### 2016-10-08 paper.tex
set up stufz

\documentclass{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{physics}
\usepackage{float}
\usepackage[margin=1.in]{geometry}
\usepackage{tocloft}
\usepackage[doublespacing]{setspace}
\renewcommand{\cfttoctitlefont}{\hfill\Large\bfseries}
\renewcommand{\cftaftertoctitle}{\hfill\hfill}
\renewcommand{\cftloftitlefont}{\hfill\Large\bfseries}
\renewcommand{\cftafterloftitle}{\hfill}
\renewcommand{\cftlottitlefont}{\hfill\Large\bfseries}
\renewcommand{\cftafterlottitle}{\hfill}
% These 6 lines will center the titles of the table of contents, list of figures and list of tables

\begin{document}

\pagenumbering{arabic}
\begin{titlepage}
\addcontentsline{toc}{section}{Abstract}
\begin{center}
\vspace{1cm}
\huge
\textbf{Nuclear energy: something something}\\
\LARGE
\vspace{.25cm}
Senior Comps Project\\
\vspace{.25cm}
\Large
\textbf{Yuping Huang}\\
\vspace{.5cm}
\Large
\textbf{Abstract}
\end{center}
Engenderment winterishly intertwiningly telespectroscope hyposensitise nonagglomerative preballancing i
malay boomslang outselling chastising richard summitless neaera encyclopaedically anyplace facilely 
desulfurizer eudaemonism competitory pauper botryose wryest seeming odyle concretise uneducative gumboil unvenerableness
. Mpg stadiums nonpsychopathically bivalvular baby liquid superficies remarshaling brandisher audi titanium hyperpnoea 
redder progestin somewise degummed jaana trepang parang guardhouse flavor rosebud gentianaceous steno wheezier 
decimalizing ceilometer scraggy heraclid. Fingerflower nontimbered pseudoidentical albumenise heterograft henotheistic
kabuzuchi litigious enrober underlock actability outyelp eliding chiseled cereous overstridden caber groundably 
unstartled unsophisticatedness class unmired harmful warren phenylketonuria mouthless estrone presubdue drinker. 
Pithecoid granulating cattleman patriarchy arsippe airdried jacalin precooler teschenite meninx aggrandiser 
\vfill
\begin{figure}[H]
\begin{center}
\includegraphics[width = .2\textwidth]{carleton.pdf}
\end{center}
\end{figure}
\begin{center}	
    \begin{onehalfspacing}
    Supervisor:  Beep Bop\\
    Department of Physics and Astronomy\\
    Carleton College\\
    Winter 2017\\
    \end{onehalfspacing} 
}
\end{center}
\end{titlepage}
\section{Introduction}

\section{oh hey}
section ehhhh
\section{acklowledgments}
\pagebreak
\pagenumbering{gobble}
\addcontentsline{toc}{section}{References}
\bibliographystyle{ieeetr}
\nocite{*}
% \bibliography{biblio.bib}
% change the file name to line up with your bibtex file
\end{document}

### 2016-11-15 paper.tex
updated bibs

@@
-\documentclass{article}
+\documentclass[usenatbib]{article}
 
@@
     \begin{onehalfspacing}
-    Supervisor:  Beep Bop\\
+    Supervisor:  Joel M. Weisberg\\
     Department of Physics and Astronomy\\
@@
     \end{onehalfspacing} 
-}
 \end{center}
@@
 
-\section{oh hey}
-section ehhhh
-\section{acklowledgments}
+\section{things}
+\begin{itemize}
+    \item history
+    \item nuclear decay mechanism: weak force and tunnelling, radioactive decay
+    \item What is nuclear waste and why is it harmful?
+    \item Above-ground disposal: look at sites, dry-casks storage
+    \item geological disposal ($\star$), Yucca mountain as an example
+    \item ocean floor disposal which was done by everyone but not legal any more
+    \item transmutation
+    \item todo: find sources on the specific health impact of nuclear waste instead of radiation
+        in general
+\end{itemize}
+\section{Acknowledgments}
+\nocite{*}
 \pagebreak
 \pagenumbering{gobble}
-\addcontentsline{toc}{section}{References}
-\bibliographystyle{ieeetr}
-\nocite{*}
-% \bibliography{biblio.bib}
-% change the file name to line up with your bibtex file
+\bibliographystyle{aip}
+\bibliography{refs}
 \end{document}

### 2017-01-04 paper_backup.tex
split body and title page for formatting

--- 2016-11-15 paper.tex
@@
-\documentclass[usenatbib]{article}
+%\documentclass{article}
+\documentclass[preprint,aip,cha]{revtex4-1}
 
 \usepackage[english]{babel}
+%\usepackage[super]{natbib}
+\usepackage{hyperref}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{graphicx}
+\graphicspath{ {images/} }
 \usepackage{physics}
@@
 \usepackage{tocloft}
-\usepackage[doublespacing]{setspace}
+\usepackage{color}
+\usepackage{setspace}
+
+\usepackage{titlesec}
 \renewcommand{\cfttoctitlefont}{\hfill\Large\bfseries}
@@
 \renewcommand{\cftafterlottitle}{\hfill}
+\newcommand{\red}[1]{\textcolor{red}{#1}}
 % These 6 lines will center the titles of the table of contents, list of figures and list of tables
@@
 \huge
-\textbf{Nuclear energy: something something}\\
+\textbf{Nuclear waste: something something}\\
 \LARGE
@@
 \begin{center}	
-    \begin{onehalfspacing}
     Supervisor:  Joel M. Weisberg\\
@@
     Winter 2017\\
-    \end{onehalfspacing} 
 \end{center}
 \end{titlepage}
+
 \section{Introduction}
+IAEA nuclear waste estimation: over 1000 tons of highly radioactive waste. \cite{iaea08}.
+Nuclear power estimate:increase by 40\% by 2020. \cite{iaea12}
+meep. Mostly focus on nuclear waste from reactors.
+global policy review \cite{r12}
 
-\section{things}
 \begin{itemize}
@@
     \item nuclear decay mechanism: weak force and tunnelling, radioactive decay
+    \item reactors?
     \item What is nuclear waste and why is it harmful?
+    \item partition and transmutation
     \item Above-ground disposal: look at sites, dry-casks storage
@@
 \end{itemize}
+
+\section{Reactors, nuclear fuel cycle, radioactivity, why harmful}
+
+\section{Formation of (High-Level) Nuclear Waste}
+    The IAEA classifies nuclear waste by considering the activity content and the half-life
+    of the radionuclides (see Figure \ref{fig:scheme}). Activity content is a measure of how much nuclear activity
+    a given amount of wastes contain: a form of waste may possesses a high activity content due to
+    a high concentration or high activity.\cite{iaea09}
+    \begin{figure}[H]
+        \centering
+        \includegraphics[width=0.5\textwidth]{wastescheme.png}
+        \caption{An illustration of the IAEA classification scheme for nuclear wastes.\cite{iaea09}}
+        \label{fig:scheme}
+    \end{figure}
+
+    \subsection{Uranium tailings}
+    ?
+
+    \subsection{Low-level wastes}
+    small amount of radioactivity
+
+    \subsection{Intermediate-level wastes}
+    higher radioactivity, needs shielding not cooling
+
+    \subsection{High-level wastes}
+    nuclear reactors. 3\% of the volume but 95\% of the radioactivity \red{(source needed)}
+
+\section{Impacts}
+
+\section{Disposal}
+
+\section{Nuclear Decay}
+k
+
 \section{Acknowledgments}
-\nocite{*}
+Thanks Joel and peeps. Arjendu and M. V. Ramana.
+
 \pagebreak

### 2017-01-05 paper.tex
initialize things
daily

@@
-\documentclass[usenatbib]{article}
-
-\usepackage[english]{babel}
-\usepackage[utf8]{inputenc}
+%\documentclass{article}
+\documentclass[preprint,aip,cha]{revtex4-1}
+\usepackage{hyperref}
 \usepackage{amsmath}
 \usepackage{graphicx}
+\graphicspath{ {images/} }
 \usepackage{physics}
 \usepackage{float}
-\usepackage[margin=1.in]{geometry}
-\usepackage{tocloft}
-\usepackage[doublespacing]{setspace}
-\renewcommand{\cfttoctitlefont}{\hfill\Large\bfseries}
-\renewcommand{\cftaftertoctitle}{\hfill\hfill}
-\renewcommand{\cftloftitlefont}{\hfill\Large\bfseries}
-\renewcommand{\cftafterloftitle}{\hfill}
-\renewcommand{\cftlottitlefont}{\hfill\Large\bfseries}
-\renewcommand{\cftafterlottitle}{\hfill}
-% These 6 lines will center the titles of the table of contents, list of figures and list of tables
+\usepackage{color}
+\newcommand{\red}[1]{\textcolor{red}{#1}}
 
@@
 
-\pagenumbering{arabic}
-\begin{titlepage}
-\addcontentsline{toc}{section}{Abstract}
-\begin{center}
-\vspace{1cm}
-\huge
-\textbf{Nuclear energy: something something}\\
-\LARGE
-\vspace{.25cm}
-Senior Comps Project\\
-\vspace{.25cm}
-\Large
-\textbf{Yuping Huang}\\
-\vspace{.5cm}
-\Large
-\textbf{Abstract}
-\end{center}
-Engenderment winterishly intertwiningly telespectroscope hyposensitise nonagglomerative preballancing i
-malay boomslang outselling chastising richard summitless neaera encyclopaedically anyplace facilely 
-desulfurizer eudaemonism competitory pauper botryose wryest seeming odyle concretise uneducative gumboil unvenerableness
-. Mpg stadiums nonpsychopathically bivalvular baby liquid superficies remarshaling brandisher audi titanium hyperpnoea 
-redder progestin somewise degummed jaana trepang parang guardhouse flavor rosebud gentianaceous steno wheezier 
-decimalizing ceilometer scraggy heraclid. Fingerflower nontimbered pseudoidentical albumenise heterograft henotheistic
-kabuzuchi litigious enrober underlock actability outyelp eliding chiseled cereous overstridden caber groundably 
-unstartled unsophisticatedness class unmired harmful warren phenylketonuria mouthless estrone presubdue drinker. 
-Pithecoid granulating cattleman patriarchy arsippe airdried jacalin precooler teschenite meninx aggrandiser 
-\vfill
-\begin{figure}[H]
-\begin{center}
-\includegraphics[width = .2\textwidth]{carleton.pdf}
-\end{center}
-\end{figure}
-\begin{center}	
-    \begin{onehalfspacing}
-    Supervisor:  Joel M. Weisberg\\
-    Department of Physics and Astronomy\\
-    Carleton College\\
-    Winter 2017\\
-    \end{onehalfspacing} 
-\end{center}
-\end{titlepage}
 \section{Introduction}
-
-\section{things}
+IAEA nuclear waste estimation: over 1000 tons of highly radioactive waste.\cite{iaea08}
+Nuclear power estimate:increase by 40\% by 2020. \cite{iaea12}
+meep. Mostly focus on nuclear waste from reactors.
+global policy review \cite{r12} 
 \begin{itemize}
@@
 \end{itemize}
-\section{Acknowledgments}
-\nocite{*}
+
+\section{Nuclear Physics}
+    Three principle prossess happen in a reactor: scattering, radiative capture and fission.
+    In each of these processes neutrons are involved.
+    \subsection{Radioactivity(and how to quantify it)}
+        \subsubsection{Quantum Theory of Radioactive decays}
+        some QM to explain nuclear excitation and ways to change into more stable system.
+        nuclear instability
+    \subsection{Properties of a neutron}
+        Because neutron plays a central role in the physical processes taking
+        place in a recator, we should first look at some of its
+        properties. A free neutron has a half-life of about $10.2$ minutes. \cite{gc01} The Standard
+        Model dictates that a free neutron can only go through beta decay, in which a neutron decays
+        into a proton, an electron, and an electron antineutrino \red{(source)}. The decay can be denoted by
+        \[n^0 \rightarrow p^+ + e^- + \bar{\nu}_e,\]
+        where $n$ denotes a neutron, $p^+$ a proton (that is positively charged), $e^-$ a
+        (negatively charged) electron and $\bar{\nu}_e$ an electron antineutrino.
+
+        While a free neutron decays relatively quickly, a neutron bound to a proton is usually stable.
+        A neutron may go through beta decay in an unstable nucleus formed by neutrons and protons(\red{more}).
+
+        A neutron can also be captured by a nucleus as it travels through materials. Since a neutron is neutrally
+        charged, it requires much less energy to combine with a nucleus than does
+        an alpha particle since it does not experience the Coulumb barrier.
+        Neutron capture usually produces a neutron that is unstable and thus
+        more radioactive than the stable nucleus found in nature.
+
+    \subsection{Neutron Capture}
+        \red{cross section?}
+    \subsection{Fission}
+        \subsubsection{Fission and fission products}
+        Conjectured by Bohr and proven by Alfred O. Nier in 1939,
+        the fission of uranium-235 induced by a thermal neutron
+        is the best-known example of induced fission. 
+        The neutron capture produces a compound nucleus ${}^{236}U$ that is in an excited state.
+        Sometimes the nucleus can decay by $\gamma$ emission in capture reaction. \red{(?)}
+        But most of the time, the excitation energy deforms the nucleus (see Figure \ref{fig:fission}).
+        When the
+        deformation reaches a certain point, the Coulumb force overcomes the short-ranged nuclear
+        force (residual strong force) and the nucleus disintegrate into two large fragments and several neutrons. The
+        positively-charged fragments(fission products) attain approximately $170$ MeV of kinetic energy as
+        Coulumb force continues to drive them apart.\cite{l01}
+        A typical reaction is described by
+        \[{}^{235}_{92}\text{U} + n \rightarrow {}^{92}_{36}\text{Kr} + {}^{142}_{56}\text{Ba} + 2 n.\]
+        The energy liberated is about $180$ MeV. The combination of fission products is not
+        unique. \cite{w98, gc01}
+        Very often, photons, electrons and nutrinos are also emitted through beta and gamma decay.
+        \red{(insert 10.2 from Lilley here? distribution of fission fragment masses)}
+        Radioactive if neutron-to-proton ratio is over 5. On average, $2.5$ neutrons are produced per fission.\cite{l01} 
+        \begin{figure}[H]
+            \centering
+            \includegraphics[width=0.8\textwidth]{fission.png}
+            \caption{Illustration of the fission process for uranium-235.\cite{l01}}
+            \label{fig:fission}
+        \end{figure}
+
+        \subsubsection{Fission energy budget}
+        We can estimate the amount of energy released in fission by considering the binding energy difference
+        and the properties of the product.
+        The binding energy per nucleon is about $7.6$ MeV u${}^{-1}$ for uranium and about $8.5$ MeV u${}^{-1}$
+        for a nucleus with atomic number around $117$. Hence, the change of binding energy per nucleon is
+        $0.9$ MeV u${}^{-1}$, which amount to $212$ MeV for uranium-235.
+        About $87\%$ of the total energy is promptly emitted because
+        the fission products, carrying over $90\%$ of the total energy, travel for only a fraction of a
+        millimeter before they are stopped. Besides, fission neutrons emit about $2$ MeV of energy by
+        successive scattering in the reactor.
+        About $13\%$ of the total energy goes into radioactivity. Energy carried by electrons and photons is
+        convertible to heat but the nutrinos, carrying about $50\%$ of the radiation energy, are not stoppable.
+
+        Henceforth, it is generally accepted that about $220$ MeV per fission is recoverabe for energy conversion.
+        This is about 2.5 million times the energy produced by burning coal of the same mass.\cite{e17}
+        \subsubsection{Chain reaction??}
+
+\section{Effect of Radiation on Human}
+    meep
+
+\section{Formation of (High-Level) Nuclear Waste}
+    The IAEA classifies nuclear waste by considering the activity content and the half-life
+    of the radionuclides (see Figure \ref{fig:scheme}). Activity content is a measure of how much nuclear activity
+    a given amount of wastes contain: a form of waste may possesses a high activity content due to
+    a high concentration or high activity.\cite{iaea09}
+    \begin{figure}[H]
+        \centering
+        \includegraphics[width=0.7\textwidth]{wastescheme.png}
+        \caption{An illustration of the IAEA classification scheme for nuclear wastes.\cite{iaea09}}
+        \label{fig:scheme}
+    \end{figure}
+
+    High-level Wastes (HLW) are highly radioactive and usually
+    hot. The typical source of HLW is spent fuel from the nuclear reactor. In a nuclear reactor,
+    fission of uranium-235 produces lighter radionuclides like cesium-137 and strontium-90, which
+    account for most of the heat and penetrating radiation (\red{penerating?}). In addition to
+    the fission process, some uranium-235 atoms go through neutron capture and produce elements that are
+    heavier than uranium ("transuranic") such as plutonium. The transuranic atoms do not generate
+    as much heat or radiation as the fission products, but they tend to have a much longer half-life.
+    For example, strontium-90 and cesium-137 have half-lives of about $30$ years, whereas
+    plutonium-239 has a half-life of $24,000$ years. https://www.nrc.gov/reading-rm/doc-collections/fact-sheets/radwaste.html
+    \red{transition needed}
+
+    \subsection{Fission products}
+    meep
+    \subsection{Transuranic radionuclides}
+    mop
+    \subsection{Trade-offs}
+    short huge dose or long small dose?
+
+\section{Impacts}
+
+\section{Disposal}
+
+\section{Nuclear Decay}
+k
+\begin{acknowledgments}
+Thanks Joel and peeps. Arjendu and M. V. Ramana.
+\end{acknowledgments}
+
 \pagebreak
-\pagenumbering{gobble}
 \bibliographystyle{aip}

### 2017-01-09 paper.tex
radioactivity and bio impact

@@
 \usepackage{physics}
-\usepackage{float}
 \usepackage{color}
-\newcommand{\red}[1]{\textcolor{red}{#1}}
+\newcommand{\red}[1]{\textcolor{red}{\bf #1}}
 
@@
 global policy review \cite{r12}
+
+Where they come from:\\
+mop
+
+what to do with them:\\
+meep
+
 \begin{itemize}
@@
 
-\section{Nuclear Physics}
-    Three principle prossess happen in a reactor: scattering, radiative capture and fission.
-    In each of these processes neutrons are involved.
-    \subsection{Radioactivity(and how to quantify it)}
-        \subsubsection{Quantum Theory of Radioactive decays}
-        some QM to explain nuclear excitation and ways to change into more stable system.
-        nuclear instability
+\section{Radioactivity and nuclear fission}
+    Here we detail the nuclear physics underlying most of the things we will talk about
+    in this work. We will apply these results in our discussion of nuclear wastes.
+    \subsection{Radioactivity}
+        \subsubsection{Nuclear shell model}
+        By solving Schr\"{o}dinger's equation\cite{k88}
+
+        \subsubsection{Quantum theory of radioactive decays}
+        We have learned that radioactivity is stochastic in nature and radioactive decay follows
+        an exponential law
+        \begin{equation} \label{eq:exp}
+            \lambda = -\frac{(dN/dt)}{N},
+        \end{equation}
+        where $\lambda$ is the decay constant, and the right hand side the probability of decay
+        per time for an atom. The fundamental assumption is that the probability of decay is constant,
+        regardless of the age of the atom. This "memoryless" assumption naturally leads to the
+        exponential law of radioactive decay
+        \begin{equation}
+            N(t) = N_0 e^{-\lambda t},
+        \end{equation}
+        where $N_0$ gives the number of nuclei at $t=0$.
+
+        Define activity $\mathcal{A}(t)$ to be the rate of decays in a sample
+        \begin{equation} \label{eq:act}
+            \mathcal{A}(t)\equiv \lambda N(t).
+        \end{equation}
+        Here we outline an quantum mechanics argument for an exponential decay law that is perhaps more
+        convincing.
+        We approximate the potential by $V+V_p$ where $V$ is the nuclear potential and $V_p$ is a small
+        time-dependent perturbation that causes the transition between the states. We know that there has
+        to be time dependence in the potential, otherwise for a superposition of eigenstates
+        $\ket{psi}=c_1\ket{\psi_1} + c_2\ket{\psi_2}$, the coefficients $c_1$ and $c_2$ would 
+        have to be constant and
+        persistent decay of one state into the other would not happen.
+        Fermi's Golden Rule \#2 (which is derived in appendix \ref{a:fermi})
+        gives the transition rate
+        \begin{equation}
+            \lambda = \frac{2\pi}{2\hbar}|\bra{\psi_f}V_p\ket{\psi_i}|^2\rho(E_f),
+        \end{equation}
+        where $\ket{\psi_i}$ denotes the initial state , $\bra{\psi_f}$ the final state, $V_p$ the perturbation
+        potential and $\rho(E_f)$ the density of state of the final state, defined by
+        \begin{equation}
+            \rho(E_f) = \frac{dn_f}{dE_f}.
+        \end{equation}
+        Note that as in perturbation theory, $\ket{\psi_i}$ and
+        $\ket{\psi_f}$ are both eigenstates of the nuclear potential $V$. Therefore, as long as
+        the decay probability is small, the result from perturbation theory should hold and the
+        decay rate should be independent of time, which is the fundamental assumption made in
+        equation~\ref{eq:exp}.
+
+        \subsubsection{Quantifying radioactivity}
+        The radioactivity of a sample $\mathcal{A}$(defined in equation~\ref{eq:act}) 
+        is the number of decays of the sample per unit time. The SI
+        unit for radioactivity is becquerel (Bq) and it equals one decay one second. Another commonly-used
+        unit is curie (Ci, with $1\text{ Ci}=3.7\times 10^{10}\text{ Bq}$) and it is defined as
+        the activity of radium-266. A fission product, cesium-137 has an activity of $3.215\times 10^{12}$ Bq
+        per gram. By contrast, the radioactivity of a human body (mainly due to the small amount of 
+        potassium-40 and carbon-14) is about $8000$ Bq.
+
+    \subsection{Biological Effects of Radiation}
+        The biological effects of radiation have been an important topic for decades. A large
+        number of studies took place after the two nuclear bomb explosions in Japan in 1945
+        and after disastrous nuclear plant accidents like Chernobyl. Therefore, we have a relatively
+        thourough understanding of the consequences of mass exposure to radiation. However, relatively
+        little is known about the long-term effects of exposure to small doses, including subtle biological
+        changes that lead to cancer and genetic deffects. Here we briefly summarize the known effects.
+        \subsubsection{Different kind of radiation?}
+        \subsubsection{Quantifying radiation on living tissues}
+        The biological impact of radiation depends on the type, power and energy of the radiation, the
+        exposure time, and the part of the body exposed. Various quantities have been introduced to
+        measure the amount of radiation and the biological impact of radiation. Here we review three
+        of these quantities: the absorbed dose, the equivalent dose and the effective dose.
+
+        The amount of radiation is usually characterized by the amount of energy it carries. Since one of
+        the ealiest understood effects of radiation is its ability to ionize gas, one can quantify the
+        strength of the radiation with the ionizing energy of the gas.
+        Absorbed dose~($D$) quantifies the 
+        radiation energy absorbed per unit mass. The modern unit is Gray(Gy)
+        and $1\text{Gy}=1\text{J kg}^{-1}$.\cite{u16}
+
+        Biological effects not only depend on the total dose applied but also the rate at which the 
+        dose is applied, because mechanisms exist within organisms that allow molecules like deoxyribonucleic acid
+        to be repaired if they are not excessively damaged. Therefore {\it dose rate} is also a useful
+        measure of radiation when considering its biological effect.
+
+        For an external $\gamma$-ray source emitting $\mathcal{A}$ million photons per second 
+        of energy $E_{\gamma}$
+        (in MeV), one can approximate the dose rate by 
+        \begin{equation}
+            \frac{dD}{dt}(\mu\text{Sv h}^{-1}) \approx \frac{\mathcal{A}\text{(MBq)}\times E_{\gamma}\text{(MeV)}}
+            {6\times [r\text{(m)}]^2}.
+        \end{equation}
+        The approximate\red{(?)} linear dependence on $E_{\gamma}$ is valid from less than $0.1$ MeV to several
+        MeV and therefore should be good for all sources but those generated in an accelerator.\cite{my68}
+        The inverse-squared dependence on distance is hardly surprising.
+
+        However, absorbed dose does not take into account different biological implications of different types
+        of radiation.
+        For example, the same dose of alpha particle can do more damage to body than gamma rays. The equivalent
+        dose ($H$) is defined as
+        \[H=w_R \times D_R,\]
+        where $D_R$ is the absorbed dose and $w_R$ the weighting factor of a given kind of radiation $R$
+        (see table \ref{tab:eq} for a list of weighing factors). If there are multiple types of radiation 
+        present, the equivalent does is given by the weighted sum over all contributions. The SI unit for
+        equivalent dose is called the sievert (Sv) and $1\text{ Sv}=1\text{ J kg}^{-1}$.
+        \\\red{Dose distribution and bio effectiveness?}
+        \begin{table}
+            \label{tab:eq}
+            \centering
+            \caption{Weighting factors for different types of radiation\cite{icrp74}}
+            \begin{ruledtabular}
+                \begin{tabular}{l c c}
+                Type of radiation & Energy range & Weighting factor, $w_R$\\
+                \hline
+                Photons, electrons & All energies & 1\\
+                Neutrons & $<10$ keV & 5 \\
+                         & $10-100$ keV & 10 \\
+                         & $100$ keV-$2$ MeV & 20 \\
+                         & $2-20$ MeV & 10 \\
+                Protons & $<20$ MeV & 5 \\
+                Alpha particles, fision fragments, heavy nuclei & & $20$\\
+            \end{tabular}
+            \end{ruledtabular}
+        \end{table}
+        Finally, certain organs and parts of the body are more vulnerable to radiation than the rest.
+        Therefore, effective dose~($E$) is defined as the equivalent dose weighted by the body part
+        radiated
+        \[E = w_T H_T.\]
+        Evidently, if the radiation affects multiple tissues, the effective dose will be the weighted
+        sum of the equivalent dose over individual body parts. Some of the recommended values of tissue
+        weighting factors are given in table \ref{tab:eff}. A weighing factor of $1$ is equivalent to the
+        whole body absorbing the radiation uniformly. The SI unit of effective dose is sievert (Sv),
+        identical to that of equivalent dose.
+        \begin{table}
+            \label{tab:eff}
+            \centering
+            \caption{Weighting factors for individual organs\cite{icrp74}}
+            \begin{ruledtabular}
+                \begin{tabular}{l c}
+                Tissue & Weighting factor, $w_T$\\
+                \hline
+                Gonads & $0.20$\\
+                Red bone marrow & $0.12$ \\
+                Liver & $0.05$ \\
+                Skin & $0.01$ \\
+                \end{tabular}
+            \end{ruledtabular}
+        \end{table}
+        \subsubsection{Effects of radiation on human}
+        Short-term effects of radiation are usually caused by extensive cell death or cell damage.\cite{u16}
+        These effects are deterministic, meaning that below a certain threshold the effect does not occur and
+        the severity of the effect depends on the dose. Some examples are gastrointestinal damage and impairment
+        of fertility.\cite{u16, l01}
+
+        Delayed health effects occur a long time after the exposure and are usually stochastic. There is
+        no apparent threshold and the severity is usually independent of the dose but
+        the chance of occurance depends on the dose.\cite{u16,l01}
+        For example, according to a review done by the UN Scientific Committee on the Effects of Atomic Radiation
+        (UNSCEAR), the additional chance of dying of cancer due to radiation exposure above $100$ mSv is
+        about 3 to 5 percent per sievert.\cite{unscear12}
+        \red{LNT vs threshold?}
+
+        Hereditary effects of radiation can occur if the reproduction cells were damaged by radiation.
+        UNSCEAR estimates that risk for severe hereditary effect is about $0.3-0.5$ percent per gray to
+        the first generation following the radiation exposure.\cite{u16, unscear12}
+
+        \subsubsection{Limitations of current studies}
     \subsection{Properties of a neutron}
@@
         place in a recator, we should first look at some of its
-        properties. A free neutron has a half-life of about $10.2$ minutes. \cite{gc01} The Standard
+        properties.
+        \subsubsection{Free neutron}
+        A free neutron has a half-life of about $10.2$ minutes. \cite{gc01} The Standard
         Model dictates that a free neutron can only go through beta decay, in which a neutron decays
@@
 
-        While a free neutron decays relatively quickly, a neutron bound to a proton is usually stable.
-        A neutron may go through beta decay in an unstable nucleus formed by neutrons and protons(\red{more}).
+        \subsubsection{Neutron Absorption}
+        easy to absorb because neutrally charged
+        The nuclear absorption reaction is the most crucial type of reactions that takes place in
+        a nuclear reactor. As neutrons travels through materials, they react readily with nuclei.
+        For fast neutrons, many reactions are possible. But for slow or thermal neutrons (as in the
+        case of a nuclear reactor), neutron absorption is the primary mode of interaction.
+        In the process of neutron absorption, a nucleus absorbs a neutron to form
+        a compound nucleus. Depending on the path that the compound nucleus takes to return to its ground
+        state, neutron absorption is classified into radiative capture and neutron-induced
+        fission(see section~\ref{sec:fission}).
 
-        A neutron can also be captured by a nucleus as it travels through materials. Since a neutron is neutrally
-        charged, it requires much less energy to combine with a nucleus than does
-        an alpha particle since it does not experience the Coulumb barrier.
-        Neutron capture usually produces a neutron that is unstable and thus
-        more radioactive than the stable nucleus found in nature.
+        In a radiative capture process, the compound nucleus emits photons, whereas in neutron-induced
+        fission, the compound nucleus splits into fragments.
+        In fact, not all nuclei can undergo fission.
+        Therefore, radiative capture is the only possible neutron absorption process for
+        nuclei that are not fissionable.\cite{l66} We will explore neutron-induced fission further in
+        section~\ref{sec:fission}. Here we give an example of radiative capture. When a stable form
+        of gold ${}^{197}$Au absorbs a neutron, it forms a highly unstable isotope ${}^{198}$Au in its
+        excited state and quickly decays into its ground state by emitting $\gamma$ rays. This can be
+        written as
+        \[{}^{197}\text{Au} + n \rightarrow {}^{198}\text{Au} + \gamma.\]
+        Moreover, the isotope ${}^{198}$Au can spontaneously decay into ${}^{198}$Hg via beta decay. This process can
+        be seen as the analogy of free neutron beta decay for bound neutrons.
 
-    \subsection{Neutron Capture}
+        i.e. how does a neutron excite a nucleus?
+        \red{fission vs radiative capture condition?}
         \red{cross section?}
+        \red{pic of the two potentials}
+
     \subsection{Fission}
+    \label{sec:fission}
+        Nuclear Fission is the process in which a heavy nucleus splits into smaller fragments and releases
+        a huge amount of energy. It is the main physical process taking place in a nuclear reactor. It is
+        also the main mechanism that generates the most radiative wastes that we will discuss later
+        in this work. Therefore, we should first familiarize ourselves with fission.
+        \red{fissile vs fissionable}
+
         \subsubsection{Fission and fission products}
@@
         Sometimes the nucleus can decay by $\gamma$ emission in capture reaction. \red{(?)}
-        But most of the time, the excitation energy deforms the nucleus (see Figure \ref{fig:fission}).
+        But most of the time, the excitation energy deforms the nucleus (see FIG. \ref{fig:fission}).
         When the
@@
         Radioactive if neutron-to-proton ratio is over 5. On average, $2.5$ neutrons are produced per fission.\cite{l01} 
-        \begin{figure}[H]
+        \begin{figure}[h]
             \centering
@@
         \end{figure}
+
+        \red{Wong has more on the fragmentation thing in his fission chapter.}
 
@@
         Henceforth, it is generally accepted that about $220$ MeV per fission is recoverabe for energy conversion.
-        This is about 2.5 million times the energy produced by burning coal of the same mass.\cite{e17}
+        The energy genreated by the fission of uranium-235 is 
+        about 2.5 million times the energy produced by burning coal of the same mass.\cite{e17}
         \subsubsection{Chain reaction??}
-
-\section{Effect of Radiation on Human}
-    meep
 
@@
     The IAEA classifies nuclear waste by considering the activity content and the half-life
-    of the radionuclides (see Figure \ref{fig:scheme}). Activity content is a measure of how much nuclear activity
+    of the radionuclides (see FIG. \ref{fig:scheme}). Activity content is a measure of how much nuclear activity
     a given amount of wastes contain: a form of waste may possesses a high activity content due to
     a high concentration or high activity.\cite{iaea09}
-    \begin{figure}[H]
+    \begin{figure}[h]
         \centering
@@
     account for most of the heat and penetrating radiation (\red{penerating?}). In addition to
-    the fission process, some uranium-235 atoms go through neutron capture and produce elements that are
+    the fission process, some uranium-235 atoms go through radiative capture and produce elements that are
     heavier than uranium ("transuranic") such as plutonium. The transuranic atoms do not generate
@@
 
-\section{Impacts}
-
 \section{Disposal}
 
-\section{Nuclear Decay}
+\section{Future and perspective?}
+
+\section{Discussions}
+
 k
@@
 \bibliography{refs}
+\appendix
+\section{Derivation of Fermi's Golden Rule \#2}
+\label{a:fermi}
+\section{meep}
 \end{document}

### 2017-01-11 paper.tex
outline done

@@
 %\documentclass{article}
-\documentclass[preprint,aip,cha]{revtex4-1}
+\documentclass[preprint,aip,pra]{revtex4-1}
 \usepackage{hyperref}
@@
 \begin{document}
+
+\title{Nuclear Wastes bla bla bla}
+
+\author{Yuping Huang}%
+\affiliation{Department of Physics and Astronomy, Carleton College}
+
+\date{\today}%
+\begin{abstract}
+    meep
+\end{abstract}
+\maketitle
+
+\tableofcontents
+\newpage
 
@@
     \subsection{Radioactivity}
-        \subsubsection{Nuclear shell model}
+    \subsubsection{Nuclear shell model vs liquid drop model(?)}
         By solving Schr\"{o}dinger's equation\cite{k88}
+    \subsubsection{Binding Energy(?)}
 
@@
         changes that lead to cancer and genetic deffects. Here we briefly summarize the known effects.
-        \subsubsection{Different kind of radiation?}
+        \subsubsection{Different kinds of radiation and shielding}
+        meep
         \subsubsection{Quantifying radiation on living tissues}
@@
         into a proton, an electron, and an electron antineutrino \red{(source)}. The decay can be denoted by
-        \[n^0 \rightarrow p^+ + e^- + \bar{\nu}_e,\]
+        \begin{equation}
+            n^0 \rightarrow p^+ + e^- + \bar{\nu}_e,
+        \end{equation}
         where $n$ denotes a neutron, $p^+$ a proton (that is positively charged), $e^-$ a
@@
         written as
-        \[{}^{197}\text{Au} + n \rightarrow {}^{198}\text{Au} + \gamma.\]
+        \begin{equation}
+            {}^{197}\text{Au} + n \rightarrow {}^{198}\text{Au} + \gamma.
+        \end{equation}
         Moreover, the isotope ${}^{198}$Au can spontaneously decay into ${}^{198}$Hg via beta decay. This process can
@@
 \section{Formation of (High-Level) Nuclear Waste}
+    \subsection{Classification of Nuclear Waste}
     The IAEA classifies nuclear waste by considering the activity content and the half-life
@@
 
-    \subsection{Fission products}
+    \subsection{Comments on lower-level waste and how to deal with them}
+    \subsection{Compositions and risks of high-level waste}
+    \subsubsection{Fission products(needs shielding)}
         meep
-    \subsection{Transuranic radionuclides}
+        \subsubsection{Transuranic radionuclides(really bad, needs long-term storage)}
         mop
-    \subsection{Trade-offs}
-    short huge dose or long small dose?
 
-\section{Disposal}
+\section{Re-processing and storage of HLW}
+    \subsection{Re-processing of HLW: France and UK}
+    \subsection{Requirements for storing HLW}
+    \subsection{At-reactor storage}
+        \subsubsection{Dry cask storage}
+    \subsection{Interim storage}
+    \subsection{Necessity of a next-step}
 
-\section{Future and perspective?}
+\section{Disposal of HLW}
+    \subsection{Comments on partitioning and transmutation}
+    \subsection{Survey of possibilities}
+    \subsubsection{mention ocean disposal and why it was banned('Ndrangheta)}
+    \subsection{Necessity of geological disposal sites}
+    \subsection{Requirements of disposal from physics}
+
+\section{Deep Geological Repository}
+    \subsection{something general}
+    \subsection{Yucca Mountain}
+    \subsection{Something related to the French Ciegeo project: licensed reversible disposal site}
+    \subsection{Risk calcs: geophysics, mention WIPP}
 
@@
 \label{a:fermi}
-\section{meep}
 \end{document}

### 2017-01-14 paper.tex
outline joel comments

@@
 
-\section{Formation of (High-Level) Nuclear Waste}
+\section{Nuclear Waste}
     \subsection{Classification of Nuclear Waste}
-    The IAEA classifies nuclear waste by considering the activity content and the half-life
+    Different organizations have slightly different classification schemes for nuclear waste.
+    The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the activity content and the half-life
     of the radionuclides (see FIG. \ref{fig:scheme}). Activity content is a measure of how much nuclear activity

### 2017-01-19 paper.tex
meep

@@
 \maketitle
-
 \tableofcontents
@@
     \item history
-    \item nuclear decay mechanism: weak force and tunnelling, radioactive decay
+    \item nuclear decay mechanism: weak force and tunneling, radioactive decay
     \item reactors?
@@
         per time for an atom. The fundamental assumption is that the probability of decay is constant,
-        regardless of the age of the atom. This "memoryless" assumption naturally leads to the
+        regardless of the age of the atom. This "memorylessness" assumption naturally leads to the
         exponential law of radioactive decay
@@
         \begin{equation}
+            \label{eq:trans}
             \lambda = \frac{2\pi}{2\hbar}|\bra{\psi_f}V_p\ket{\psi_i}|^2\rho(E_f),
@@
         $\ket{\psi_f}$ are both eigenstates of the nuclear potential $V$. Therefore, as long as
-        the decay probability is small, the result from perturbation theory should hold and the
+        the decay probability is small, the result from perturbation theory should hold. Since
+        there is no time dependency in equation~\ref{eq:trans},the
         decay rate should be independent of time, which is the fundamental assumption made in
@@
         and after disastrous nuclear plant accidents like Chernobyl. Therefore, we have a relatively
-        thourough understanding of the consequences of mass exposure to radiation. However, relatively
+        thorough understanding of the consequences of mass exposure to radiation. However, relatively
         little is known about the long-term effects of exposure to small doses, including subtle biological
-        changes that lead to cancer and genetic deffects. Here we briefly summarize the known effects.
+        changes that lead to cancer and genetic defects. Here we briefly summarize the known effects.
         \subsubsection{Different kinds of radiation and shielding}
@@
         The amount of radiation is usually characterized by the amount of energy it carries. Since one of
-        the ealiest understood effects of radiation is its ability to ionize gas, one can quantify the
+        the earliest understood effects of radiation is its ability to ionize gas, one can quantify the
         strength of the radiation with the ionizing energy of the gas.
@@
         dose ($H$) is defined as
-        \[H=w_R \times D_R,\]
+        \begin{equation}
+        H=w_R \times D_R,
+        \end{equation}
         where $D_R$ is the absorbed dose and $w_R$ the weighting factor of a given kind of radiation $R$
@@
                 Protons & $<20$ MeV & 5 \\
-                Alpha particles, fision fragments, heavy nuclei & & $20$\\
+                Alpha particles, fission fragments, heavy nuclei & & $20$\\
             \end{tabular}
@@
         radiated
-        \[E = w_T H_T.\]
+        \begin{equation}
+            E = w_T H_T.
+        \end{equation}
         Evidently, if the radiation affects multiple tissues, the effective dose will be the weighted
@@
         no apparent threshold and the severity is usually independent of the dose but
-        the chance of occurance depends on the dose.\cite{u16,l01}
+        the chance of occurrence depends on the dose.\cite{u16,l01}
         For example, according to a review done by the UN Scientific Committee on the Effects of Atomic Radiation
@@
         Because neutron plays a central role in the physical processes taking
-        place in a recator, we should first look at some of its
+        place in a reactor, we should first look at some of its
         properties.
@@
         Model dictates that a free neutron can only go through beta decay, in which a neutron decays
-        into a proton, an electron, and an electron antineutrino \red{(source)}. The decay can be denoted by
+        into a proton, an electron, and an electron anti-neutrino \red{(source)}. The decay can be denoted by
         \begin{equation}
@@
         \subsubsection{Neutron Absorption}
+        \label{sec:capture}
         easy to absorb because neutrally charged
@@
         When the
-        deformation reaches a certain point, the Coulumb force overcomes the short-ranged nuclear
+        deformation reaches a certain point, the Coulomb force overcomes the short-ranged nuclear
         force (residual strong force) and the nucleus disintegrate into two large fragments and several neutrons. The
         positively-charged fragments(fission products) attain approximately $170$ MeV of kinetic energy as
-        Coulumb force continues to drive them apart.\cite{l01}
+        Coulomb force continues to drive them apart.\cite{l01}
         A typical reaction is described by
@@
         unique. \cite{w98, gc01}
-        Very often, photons, electrons and nutrinos are also emitted through beta and gamma decay.
+        Very often, photons, electrons and neutrinos are also emitted through beta and gamma decay.
         \red{(insert 10.2 from Lilley here? distribution of fission fragment masses)}
@@
         About $13\%$ of the total energy goes into radioactivity. Energy carried by electrons and photons is
-        convertible to heat but the nutrinos, carrying about $50\%$ of the radiation energy, are not stoppable.
+        convertible to heat but the neutrinos, carrying about $50\%$ of the radiation energy, are not stoppable.
 
-        Henceforth, it is generally accepted that about $220$ MeV per fission is recoverabe for energy conversion.
-        The energy genreated by the fission of uranium-235 is 
+        Henceforth, it is generally accepted that about $220$ MeV per fission is recoverable for energy conversion.
+        The energy generated by the fission of uranium-235 is 
         about 2.5 million times the energy produced by burning coal of the same mass.\cite{e17}
@@
 
-\section{Nuclear Waste}
-    \subsection{Classification of Nuclear Waste}
+\section{Nuclear waste: classification and properties}
+    \red{how much of each is generated and what the impact is; industry vs military}\\
     Different organizations have slightly different classification schemes for nuclear waste.
-    The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the activity content and the half-life
-    of the radionuclides (see FIG. \ref{fig:scheme}). Activity content is a measure of how much nuclear activity
-    a given amount of wastes contain: a form of waste may possesses a high activity content due to
-    a high concentration or high activity.\cite{iaea09}
+    However, their definitions of high-level waste (HLW), the most detrimental kind, are relatively consistent.
+    The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the activity
+    content and the half-life of the radionuclides (see FIG. \ref{fig:scheme}).
+    Activity content is a measure of how much radioactivity
+    a given amount of waste contains: a form of waste may possesses a high activity content due to
+    \red{a high concentration of radionuclides or the presence of radionuclides with high activity.}\cite{iaea09}
+    Each form of waste, depending on the half-life, activity content, and temperature, requires different
+    disposal methods.
     \begin{figure}[h]
@@
         \includegraphics[width=0.7\textwidth]{wastescheme.png}
-        \caption{An illustration of the IAEA classification scheme for nuclear wastes.\cite{iaea09}}
+        \caption{An illustration of the IAEA classification scheme for nuclear waste along with suggested
+        disposal methods.\cite{iaea09}}
         \label{fig:scheme}
     \end{figure}
+    We will focus on high-level waste (HLW, defined below) in the rest of this work.
+    In this section, we will first examine the composition and properties of HLW and then
+    comment on other categories of waste and disposal methods.
 
-    High-level Wastes (HLW) are highly radioactive and usually
-    hot. The typical source of HLW is spent fuel from the nuclear reactor. In a nuclear reactor,
+    \subsection{Composition and risks of high-level waste}
+    High-level Wastes (HLW) has high activity contents and usually generates a large amount of
+    heat as it radiates. The typical source of HLW is spent fuel from the nuclear reactor.
+    In fact, since there is no commercial re-processing facilities in the US (see section~\ref{sec:reproc}),
+    almost all of the HLW is in the form of spent nuclear fuel from the reactor. In a nuclear reactor,
     fission of uranium-235 produces lighter radionuclides like cesium-137 and strontium-90, which
-    account for most of the heat and penetrating radiation (\red{penerating?}). In addition to
-    the fission process, some uranium-235 atoms go through radiative capture and produce elements that are
+    account for most of the heat and penetrating radiation (\red{penetrating?}). In addition to
+    the fission process, some uranium-235 atoms go through radiative capture
+    (see section~\ref{sec:capture}) and produce elements that are
     heavier than uranium ("transuranic") such as plutonium. The transuranic atoms do not generate
@@
     For example, strontium-90 and cesium-137 have half-lives of about $30$ years, whereas
-    plutonium-239 has a half-life of $24,000$ years. https://www.nrc.gov/reading-rm/doc-collections/fact-sheets/radwaste.html
-    \red{transition needed}
+    plutonium-239 has a half-life of $24,000$ years.
+    \subsubsection{Fission products}
+    hot where does that come from?
+    \subsubsection{Transuranic waste}
+    \red{ahhhhhhh}
+    Transuranic waste is defined as materials contaminated by alpha-emitting radionuclides with
+    sufficiently long half-life ($<20$ years) of elements with atomic number $92$ (atomic number of uranium)
+    or larger and concentration of activity larger than $3700$ Bq/g.\cite{j83}
 
-    \subsection{Comments on lower-level waste and how to deal with them}
-    \subsection{Compositions and risks of high-level waste}
-    \subsubsection{Fission products(needs shielding)}
-        meep
-        \subsubsection{Transuranic radionuclides(really bad, needs long-term storage)}
-        mop
+    \subsection{Low-level and intermediate-level waste}
+    Low-level waste (LLW) has relatively little radioactivity, produces almost no heat,
+    and contains practically no
+    transuranic elements. Examples are papers and tools used in cleanup operations in the reactor.
+    Most LLW does not require shielding and may be buried in landfill 
+    facilities after on-site short-term decay storage.
+    Some LLW (for example, metal nuclear fuel cladding) may have a slightly
+    higher radioactivity such that they are classified
+    as intermediate-level waste (ILW). ILL are solidified in concrete and buried in shallow
+    repositories.\cite{s01} However, US regulations do not make the distinction
+    between LLW and ILW but different disposal methods are practiced.\cite{nrc09, s01}
 
 \section{Re-processing and storage of HLW}
+    Re-processing extracts the uranium and plutonium from the spent
+    nuclear fuel.
+    \label{sec:reproc}
     \subsection{Re-processing of HLW: France and UK}
@@
     \subsection{At-reactor storage}
+        \subsection{Spent Fuel Pools}
+        They are usually about 15m deep, with storage rack at the bottom to hold the fuel assemblies.
+        The water absorbs the heat and shields the radiation.
         \subsubsection{Dry cask storage}

### 2017-01-22 paper.tex
stuck

@@
 \usepackage{amsmath}
+\usepackage{csquotes}
+\usepackage{gensymb}
 \usepackage{graphicx}
@@
 
-\title{Nuclear Wastes bla bla bla}
+\title{Nuclear Wastes: we're kinda doomed}
 
@@
 meep. Mostly focus on nuclear waste from reactors.
-global policy review \cite{r12}
+Global policy review \cite{r12}
 
@@
         per time for an atom. The fundamental assumption is that the probability of decay is constant,
-        regardless of the age of the atom. This "memorylessness" assumption naturally leads to the
+        regardless of the age of the atom. This ``memorylessness'' assumption naturally leads to the
         exponential law of radioactive decay
@@
         \subsubsection{Quantifying radioactivity}
-        The radioactivity of a sample $\mathcal{A}$(defined in equation~\ref{eq:act}) 
+        The radioactivity of a sample $\mathcal{A}$ (defined in equation~\ref{eq:act}) 
         is the number of decays of the sample per unit time. The SI
-        unit for radioactivity is becquerel (Bq) and it equals one decay one second. Another commonly-used
+        unit for radioactivity is becquerel (Bq) and it equals one decay per second. Another commonly used
         unit is curie (Ci, with $1\text{ Ci}=3.7\times 10^{10}\text{ Bq}$) and it is defined as
@@
         \begin{equation}
-            E = w_T H_T.
+            E = w_T H_T,
         \end{equation}
+        where $w_T$ is the body tissue weighting factor, and $H_T$ the equivalent dose on the tissue
+        at question.
         Evidently, if the radiation affects multiple tissues, the effective dose will be the weighted
@@
         Therefore, radiative capture is the only possible neutron absorption process for
-        nuclei that are not fissionable.\cite{l66} We will explore neutron-induced fission further in
+        nuclei that are not fissionable.\cite{lb01} We will explore neutron-induced fission further in
         section~\ref{sec:fission}. Here we give an example of radiative capture. When a stable form
@@
 
-    \subsection{Composition and risks of high-level waste}
+    \subsection{High-level waste(HLW)}
+    \red{reprocessing or not}\\
+    \red{where does the heat come from?}
     High-level Wastes (HLW) has high activity contents and usually generates a large amount of
     heat as it radiates. The typical source of HLW is spent fuel from the nuclear reactor.
-    In fact, since there is no commercial re-processing facilities in the US (see section~\ref{sec:reproc}),
+    In fact, since there is no commercial reprocessing facilities in the US (see section~\ref{sec:reproc}),
     almost all of the HLW is in the form of spent nuclear fuel from the reactor. In a nuclear reactor,
@@
     (see section~\ref{sec:capture}) and produce elements that are
-    heavier than uranium ("transuranic") such as plutonium. The transuranic atoms do not generate
+    heavier than uranium (``transuranic'') such as plutonium. The transuranic atoms do not generate
     as much heat or radiation as the fission products, but they tend to have a much longer half-life.
@@
     plutonium-239 has a half-life of $24,000$ years.
-    \subsubsection{Fission products}
-    hot where does that come from?
-    \subsubsection{Transuranic waste}
-    \red{ahhhhhhh}
-    Transuranic waste is defined as materials contaminated by alpha-emitting radionuclides with
+
+    \subsection{Transuranic waste(TRU)}
+    Transuranic waste is defined as materials that do not belong to high-level waste and are
+    contaminated by alpha-emitting radionuclides with
     sufficiently long half-life ($<20$ years) of elements with atomic number $92$ (atomic number of uranium)
-    or larger and concentration of activity larger than $3700$ Bq/g.\cite{j83}
+    or larger and concentration of activity larger than $3700$ Bq/g.\cite{j83,s01}. The activity
+    of TRU is usually low but it remains active for long period of time.
+    \red{WIPP is only for TRU}
 
-    \subsection{Low-level and intermediate-level waste}
+    \subsection{Low-level and intermediate-level waste (LLW and ILW)}
     Low-level waste (LLW) has relatively little radioactivity, produces almost no heat,
@@
 
-\section{Re-processing and storage of HLW}
-    Re-processing extracts the uranium and plutonium from the spent
-    nuclear fuel.
+    \subsection{Uranium tailings}
+    Uranium mill tailings from uranium mills are another form of waste with low activity, but
+    they are not usually classified as LLW. The tailing is what is left after the uranium has been
+    extracted from the ore. The decay of uranium-238 from the mill decays into
+    radioactive thorium and radium and these radionuclides stay in the tailing. Notably, radium
+    can decay into radon gas which is an alpha-emitter. Hence, the alpha radiation can do additional
+    damage to body if the radon enters the body as one breathes.
+    The uranium tailings leave the uranium mills in the form of a liquid sludge
+    and can dry. Therefore, it is necessary to prevent the uranium tailings
+    from contaminating water and air.
+
+    \section{Spent fuel management}
+    Spent fuel from nuclear reactors constitutes the major source of HLW. Spen fuel pools at the
+    reactors hold spen fuel cores to cool them down and isolate them for a few years. After that,
+    reprocessing can occur to extract the most radioactive plutonium and uranium isotopes and
+    convert the waste into more managable form (vitrification). However, reprocessing of commercial
+    nuclear spent fueld did not
+    happen in this country for various historical reasons. 
+    \subsection{Spent fuel pools}
+    \red{boron-10 and radiation shielding.}
+    The heat output of the spent fuel decreases by $99\%$ after the first year and by another factor of $5$
+    by the 5-year mark.\cite{aa12} Therefore, short-term cooling of spent fuel is crucial before
+    futher disposal or reprocessing can be done.
+    US nuclear power plants generally have an operation cycle of about 2 years. At the end of each operation cycle,
+    one-quater to one-third of its fuel assemblies (spent fuel) are removed. The removed spent fuel spend a minimum
+    of five years in these water pools to cool.
+    Most of the nuclear industry's spent fuel is now stored in spent fuel water pools ``at-reactor (AR)'',
+    which means either within the reactor building or in an ajacent spent fuel building linked to the reactor
+    AR pools minimize the difficulty of transporting spent fuel.\cite{iaea99}
+    US has generated about $65,000$ metric tons of spent fuel, of which $75\%$ is stored in AR pools.\cite{a11}
+    by a transfer tunnel. 
+    Spent fuel pools  are usually about 15-meter deep water pools, with storage rack at the bottom to hold the fuel assemblies.
+    The water absorbs the heat and shields the radiation. Water is the chosen medium because it is cheap and
+    can provide coooling by circulation. Moreover, water can serve as a shielding from radiation, because it has a
+    fairly high density and hence forces the radiated particles (especially the neutrons) to interact with
+    the water molecules and lose energy \red{(more precise terms needed, cladding materials)}. Water temperature is maintained
+    below 50 $^\circ$C via circulation and cooling. The integrity of the storage can persist well beyond 
+    50 years.\cite{a11, iaea99}
+
+    \subsection{Reprocessing: an option}
     \label{sec:reproc}
-    \subsection{Re-processing of HLW: France and UK}
-    \subsection{Requirements for storing HLW}
-    \subsection{At-reactor storage}
-        \subsection{Spent Fuel Pools}
-        They are usually about 15m deep, with storage rack at the bottom to hold the fuel assemblies.
-        The water absorbs the heat and shields the radiation.
-        \subsubsection{Dry cask storage}
-    \subsection{Interim storage}
-    \subsection{Necessity of a next-step}
+    After spending some time in the nuclear fuel pools and cooling down, the spent fuel may be ready
+    for reprocessing.
+    Spent nuclear fuel contains a large amount of plutonium and uranium that are suitable as nuclear fuel
+    or fissile weapon material (plutonium-239). Therefore, reprocessing the spent fuel generates plutonium and
+    uranium that may be recycled. The US currently does not have any commercial nuclear fuel
+    reprocessing facility due to the demonstrated high cost of reprocessing
+    during the West Valley Demonstration Project and water contamination issues at the Hanford Site.
+    Concerns about proliferation issues also arise, as the 
+    reprocessed plutonium-239 is the primary fissile element in the production of nuclear weapons.\cite{aa12}
+    On the other hand, France (the world leader in reprocessing technology), India, Japan, Russia, the UK
+    and China (since 2010) continues to reprocess spent fuel. To understand the details of reprocessing
+    requires advanced chemistry and is beyond the scope of this work. However, the issue remains interesting
+    and hence we provide an overview of the process as well as some of the related concerns.
+
+    Almost all of the reprocessing plants use the PUREX (Plutonium Uranium Redox EXtraction) process
+    to reprocess spent fuel. This process utilizes the differences in oxidation potential 
+    between the heavy elements (uranium and plutonium) from the other lighter fission products.
+    The process involves three steps: (a) adding appropriate solvent (usually nitric acid) to absorb and oxidize 
+    plutonium and uranium while leaving out other elements;
+    (b) adding a reducing (de-oxidizing) agent that reduces (de-oxidizes) only the
+    oxidation state of plutonium to one that is insoluble
+    and thus seperable; and finally (c) extracting the plutonium from the mixture. In this way
+    the oxidized plutonium and uranium are separated and can be reduced back to reusable radionuclides.\cite{lb01} 
+    The extracted plutonium then returns to the reactor and transforms into short-lived fission products
+    by fission. Un-reprocessed fuel usually takes 10 times as much time to reach the safe
+    activity level (\sim$10^5$ years) as reprocessed fuel (\sim$10^4$ years). Moreover, since fission products
+    make up a small portion of the spent fuel, reprocessing greatly reduces the mass of the waste. The reprocessed
+    waste can be turned into liquid form, mixed with frit (the substance from which glass is made) and then
+    vitrified (made into glass). The vitrification process immobilizes the radioactive particles and makes
+    the waste much more safe and manageable, but vitrification can only be applied to reprocessed spent fuel.
+
+    It is worthwhile to note that the PUREX processed was designed specifically to seperate plutonium for
+    bombs. Therefore, literally any other reprocessing approach would be less prone to proliferation.
+    Researchers are actively investigating alternatives to the PUREX technology. For example, the UREX+ (for
+    uranium extraction) technique extracts the uranium but leaves the plutonium mixed with other
+    radioactive materials. Another popular alternative, pyrochemical reprocessing, inserts giant electrodes
+    into the chemical bath of nuclear waste and electroplates the plutonium and other transuranics onto
+    the electrode. However, none of these alternatives have been put into practice due to the high construction
+    and operating costs of reprocessing facilities.\cite{aa12}
+
+    \red{fast neutron reactors}
+
+    \subsection{Dry cask storage}
+    Spent fuel pools have very limited capacities. 
+    Overground dry cask storage provides a low-cost temporary storage before the spent fuel is disposed of.
+    After cooling in the spent fuel pools for a few
+    years, the spent fuel can be stored in dry casks filled with inert gas. 
+    The dry casks are often metal cyliders welded to prevent leakage. Each cask is surrounded by
+    additional steel and concrete to provide shielding.
+    The heat from the cask is removed by passive
+    cooling (air circulation). The Nuclear Regulatory Comission is hoping that the casks will
+    last safely for as long as 100 years, yet corrosions and cracking occur in 30 years
+    or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
+    overground dry cask storage does not seem to be a long-term solution. The doubts were expressed
+    dudring  the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
+    \begin{displayquote}
+    Leaving the spent fuel onsite for extended periods of time was never intended and is not
+    responsible. (Dry cask manufacturer)
+
+    It is not ethical, basically to plan for long-term storage without pursuing a well defined repository
+    program. (Department of Energy employee speaking as a citizen)
+    \end{displayquote}
+    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in pools
+    that are close or over to their designed limits.\cite{a11,aa12}
+    Therefore, there is clearly a need for a long-term disposal plan, which we will epxplore in
+    the next section.\\
+    \red{(A lot more drama with the NRC going on here, loss of credibility.)}
 
 \section{Disposal of HLW}
-    \subsection{Comments on partitioning and transmutation}
+    The timeframe of nuclear waste disposal is about $10^4-10^6$ years, after which the radioactivity
+    of the waste decreases down to a safe level.
+    Deep geological repositories that are stable over timescales naturally offer a good option for disposal.
+    It is accepted that geological repositories in a stable environment over 300 meters below ground
+    is safter than indefinite overground storage of HLW, reprocessed or not.\cite{fmr11}
+
+    Reversible repositories hm.
+%    \subsection{Comments on partitioning and transmutation}
+    \subsection{Requirements for HLW disposal}
      \subsection{Survey of possibilities}
-    \subsubsection{mention ocean disposal and why it was banned('Ndrangheta)}
-    \subsection{Necessity of geological disposal sites}
-    \subsection{Requirements of disposal from physics}
-
+     transmutation, ocean floor, deep geological disposal
+%    \subsubsection{mention ocean disposal and why it was banned('Ndrangheta)}
+%    \subsection{Necessity of geological disposal sites}
+%    \subsection{Requirements of disposal from physics}
+     \subsection{Deep borehole disposal}
+     \subsection{Deep geological repository}
+%
 \section{Deep Geological Repository}
+    Therefore, long-term storage before final disposal is crucial. 
     \subsection{something general}

### 2017-01-23 paper.tex
intro done

@@
 %\documentclass{article}
-\documentclass[preprint,aip,pra]{revtex4-1}
+\documentclass[nofootinbib,preprint,aip,pra]{revtex4-1}
 \usepackage{hyperref}
@@
 \section{Introduction}
-IAEA nuclear waste estimation: over 1000 tons of highly radioactive waste.\cite{iaea08}
-Nuclear power estimate:increase by 40\% by 2020. \cite{iaea12}
-meep. Mostly focus on nuclear waste from reactors.
-Global policy review \cite{r12}
+Nuclear power continues to play an important role in global energy supply.
+The International Atomic Energy Agency (IAEA) estimated that the power generated by nuclear power
+is continuing to grow and will increase by $40\%$ by 2030.\cite{iaea12}
+Whereas the ``front end'' of the nuclear fuel cycle, for example reactor physics and power output, draws
+the most attention from the public, the ``back end'' of the fuel cycle, for example waste management and
+disposal, deserves equal, if not more, attention.
+Currently, the IAEA estimates that there are over 10,000 metric tons of highly radioactive nuclear waste 
+generated worldwide annually,
+most of which is spent nuclear fuel from nuclear reactors.\cite{iaea08, r12}
+The world will have over $373,800$ metric tons of highly radioactive waste stock by the end of 2020.\cite{r12}
+Highly radioactive waste remains hazardous for up to $10^6$ years after it is produced and thus requires extremely
+careful 
+Nevertheless, there is currently no permanent repository in the world to dispose of these highly radioactive
+waste.~\footnote{The Waste Isolation Pilot Plant (WIPP) in New Mexico is designed to be a permanent radioactive
+    waste repository, but only
+for waste classified as Transuranic, which, by the Nuclear Regulatory Commission standard, does not have particularly
+high radioactivity but only long half-life (See Sec.~\ref{sec:tru}). In particular, WIPP does not accept spent nuclear
+fuel.}  More than $75\%$ of the spent nuclear fuel worldwide is stored onsite at the nuclear power plants in
+water pools that are close to their limiting capacity.\cite{aa12}
 
-Where they come from:\\
-mop
+Two weeks after the davastating earthquake and tsunami hit the Fukushima Daiichi Nuclear Power Plant in Japan,
+the water pool cooling system was cut off from power and officials feared that
+the highly radioactive waste in the pools would catch fire and send radioactive smoke across a much larger part of Japan, including Tokyo.
+The Nuclear Regulatory Comission (NRC), the independent government agency of the U.S. government overseeing 
+nuclear energy safety,
+estimates that a major fire in a U.S. spent fuel pool is expected to
+displace 3.4 million people in an area larger than New Jersy.\cite{s16}
 
-what to do with them:\\
-meep
-
-\begin{itemize}
-    \item history
-    \item nuclear decay mechanism: weak force and tunneling, radioactive decay
-    \item reactors?
-    \item What is nuclear waste and why is it harmful?
-    \item partition and transmutation
-    \item Above-ground disposal: look at sites, dry-casks storage
-    \item geological disposal ($\star$), Yucca mountain as an example
-    \item ocean floor disposal which was done by everyone but not legal any more
-    \item transmutation
-    \item todo: find sources on the specific health impact of nuclear waste instead of radiation
-        in general
-\end{itemize}
+As we can see, nuclear waste management and disposal is becoming a increasingly urgent matter globally.
+In light of the high stake,
+the uncertainty of the stability of a geological repository over a extremely long timescale has evolved into 
+a topic not only of scientific contention but also of policy debate.
+How well is the predicative power of science? How much trust should we put into the uncertainty estimates
+(i.e. how well can science predict its predicative power)?
+In this work, we will attempt to examine some of the scientific issues underlying nuclear waste management
+and disposal. We will start by building up some
+language and knowledge necessary to understand how nuclear waste is generated (Section~\ref{sec:phys}),
+then apply those principles to study
+what physical properties nuclear waste possesses (Section~\ref{sec:waste}), its management and temporary storage
+(Section~\ref{sec:temp}), and its disposal (Section~\ref{sec:disposal}).
 
 \section{Radioactivity and nuclear fission}
+\label{sec:phys}
     Here we detail the nuclear physics underlying most of the things we will talk about
@@
 \section{Nuclear waste: classification and properties}
+\label{sec:waste}
     \red{how much of each is generated and what the impact is; industry vs military}\\
@@
     \subsection{Transuranic waste(TRU)}
+    \label{sec:tru}
     Transuranic waste is defined as materials that do not belong to high-level waste and are
@@
     \section{Spent fuel management}
+    \label{sec:temp}
     Spent fuel from nuclear reactors constitutes the major source of HLW. Spen fuel pools at the
@@
     US nuclear power plants generally have an operation cycle of about 2 years. At the end of each operation cycle,
-    one-quater to one-third of its fuel assemblies (spent fuel) are removed. The removed spent fuel spend a minimum
-    of five years in these water pools to cool.
+    one-quater to one-third of its fuel assemblies (spent fuel) are removed.
     Most of the nuclear industry's spent fuel is now stored in spent fuel water pools ``at-reactor (AR)'',
@@
     by a transfer tunnel. 
-    Spent fuel pools  are usually about 15-meter deep water pools, with storage rack at the bottom to hold the fuel assemblies.
+    Spent fuel pools  are usually about 15-meter deep borated water pools, with storage rack at the bottom to hold the fuel assemblies.
     The water absorbs the heat and shields the radiation. Water is the chosen medium because it is cheap and
@@
     Overground dry cask storage provides a low-cost temporary storage before the spent fuel is disposed of.
-    After cooling in the spent fuel pools for a few
-    years, the spent fuel can be stored in dry casks filled with inert gas. 
+    After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks filled with inert gas. 
     The dry casks are often metal cyliders welded to prevent leakage. Each cask is surrounded by
@@
     \end{displayquote}
-    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in pools
-    that are close or over to their designed limits.\cite{a11,aa12}
+    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in densely
+    packed pools
+    that are close to or over their designed limits.\cite{a11,aa12}
     Therefore, there is clearly a need for a long-term disposal plan, which we will epxplore in
@@
 \section{Disposal of HLW}
+\label{sec:disposal}
     The timeframe of nuclear waste disposal is about $10^4-10^6$ years, after which the radioactivity
-    of the waste decreases down to a safe level.
+    of the waste is predicted to decline to a safe level.
     Deep geological repositories that are stable over timescales naturally offer a good option for disposal.
@@
     is safter than indefinite overground storage of HLW, reprocessed or not.\cite{fmr11}
-
-    Reversible repositories hm.
-%    \subsection{Comments on partitioning and transmutation}
+    In this section, we first state the requirements for a HLW geological disposal solution, and then
+    we look at  deep geological repository, and its alternatives (or variations): deep boreholes storage
+    and reversible repository.
     \subsection{Requirements for HLW disposal}
@@
     transmutation, ocean floor, deep geological disposal
-%    \subsubsection{mention ocean disposal and why it was banned('Ndrangheta)}
-%    \subsection{Necessity of geological disposal sites}
-%    \subsection{Requirements of disposal from physics}
     \subsection{Deep borehole disposal}
-     \subsection{Deep geological repository}
-%
-\section{Deep Geological Repository}
-    Therefore, long-term storage before final disposal is crucial. 
-    \subsection{something general}
-    \subsection{Yucca Mountain}
-    \subsection{Something related to the French Ciegeo project: licensed reversible disposal site}
-    \subsection{Risk calcs: geophysics, mention WIPP}
+        \subsection{Deep geological repository: Yucca Mountain}
+        \subsubsection{Risk}
+    \subsection{Reversible disposal sites}
 

### 2017-01-23 paper.tex
spell checked

@@
 is continuing to grow and will increase by $40\%$ by 2030.\cite{iaea12}
-Whereas the ``front end'' of the nuclear fuel cycle, for example reactor physics and power output, draws
-the most attention from the public, the ``back end'' of the fuel cycle, for example waste management and
+Whereas the ``front end'' of the nuclear fuel cycle, such as reactor physics and power output, draws
+the most attention from the public, the ``back end'' of the fuel cycle, like waste management and
 disposal, deserves equal, if not more, attention.
@@
 most of which is spent nuclear fuel from nuclear reactors.\cite{iaea08, r12}
-The world will have over $373,800$ metric tons of highly radioactive waste stock by the end of 2020.\cite{r12}
-Highly radioactive waste remains hazardous for up to $10^6$ years after it is produced and thus requires extremely
-careful 
+The world will have over $457,000$ metric tons of highly radioactive waste stock by the end of 2020.\cite{r12}
+Highly radioactive waste remains hazardous for up to $10^6$ years after it is produced and thus
+requires extremely careful disposal methods.
 Nevertheless, there is currently no permanent repository in the world to dispose of these highly radioactive
@@
 high radioactivity but only long half-life (See Sec.~\ref{sec:tru}). In particular, WIPP does not accept spent nuclear
-fuel.}  More than $75\%$ of the spent nuclear fuel worldwide is stored onsite at the nuclear power plants in
-water pools that are close to their limiting capacity.\cite{aa12}
+fuel.}
 
-Two weeks after the davastating earthquake and tsunami hit the Fukushima Daiichi Nuclear Power Plant in Japan,
-the water pool cooling system was cut off from power and officials feared that
-the highly radioactive waste in the pools would catch fire and send radioactive smoke across a much larger part of Japan, including Tokyo.
-The Nuclear Regulatory Comission (NRC), the independent government agency of the U.S. government overseeing 
-nuclear energy safety,
+It is alarming that more than $75\%$ of the spent nuclear fuel worldwide is stored on-site at the nuclear power plants
+in water pools that have exceeded their designed capacity and are reaching their limiting capacity.\cite{aa12}
+Two weeks after the devastating earthquake and tsunami hit the Fukushima Daiichi Nuclear Power Plant in Japan,
+the water pool cooling system was cut off from power. Officials feared that
+the highly radioactive waste in the pools would catch fire and send radioactive smoke
+across a much larger part of Japan, including Tokyo. For some ``fortuitous'' reasons, the fear did not come true.
+However, the Nuclear Regulatory Commission (NRC), the independent government agency
+of the U.S. government overseeing nuclear energy safety,
 estimates that a major fire in a U.S. spent fuel pool is expected to
-displace 3.4 million people in an area larger than New Jersy.\cite{s16}
+displace 3.4 million people in an area larger than New Jersey.\cite{s16}
 
@@
 (i.e. how well can science predict its predicative power)?
+
 In this work, we will attempt to examine some of the scientific issues underlying nuclear waste management
@@
 what physical properties nuclear waste possesses (Section~\ref{sec:waste}), its management and temporary storage
-(Section~\ref{sec:temp}), and its disposal (Section~\ref{sec:disposal}).
+(Section~\ref{sec:temp}), and the debates around its eventual disposal (Section~\ref{sec:disposal}).
 
@@
     \subsection{Radioactivity}
-    \subsubsection{Nuclear shell model vs liquid drop model(?)}
-        By solving Schr\"{o}dinger's equation\cite{k88}
-    \subsubsection{Binding Energy(?)}
-
+    %\subsubsection{Nuclear shell model vs liquid drop model(?)}
+    %\subsubsection{Binding Energy(?)}
         \subsubsection{Quantum theory of radioactive decays}
@@
         \end{equation}
+        In simple terms, $\rho(E_f)dE$ gives the proportion of the particles with energy in the
+        really small interval between
+        $E_f$ and $(E_f+dE)$.
         Note that as in perturbation theory, $\ket{\psi_i}$ and
@@
         the decay probability is small, the result from perturbation theory should hold. Since
-        there is no time dependency in equation~\ref{eq:trans},the
+        there is no time dependency in Eq.~\ref{eq:trans},the
         decay rate should be independent of time, which is the fundamental assumption made in
-        equation~\ref{eq:exp}.
+        Eq.~\ref{eq:exp}.
 
         \subsubsection{Quantifying radioactivity}
-        The radioactivity of a sample $\mathcal{A}$ (defined in equation~\ref{eq:act}) 
+        The radioactivity of a sample $\mathcal{A}$ (defined in Eq.~\ref{eq:act}) 
         is the number of decays of the sample per unit time. The SI
@@
         Sometimes the nucleus can decay by $\gamma$ emission in capture reaction. \red{(?)}
-        But most of the time, the excitation energy deforms the nucleus (see FIG. \ref{fig:fission}).
+        But most of the time, the excitation energy deforms the nucleus (see Fig. \ref{fig:fission}).
         When the
@@
     The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the activity
-    content and the half-life of the radionuclides (see FIG. \ref{fig:scheme}).
+    content and the half-life of the radionuclides (see Fig. \ref{fig:scheme}).
     Activity content is a measure of how much radioactivity
@@
     \label{sec:temp}
-    Spent fuel from nuclear reactors constitutes the major source of HLW. Spen fuel pools at the
-    reactors hold spen fuel cores to cool them down and isolate them for a few years. After that,
+    Spent fuel from nuclear reactors constitutes the major source of HLW. Spent fuel pools at the
+    reactors hold spent fuel cores to cool them down and isolate them for a few years. After that,
     reprocessing can occur to extract the most radioactive plutonium and uranium isotopes and
-    convert the waste into more managable form (vitrification). However, reprocessing of commercial
-    nuclear spent fueld did not
+    convert the waste into more manageable form (vitrification). However, reprocessing of commercial
+    nuclear spent fuel did not
     happen in this country for various historical reasons. 
@@
     by the 5-year mark.\cite{aa12} Therefore, short-term cooling of spent fuel is crucial before
-    futher disposal or reprocessing can be done.
+    further disposal or reprocessing can be done.
     US nuclear power plants generally have an operation cycle of about 2 years. At the end of each operation cycle,
-    one-quater to one-third of its fuel assemblies (spent fuel) are removed.
+    one-quarter to one-third of its fuel assemblies (spent fuel) are removed.
     Most of the nuclear industry's spent fuel is now stored in spent fuel water pools ``at-reactor (AR)'',
-    which means either within the reactor building or in an ajacent spent fuel building linked to the reactor
+    which means either within the reactor building or in an adjacent spent fuel building linked to the reactor
     AR pools minimize the difficulty of transporting spent fuel.\cite{iaea99}
@@
     The water absorbs the heat and shields the radiation. Water is the chosen medium because it is cheap and
-    can provide coooling by circulation. Moreover, water can serve as a shielding from radiation, because it has a
+    can provide cooling by circulation. Moreover, water can serve as a shielding from radiation, because it has a
     fairly high density and hence forces the radiated particles (especially the neutrons) to interact with
@@
     oxidation state of plutonium to one that is insoluble
-    and thus seperable; and finally (c) extracting the plutonium from the mixture. In this way
+    and thus separable; and finally (c) extracting the plutonium from the mixture. In this way
     the oxidized plutonium and uranium are separated and can be reduced back to reusable radionuclides.\cite{lb01} 
@@
 
-    It is worthwhile to note that the PUREX processed was designed specifically to seperate plutonium for
+    It is worthwhile to note that the PUREX processed was designed specifically to separate plutonium for
     bombs. Therefore, literally any other reprocessing approach would be less prone to proliferation.
@@
     radioactive materials. Another popular alternative, pyrochemical reprocessing, inserts giant electrodes
-    into the chemical bath of nuclear waste and electroplates the plutonium and other transuranics onto
+    into the chemical bath of nuclear waste and electroplates the plutonium and other transuranic elements onto
     the electrode. However, none of these alternatives have been put into practice due to the high construction
@@
     After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks filled with inert gas. 
-    The dry casks are often metal cyliders welded to prevent leakage. Each cask is surrounded by
+The dry casks are often metal cylinders welded to prevent leakage. Each cask is surrounded by
     additional steel and concrete to provide shielding.
     The heat from the cask is removed by passive
-    cooling (air circulation). The Nuclear Regulatory Comission is hoping that the casks will
+    cooling (air circulation). The NRC is hoping that the casks will
     last safely for as long as 100 years, yet corrosions and cracking occur in 30 years
@@
     overground dry cask storage does not seem to be a long-term solution. The doubts were expressed
-    dudring  the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
+    during the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
     \begin{displayquote}
-    Leaving the spent fuel onsite for extended periods of time was never intended and is not
+    Leaving the spent fuel on-site for extended periods of time was never intended and is not
     responsible. (Dry cask manufacturer)
@@
     that are close to or over their designed limits.\cite{a11,aa12}
-    Therefore, there is clearly a need for a long-term disposal plan, which we will epxplore in
+    Therefore, there is clearly a need for a long-term disposal plan, which we will explore in
     the next section.\\
@@
 \label{sec:disposal}
-    The timeframe of nuclear waste disposal is about $10^4-10^6$ years, after which the radioactivity
+    The time frame of nuclear waste disposal is about $10^4-10^6$ years, after which the radioactivity
     of the waste is predicted to decline to a safe level.
@@
     It is accepted that geological repositories in a stable environment over 300 meters below ground
-    is safter than indefinite overground storage of HLW, reprocessed or not.\cite{fmr11}
+    is safer than indefinite overground storage of HLW, reprocessed or not.\cite{fmr11}
     In this section, we first state the requirements for a HLW geological disposal solution, and then
@@
 \section{Discussions}
+meep
 
-k
 \begin{acknowledgments}

### 2017-01-25 paper.tex
first version complete

@@
 
-\title{Nuclear Wastes: we're kinda doomed}
+\title{Nuclear Waste Management and Disposal}
 
@@
 \begin{abstract}
-    meep
+    As nuclear reactors around the world keep operating with no disposal method available for highly
+    radioactive nuclear waste, how we manage nuclear waste is becoming more important. In this work,
+    we examine a few scientific and technical related to nuclear waste management and disposal.
+    The principles of radioactivity and nuclear fission helps us understand the source and properties
+    of different types of nuclear waste. Focusing on the highly radioactive spent fuel, the rest
+    of this work outlines the advantages and drawbacks of different spent fuel management and disposal
+    methods. We study the methods used for performance assessments for Yucca Mountain Nuclear Waste Repository,
+    especially in the context of volcanic disruption. Finally, we discuss some possible options for 
+    nuclear waste management and disposal after
+    the termination of the Yucca Mountain project.\\
+    Word count: 6592\\
 \end{abstract}
@@
 Nuclear power continues to play an important role in global energy supply.
-The International Atomic Energy Agency (IAEA) estimated that the power generated by nuclear power
+The International Atomic Energy Agency (IAEA) estimates that the power generated by nuclear power
 is continuing to grow and will increase by $40\%$ by 2030.\cite{iaea12}
@@
 the most attention from the public, the ``back end'' of the fuel cycle, like waste management and
-disposal, deserves equal, if not more, attention.
+disposal, deserves equal, if not more attention.
 Currently, the IAEA estimates that there are over 10,000 metric tons of highly radioactive nuclear waste 
@@
 Nevertheless, there is currently no permanent repository in the world to dispose of these highly radioactive
-waste.~\footnote{The Waste Isolation Pilot Plant (WIPP) in New Mexico is designed to be a permanent radioactive
+waste.\footnote{The Waste Isolation Pilot Plant (WIPP) in New Mexico is designed to be a permanent radioactive
     waste repository, but only
@@
 
-As we can see, nuclear waste management and disposal is becoming a increasingly urgent matter globally.
+Evidently, nuclear waste management and disposal is becoming a increasingly urgent matter globally.
 In light of the high stake,
@@
     in this work. We will apply these results in our discussion of nuclear wastes.
+    %\subsection{The nuclear shell model}
+    %The nuclear shell model provides an adequate framework to describe nuclear excitation and
+    %decay. Here we briefly introduce the model and the evidences supporting it.
     \subsection{Radioactivity}
@@
         \subsubsection{Quantum theory of radioactive decays}
-        We have learned that radioactivity is stochastic in nature and radioactive decay follows
+        We have learned that the radioactive decay is stochastic in nature and radioactive decay follows
         an exponential law
-        \begin{equation} \label{eq:exp}
-            \lambda = -\frac{(dN/dt)}{N},
+        \begin{equation}
+            \label{eq:lambda}
+            \lambda = -\frac{(dN(t)/dt)}{N},
         \end{equation}
-        where $\lambda$ is the decay constant, and the right hand side the probability of decay
-        per time for an atom. The fundamental assumption is that the probability of decay is constant,
-        regardless of the age of the atom. This ``memorylessness'' assumption naturally leads to the
-        exponential law of radioactive decay
+        where $N(t)$ is the number of the nuclei at time $t$, and $\lambda$ the decay constant.
+        Equation~\ref{eq:lambda} states that the rate of decay from the higher energy state to
+        lower energy state is independent of time (since $\lambda$ is constant).
+        It corresponds to the fundamental assumption is that the process of radioactive decay is stochastic
+        in nature and the probability of decay for any given nucleus is constant,
+        regardless of the age of the atom.
+
+        Here we outline an quantum mechanics argument for the time-independent decay probability that is perhaps more
+        satisfying.
+        We can approximate the nucleus as a particle situated in a potential well with potential given by
         \begin{equation}
-            N(t) = N_0 e^{-\lambda t},
+            \label{eq:vp}
+            \tilde{V}(t) = V + V_p,
         \end{equation}
-        where $N_0$ gives the number of nuclei at $t=0$.
+        where $V$ is the time-independent nucleus potential that gives rise to eigenstates $\ket{\psi_1}$
+        and $\ket{\psi_2}$ with $\ket{\psi_2}$ having a lower (expectation value of) energy; $V_p$ is a small
+        perturbation that causes the decay.
+        Suppose the particle is in a state given by
+        \begin{equation}
+            \label{eq:state}
+            \ket{\psi(t)} = c_1(t)\ket{\psi_1} + c_2(t)\ket{\psi_2},
+        \end{equation}
+        where $c_1,c_2$ are the coefficients of superposition.
+        A decay from $\ket{\psi_1}$ to $\ket{\psi_2}$ is equivalent to
+        having $|c_2(t)|^2\gg |c_1(t)|^2$ as time increases. Since we know that time-independent potentials
+        only give rise to time-independent states, the fact that Eq.~\ref{eq:state} has time dependence requires
+        our perturbation $V_p$ in Eq.~\ref{eq:vp} to be time dependent.
 
-        Define activity $\mathcal{A}(t)$ to be the rate of decays in a sample
-        \begin{equation} \label{eq:act}
-            \mathcal{A}(t)\equiv \lambda N(t).
-        \end{equation}
-        Here we outline an quantum mechanics argument for an exponential decay law that is perhaps more
-        convincing.
-        We approximate the potential by $V+V_p$ where $V$ is the nuclear potential and $V_p$ is a small
-        time-dependent perturbation that causes the transition between the states. We know that there has
-        to be time dependence in the potential, otherwise for a superposition of eigenstates
-        $\ket{psi}=c_1\ket{\psi_1} + c_2\ket{\psi_2}$, the coefficients $c_1$ and $c_2$ would 
-        have to be constant and
-        persistent decay of one state into the other would not happen.
-        Fermi's Golden Rule \#2 (which is derived in appendix \ref{a:fermi})
-        gives the transition rate
+        Fermi's Golden Rule \#2 (which we can derive from time-dependent
+        perturbation theory)
+        gives the transition probability that is to the first order, independent of time,
         \begin{equation}
@@
         Note that as in perturbation theory, $\ket{\psi_i}$ and
-        $\ket{\psi_f}$ are both eigenstates of the nuclear potential $V$. Therefore, as long as
-        the decay probability is small, the result from perturbation theory should hold. Since
+        $\ket{\psi_f}$ are both eigenstates of the nuclear potential $V$. As long as
+        the decay probability is small (meaning $V_p\ll V$), the result from perturbation theory should hold. Since
         there is no time dependency in Eq.~\ref{eq:trans},the
         decay rate should be independent of time, which is the fundamental assumption made in
-        Eq.~\ref{eq:exp}.
+        Eq.~\ref{eq:lambda}.
+
+        One can easily solve the differential equation given by Eq.~\ref{eq:lambda} to get
+        the law of radioactive decay
+        \begin{equation}
+            N(t) = N_0 e^{-\lambda t},
+        \end{equation}
+        where $N_0$ gives the number of nuclei in the initial state at $t=0$.
 
         \subsubsection{Quantifying radioactivity}
-        The radioactivity of a sample $\mathcal{A}$ (defined in Eq.~\ref{eq:act}) 
-        is the number of decays of the sample per unit time. The SI
+        The radioactivity (or simply activity) of a sample $\mathcal{A}(t)$ is given by
+        \begin{equation}
+            \label{eq:act}
+            \mathcal{A}(t)\equiv \lambda N(t),
+        \end{equation}
+        which is the number of decays occurring in the sample per unit time. The SI
         unit for radioactivity is becquerel (Bq) and it equals one decay per second. Another commonly used
         unit is curie (Ci, with $1\text{ Ci}=3.7\times 10^{10}\text{ Bq}$) and it is defined as
-        the activity of radium-266. A fission product, cesium-137 has an activity of $3.215\times 10^{12}$ Bq
-        per gram. By contrast, the radioactivity of a human body (mainly due to the small amount of 
-        potassium-40 and carbon-14) is about $8000$ Bq.
+        the activity of radium-266. A commonly found isotope in highly radioactive waste,
+        cesium-137 has an activity of $3.215\times 10^{12}$ Bq
+        per gram. By contrast, the radioactivity of an adult human body (mainly due to the small amount of 
+        potassium-40 and carbon-14), regardless of the weight of the human, is about $8000$ Bq.
+
+        \subsubsection{Different kinds of radiation and shielding}
+        Here we review different types of radiation as well as how shielding works. In general, shielding
+        of radiation forces the energetic radiated particles (electrons, alpha particles and gamma rays) to
+        interact with the shielding material such that the radiation loses much of its energy before it
+        reaches a living tissue. It is worth noting that water can serve as a decent shielding due to its
+        relatively high density.
+
+        \paragraph{Alpha decay}
+        An alpha particle is a helium-4 atom ${}^4_2$He. In an alpha decay, a nucleus emits an alpha
+        particle and decays into a different element. The emission of an alpha particle is
+        particularly favorable because it does not change the parity of the nucleus and has a particularly
+        high binding energy. The process can be represented by
+        \begin{equation}
+            {}^A_Z X_N \rightarrow {}^{A-4}_{Z-2}X'_{N-2} + \alpha.
+        \end{equation}
+        Since the alpha particle is charged and massive, it is the easiest radiation to shield.
+        alpha particles can be stopped after traveling through 2 millimeters of water. The human skin
+        also provides adequate shielding against alpha particles. However, our internal organs are
+        vulnerable against alpha particles. Therefore, if the alpha emitter is inhaled or otherwise enters
+        one's body, it may cause more harm than other types of radiations because it is more easily stopped
+        by the tissues and hence deposits more energy.
+
+        \paragraph{Beta decay}
+        There are a variety of beta decay processes. Here we only describe the negative beta decay that is
+        relevant in this work. The basic decay process usually comes from the decay of a neutron
+        \begin{equation}
+            n \rightarrow p + e^- + \bar{\nu}_e,
+        \end{equation}
+        where $n$ denote  a neutron, $p$ a proton, $e^-$ an electron and $\bar{\nu}$ an electron antineutrino.
+        This can also happen inside a nucleus (bound neutron), represented by
+        \begin{equation}
+            {}^A_Z X_N \rightarrow {}^A_{Z+1} X'_{N+1} + e^- + \bar{\nu}_e.
+        \end{equation}
+        In a beta decay, the electron emitted carries the most energy. Over 5 centimeter of water usually provides
+        good shielding against beta radiation. Metals are extremely effective for shielding beta radiation.
+
+        \paragraph{Gamma decay}
+        Most nuclear reactions, including the alpha and beta decay, leave the resulting nucleus in an excited state.
+        These excited states quickly decay to the ground state via emitting gamma rays with energies in the range
+        of $0.1$ to $10$ MeV, which are characteristic of the energy differences between nuclear states.
+        This process is very much identical to its atomic counterpart. As high energy massless particles, gamma
+        radiation is the hardest to shield, and the thickness of the shield depends on the energy of the photons.
+        Usually it takes 5 meters of water, or about 2 meters of concrete,  to reduce the energy of the photon
+        to one billionth of its original energy. Metal is usually more effective against gamma radiation since
+        gamma radiation is a form of electromagnetic wave.\cite{l01}
 
@@
         changes that lead to cancer and genetic defects. Here we briefly summarize the known effects.
-        \subsubsection{Different kinds of radiation and shielding}
-        meep
         \subsubsection{Quantifying radiation on living tissues}
@@
         of these quantities: the absorbed dose, the equivalent dose and the effective dose.
+        The SI units for all of these quantities are equivalent to J/Kg, the amount of energy deposited into
+        per unit mass of living tissue. However, this unit is called gray (Gy) for absorbed dose but sievert
+        for equivalent dose and effective dose.
 
@@
         \end{equation}
-        The approximate\red{(?)} linear dependence on $E_{\gamma}$ is valid from less than $0.1$ MeV to several
+        The approximate linear dependence on $E_{\gamma}$ is valid from less than $0.1$ MeV to several
         MeV and therefore should be good for all sources but those generated in an accelerator.\cite{my68}
@@
         equivalent dose is called the sievert (Sv) and $1\text{ Sv}=1\text{ J kg}^{-1}$.
-        \\\red{Dose distribution and bio effectiveness?}
         \begin{table}
@@
         \end{table}
+
         Finally, certain organs and parts of the body are more vulnerable to radiation than the rest.
@@
         about 3 to 5 percent per sievert.\cite{unscear12}
-        \red{LNT vs threshold?}
 
@@
 
-        \subsubsection{Limitations of current studies}
-    \subsection{Properties of a neutron}
-        Because neutron plays a central role in the physical processes taking
-        place in a reactor, we should first look at some of its
-        properties.
-        \subsubsection{Free neutron}
-        A free neutron has a half-life of about $10.2$ minutes. \cite{gc01} The Standard
-        Model dictates that a free neutron can only go through beta decay, in which a neutron decays
-        into a proton, an electron, and an electron anti-neutrino \red{(source)}. The decay can be denoted by
-        \begin{equation}
-            n^0 \rightarrow p^+ + e^- + \bar{\nu}_e,
-        \end{equation}
-        where $n$ denotes a neutron, $p^+$ a proton (that is positively charged), $e^-$ a
-        (negatively charged) electron and $\bar{\nu}_e$ an electron antineutrino.
+%        \subsubsection{Limitations of current studies}
+%        \red{TODO}
 
-        \subsubsection{Neutron Absorption}
+    \subsection{Fission}
+    \label{sec:fission}
+        Nuclear Fission is the process in which a heavy nucleus splits into smaller fragments and releases
+        a huge amount of energy. It is the main physical process taking place in a nuclear reactor. It is
+        also the main mechanism that generates the most radiative wastes that we will discuss later
+        in this work. Therefore, we should first familiarize ourselves with fission.
+    \subsubsection{Neutron absorption}
     \label{sec:capture}
-        easy to absorb because neutrally charged
+    Thermal (slow) neutrons induce fissions in a reactor.
+    Since a neutron does not carry any charge, it experiences a much smaller potential when it interacts
+    with a nucleus. In the nuclear absorption reaction, a neutron crosses the potential barrier to form
+    an excited nucleus.
     The nuclear absorption reaction is the most crucial type of reactions that takes place in
@@
     case of a nuclear reactor), neutron absorption is the primary mode of interaction.
-        In the process of neutron absorption, a nucleus absorbs a neutron to form
-        a compound nucleus. Depending on the path that the compound nucleus takes to return to its ground
+    Depending on the path that the compound nucleus takes to return to its ground
     state, neutron absorption is classified into radiative capture and neutron-induced
@@
 
-        In a radiative capture process, the compound nucleus emits photons, whereas in neutron-induced
+    In a radiative capture process, the compound nucleus emits photons in the form of gamma rays, whereas in neutron-induced
     fission, the compound nucleus splits into fragments.
@@
     \end{equation}
-        Moreover, the isotope ${}^{198}$Au can spontaneously decay into ${}^{198}$Hg via beta decay. This process can
-        be seen as the analogy of free neutron beta decay for bound neutrons.
-
-        i.e. how does a neutron excite a nucleus?
-        \red{fission vs radiative capture condition?}
-        \red{cross section?}
-        \red{pic of the two potentials}
-
-    \subsection{Fission}
-    \label{sec:fission}
-        Nuclear Fission is the process in which a heavy nucleus splits into smaller fragments and releases
-        a huge amount of energy. It is the main physical process taking place in a nuclear reactor. It is
-        also the main mechanism that generates the most radiative wastes that we will discuss later
-        in this work. Therefore, we should first familiarize ourselves with fission.
-        \red{fissile vs fissionable}
-
+    Moreover, the isotope ${}^{198}$Au can spontaneously decay into ${}^{198}$Hg via beta decay. This is not
+    uncommon, as the nucleus formed by neutron capture is usually unstable.
         \subsubsection{Fission and fission products}
@@
         The neutron capture produces a compound nucleus ${}^{236}U$ that is in an excited state.
-        Sometimes the nucleus can decay by $\gamma$ emission in capture reaction. \red{(?)}
+        Sometimes the nucleus can decay by $\gamma$ emission in capture reaction.
         But most of the time, the excitation energy deforms the nucleus (see Fig. \ref{fig:fission}).
@@
         A typical reaction is described by
-        \[{}^{235}_{92}\text{U} + n \rightarrow {}^{92}_{36}\text{Kr} + {}^{142}_{56}\text{Ba} + 2 n.\]
+        \begin{equation}
+        {}^{235}_{92}\text{U} + n \rightarrow {}^{92}_{36}\text{Kr} + {}^{142}_{56}\text{Ba} + 2 n.
+        \end{equation}
         The energy liberated is about $180$ MeV. The combination of fission products is not
@@
         Very often, photons, electrons and neutrinos are also emitted through beta and gamma decay.
-    \red{(insert 10.2 from Lilley here? distribution of fission fragment masses)}
-        Radioactive if neutron-to-proton ratio is over 5. On average, $2.5$ neutrons are produced per fission.\cite{l01} 
+        On average, $2.5$ neutrons are produced per fission.\cite{l01} 
         \begin{figure}[h]
@@
 
-        \red{Wong has more on the fragmentation thing in his fission chapter.}
-
         \subsubsection{Fission energy budget}
         We can estimate the amount of energy released in fission by considering the binding energy difference
-        and the properties of the product.
+        and the properties of the product. Essentially, any energy carried by particles that can be slowed down 
+        in the reactor can be harvested as heat to generate power.
+
         The binding energy per nucleon is about $7.6$ MeV u${}^{-1}$ for uranium and about $8.5$ MeV u${}^{-1}$
@@
         about 2.5 million times the energy produced by burning coal of the same mass.\cite{e17}
-        \subsubsection{Chain reaction??}
+%        \subsubsection{Chain reaction??}
 
@@
 \label{sec:waste}
-    \red{how much of each is generated and what the impact is; industry vs military}\\
     Different organizations have slightly different classification schemes for nuclear waste.
@@
     The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the activity
-    content and the half-life of the radionuclides (see Fig. \ref{fig:scheme}).
-    Activity content is a measure of how much radioactivity
-    a given amount of waste contains: a form of waste may possesses a high activity content due to
-    \red{a high concentration of radionuclides or the presence of radionuclides with high activity.}\cite{iaea09}
+    content and the half-life of the radionuclides.
+    Activity content is a measure of how much radioactivity $\mathcal{A}(t)$, as defined in Eq.~\ref{eq:act},
+    a given amount of waste contains.
+    A form of waste may possesses a high activity content due to
+    a high concentration of radionuclides or the presence of radionuclides with high activity.\cite{iaea09}
     Each form of waste, depending on the half-life, activity content, and temperature, requires different
     disposal methods.
-    \begin{figure}[h]
-        \centering
-        \includegraphics[width=0.7\textwidth]{wastescheme.png}
-        \caption{An illustration of the IAEA classification scheme for nuclear waste along with suggested
-        disposal methods.\cite{iaea09}}
-        \label{fig:scheme}
-    \end{figure}
+
     We will focus on high-level waste (HLW, defined below) in the rest of this work.
     In this section, we will first examine the composition and properties of HLW and then
-    comment on other categories of waste and disposal methods.
-
+    comment on other categories of waste and disposal methods (following the NRC classification scheme). 
     \subsection{High-level waste(HLW)}
-    \red{reprocessing or not}\\
-    \red{where does the heat come from?}
     High-level Wastes (HLW) has high activity contents and usually generates a large amount of
@@
     fission of uranium-235 produces lighter radionuclides like cesium-137 and strontium-90, which
-    account for most of the heat and penetrating radiation (\red{penetrating?}). In addition to
+    account for most of the heat and penetrating radiation (gamma radiations). In addition to
     the fission process, some uranium-235 atoms go through radiative capture
-    (see section~\ref{sec:capture}) and produce elements that are
+    (see section~\ref{sec:capture}) and produce radionuclides that are
     heavier than uranium (``transuranic'') such as plutonium. The transuranic atoms do not generate
-    as much heat or radiation as the fission products, but they tend to have a much longer half-life.
+    as much heat or radiation from their subsequent decay as the fission products, but they tend to have
+    a much longer half-life.
     For example, strontium-90 and cesium-137 have half-lives of about $30$ years, whereas
-    plutonium-239 has a half-life of $24,000$ years.
+    the transuranic plutonium-239 has a half-life of $24,000$ years.
 
-    \subsection{Transuranic waste(TRU)}
+    \subsection{Transuranic waste (TRU)}
     \label{sec:tru}
@@
     contaminated by alpha-emitting radionuclides with
-    sufficiently long half-life ($<20$ years) of elements with atomic number $92$ (atomic number of uranium)
-    or larger and concentration of activity larger than $3700$ Bq/g.\cite{j83,s01}. The activity
+    sufficiently long half-life ($>20$ years) of elements with atomic number $92$ (atomic number of uranium)
+    or larger and concentration of activity larger than $3700$ Bq/g.\cite{j83,s01} The activity
     of TRU is usually low but it remains active for long period of time.
-    \red{WIPP is only for TRU}
 
@@
     \subsection{Spent fuel pools}
-    \red{boron-10 and radiation shielding.}
     The heat output of the spent fuel decreases by $99\%$ after the first year and by another factor of $5$
@@
     Spent fuel pools  are usually about 15-meter deep borated water pools, with storage rack at the bottom to hold the fuel assemblies.
+    The borated water contains boron which are more prone to neutron captures.
     The water absorbs the heat and shields the radiation. Water is the chosen medium because it is cheap and
@@
     fairly high density and hence forces the radiated particles (especially the neutrons) to interact with
-    the water molecules and lose energy \red{(more precise terms needed, cladding materials)}. Water temperature is maintained
+    the water molecules and lose energy. Water temperature is maintained
     below 50 $^\circ$C via circulation and cooling. The integrity of the storage can persist well beyond 
@@
     by fission. Un-reprocessed fuel usually takes 10 times as much time to reach the safe
-    activity level (\sim$10^5$ years) as reprocessed fuel (\sim$10^4$ years). Moreover, since fission products
+    activity level ($\sim 10^5$ years) as reprocessed fuel ($\sim 10^4$ years). Moreover, since fission products
     make up a small portion of the spent fuel, reprocessing greatly reduces the mass of the waste. The reprocessed
@@
 
-    \red{fast neutron reactors}
+    %\red{fast neutron reactors}
 
@@
     cooling (air circulation). The NRC is hoping that the casks will
-    last safely for as long as 100 years, yet corrosions and cracking occur in 30 years
+    last safely for as long as 100 years, yet it has been observed that corrosions and cracking occur in 30 years
     or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
@@
     packed pools
-    that are close to or over their designed limits.\cite{a11,aa12}
+    that are close to or over their limiting capacity.\cite{a11,aa12}
     Therefore, there is clearly a need for a long-term disposal plan, which we will explore in
-    the next section.\\
-    \red{(A lot more drama with the NRC going on here, loss of credibility.)}
+    the next section.
 
@@
     of the waste is predicted to decline to a safe level.
-    Deep geological repositories that are stable over timescales naturally offer a good option for disposal.
-    It is accepted that geological repositories in a stable environment over 300 meters below ground
+    Deep geological repositories that are stable over geological timescales naturally offer a good option for disposal.
+    In fact, it is accepted by the international science community that geological repositories in a stable environment over 300 meters below ground
     is safer than indefinite overground storage of HLW, reprocessed or not.\cite{fmr11}
-    In this section, we first state the requirements for a HLW geological disposal solution, and then
-    we look at  deep geological repository, and its alternatives (or variations): deep boreholes storage
-    and reversible repository.
-    \subsection{Requirements for HLW disposal}
-    \subsection{Survey of possibilities}
-    transmutation, ocean floor, deep geological disposal
+    In this section, we first use the Yucca Mountain Nuclear Waste
+    Repository as a case study to understand the requirements for a permanent HLW disposal site and the debate
+    surrounding it.
+    Then we look at an alternative to a permanent deep geological solution:
+    deep boreholes storage.
+    \subsection{Deep geological repository: Yucca Mountain}
+        The defining characteristic for a nuclear waste disposal site is its ability to minimize human
+        exposure to radiation over the next ten thousand to a million years or more.
+        For a deep geological repository, this means preventing the release
+        of radionuclides to the ground water or the surface.
+        Due to the amount of radiation the repository would contain, it would be difficult to perform
+        maintenance once the waste was deposited. Therefore, scientists and engineers are challenged
+        to predict the geological environment of the Yucca Mountain over a million years and come up
+        with a design that they will have very little chance to alter once it is implemented.
+
+        The Yucca Mountain Nuclear Waste Repository was designated by the Nuclear Waste Policy Act
+        of 1987 to be the nation's permanent deep underground repository for spent nuclear fuel. It is
+        located near the Nevada Test Site in Nevada. The project has been highly contested since its
+        inception and the Obama administration ended federal funding for the project in 2011 for political,
+        not safety or technical, reasons. We will take a glimpse at the suitability assessment of the site
+        as a nuclear waste repository to get a sense of its complexity.
+        \subsubsection{Geological stability of Yucca Mountain}
+        The chief concern of the suitability of Yucca Mountain was about water. If moist air comes to
+        contact with the waste, the waste may get oxidized and become chemically and physically unstable, thereby
+        releasing radionuclides.
+        The community seemed to agree that the cladding will have to carry the primary
+        responsibility of preventing the oxidation of the waste, but predicting the performance of some
+        engineered product over the timescale of a million years is not something with which we have had much 
+        experience. Over time, the radiation from the waste will also alter the rock (by altering its composition
+        and creating fractures) and the water in the rock.\cite{m06}
+
+        Many other factors are considered. Climate change will influence precipitation and thus the
+        rate at which water percolates the rock and reaches the waste. 
+        Tectonic motion may change how much the rocks are fractured and thus the percolation rate.
+        Volcanic activity, meteorite hit and earthquakes may seriously undermine the integrity of the repository.
+        Hot upwelling water from below the repository may corrode the cladding and release radionuclides into
+        the ground water system. It is impossible to address all of these concerns in this work, but we will look
+        at how the risk assessment is carried out in the case of volcanism.
+        \subsubsection{Total System Performance Assessment (TSPA)}
+        The Department of Energy (DOE) created Total System Performance Assessment (TSPA) as a huge probabilistic
+        computer model that predicts the potential annual dose to the potential receptor.
+        Essentially, the TSPA takes a huge set of parameters characterizing features that DOE deems as important,
+        carry out the geophysical and geochemical simulations,
+        and outputs the expected annual dose leaked from the repository.
+        DOE expects that
+        uncertainty is fully integrated into the analysis done by TSPA such that even if the model itself change
+        (for example if new parameters and factors are included), the results from TSPA will not change
+        significantly.\cite{me06,ocrwm02}
+
+        \subsubsection{Risk and uncertainty estimation: volcanism}
+        \label{sec:volcano}
+        Yucca Mountain is composed of volcanic rocks created from a series of large-volume eruptions between
+        fifteen million and eight million years ago. However, it is the small-volume eruptions, the last of
+        which occurred about $75,000$ years ago, that would pose risk to the repository in its life span.
+        The central question is whether the risk of volcanic eruption is acceptable. Mathematically,
+        according to the Environmental Protection Agency (EPA)'s 1993 guideline,
+        if the probability of a volcanic eruption seriously damaging the repository is less than
+        one in one hundred million per year, it is acceptable.\cite{epa93}
+        Of course, one has to also consider consequences of a potential radiological releases if an eruption
+        occur. Here we focus on the estimation of occurrence probability.
+
+        The probability that a future volcanic eruption disrupting a repository is given by
+        \begin{equation}
+            \label{eq:prob}
+            P(D) = P(E2|E1)P(E1),
+        \end{equation}
+        with the standard probability notation: 
+        $P(A)$ means probability of event $A$ happening;
+        $P(A|B)$ gives the probability of event $A$ happening given that event $B$ happened;
+        $D$ denote the disruption event,
+        $E2$ the event that the eruption affects the repository, $E1$ the event that an eruption occurs.
+        Essentially, we are looking at the probability of a volcano occurring and affecting the repository.
+
+        \begin{figure}[h]
+            \label{fig:volcano}
+            \centering
+            \includegraphics[width=0.9\textwidth]{volcano.png}
+            \caption{A flowchart for the volcanism risk estimation problem. The oval-shaped
+            parts are all probabilistic. See Sec.~\ref{sec:volcano} for details.\cite{me06}}
+        \end{figure}
+
+        The information necessary to estimate the probability is listed in the flowchart Fig.~\ref{fig:volcano}.
+        The boxes represent decision parameters that depends on the judgement of the decision makers and
+        the scientists. For example the ``Volcanic Zone Models'' should be picked by the scientists, and
+        ``Repository Area'' decided by policy makers.
+
+        The ovals are probabilistic parameters in which uncertainty
+        is represented as a probability distribution over a range of possible values.
+        Each oval is a complex computational model.
+        The large bolded ovals are the $E1$ and $E2$ parameters in Eq.\ref{eq:prob}.
+        The $E2$ oval makes
+        use of the events simulated in $E1$ to calculate the conditional probability $P(E2|E1)$ and output
+        the disruption probability $P(D)$ to TSPA, which calculates the expected leaked dose over the
+        life span of the repository. Here we look in some detail at the calculation of those parameters.
+        \paragraph{Monte-Carlo simulation}
+        Estimation of uncertainty is key in this problem, because over the timescale of a million years,
+        a small uncertainty would amount to large measurable change. Recall the traditional method of
+        propagation of uncertainty. Let $\delta_X$ denote the uncertainty in variable $X$. For a function
+        with analytic form $y(x_1,x_2,...x_n)$, the uncertainty in $y$ is given by
+        \begin{equation}
+            \delta_y = \sqrt{\sum\limits_{i=1}^n \left(\left.\frac{\partial y}{\partial x_i}
+            \right\rvert_{x_i} \delta_x\right)^2},
+        \end{equation}
+        which uses a linear approximation of $y(x_1,x_2,...,x_n)$ near the measured values of
+        $(x_1,x_2,...,x_n)$.
+        However, if the uncertainty of the input parameter if expressed as a distribution, or if $y$ is
+        not an analytic function, then the propagation of uncertainty method fails. In our volcano disruption
+        problem, $y$ can be seen as the complex computational model like TSPA. However, there is a
+        straightforward algorithm called Monte-Carlo simulation that can estimate the uncertainty of $y$.
+        The steps of the algorithm is as follows.
+        \begin{itemize}
+            \item Generate a large number of $(x_1,x_2,...x_n)$ samples according to their respective distributions.
+            \item For each $(x_1,x_2,...,x_n)$ sample, calculate a $y$ value (with the computational model).
+            \item Aggregate all the $y$'s generated to estimate its distribution or uncertainty.
+        \end{itemize}
+        This method usually converges reasonably fast -- the error goes down as $1/\sqrt{N}$ where $N$ is the number
+        of samples generated. It is the algorithm of choice in Fig.~\ref{fig:volcano} to
+        propagate the uncertainty information in one oval to another.
+        \paragraph{Recurrence rate} The recurrence rate $E1$ only uses information from the ``Event Count''
+        oval, which estimates the true number of eruptions in the Yucca Mountain region in a given time range
+        from observed data. The choice of time range is not trivial: longer time range certainly gives us
+        more data and hence improved statistics, but longer time interval dated far back may fail to represent
+        future event rates since the geology evolves. The best approach seems to be choosing time interval
+        based on geological cycles.
+
+        Finally, we need the undetected volcanic intrusion rate to establish the recurrence rate. Volcanic
+        intrusions are essentially underground eruptions that can go undetected. The DOE field studies conclude
+        that intrusions are possible but are presumed to be associated with surface eruption in space and time. 
+        By contrast, the NRC argues that intrusions need not be associated with eruption and there
+        may be three to five times as many undetected intrusions as detected surface eruptions. The divergence
+        of the two agencies' viewpoints partly attributes to their different estimated recurrence rate.
+        From different studies adopting different input models, estimated recurrence rate ranges from
+        $1.5\times 10^{-6}$ to $8\times 10^{-6}$ events per year. These estimates translate to an estimated
+        time between events from 125 thousand years to 250 thousand years. We can see that the choices of
+        models introduce a significant uncertainty to the estimation.\cite{me06}
+
+        \paragraph{The probability of repository intersection}
+        Given a probability distribution for $E1$, we can estimate $P(E2|E1)$ -- the probability of
+        a given event intersecting the Yucca Mountain site. It is simply the ratio of the area of
+        the repository to the entire volcanic zone of a given event. Since the area of the repository
+        has not been decided, a probability distribution over a range of areas is used.
+        However, since not enough volcano events have been observed in the region, the model
+        to predict the size of the volcanic zone is not well determined.
+        Different models predict $P(E2|E1)$ ranging from $0.015$ to $0.0008$, which again has a high
+        uncertainty.
+
+        \paragraph{The probability of a disruption} The probability distribution of $E1$ and $E2$
+        are combined through a Monte-Carlo simulation to produce the disruption probability/rate $P(D)$
+        at the Yucca Mountain repository. DOE estimates a mean disruption rate of $1.5\times 10^8$ with
+        a 90 percent confidence interval of $5.4\times 10^{-9}-4.9\times 10^{-8}$. Hence the DOE's result
+        states that most likely the disruption won't happen more than once every 20 million years.
+
+        The NRC emphasizes the importance of the bounding nature of the problem and uses only a subset
+        of the DOE model that reflect conservatism. It concludes a disruption rate of
+        $10^{-7}$ to $10^{-8}$ per year. Furthermore, it argues that it should not distinguish the likelihood
+        of disruption rate of different values within this range. Therefore it picks the upper bound
+        of $10^{-7}$ as the estimated event rate, representing the worst case estimate.
+
+        Both the NRC and the DOE conclude with a very low volcanic disruption rate that is below the EPA
+        threshold. Nevertheless, the consequences of such a disruption event is still not well understood.
+        If the leaked dose is low in such an event, then the risk is negligible. If the leaked dose is high,
+        then the NRC estimate of $10^{-7}$ events per year would necessitate further investigation into such
+        a scenario.\cite{me06}
         \subsection{Deep borehole disposal}
-        \subsection{Deep geological repository: Yucca Mountain}
-        \subsubsection{Risk}
-    \subsection{Reversible disposal sites}
+        Deep borehole disposal utilizes deep and narrow boreholes to dispose of HLW. The boreholes
+        will place the cladded waste as much as 5 kilometers under the surface of the ground and seal it
+        with concrete clay.
+        The engineered clad, along with the granite rock with few cracks on which the borehole is drilled,
+        provides
+        good isolation for the radionuclides in the waste (See Fig.~\ref{fig:borehole}).
+        \begin{figure}[h]
+            \label{fig:borehole}
+            \centering
+            \includegraphics[width=0.8\textwidth]{borehole.jpg}
+            \caption{Illustration of borehole disposal, with deep repositories' depths for comparison.\cite{c15}}
+        \end{figure}
+
+        Deep boreholes have several advantages over deep geological repository. For example, suitable
+        rocks can be more readily found for boreholes (since it does not take up a lot of space and can be
+        of smaller scale),
+        whereas large repository sites are hard to identified. \cite{b09,c15}
+        However, the borehole solution has its own limitations.
+        Drilling a hole as deep as 5 kilometers poses a daunting engineering challenge.
+        Moreover, reactor rods from commercial nuclear plants in the U.S. usually size up to 2 meters in diameter,
+        which would be too large for the borehole and require extensive repackaging.
+        Nevertheless, deep borehole disposal remains a valid option.
+        Sandia lab has embarked on a five-year plan to experiment with the deep borehole design.\cite{c15}
+
+        %\subsection{reversible disposal}
 
 \section{Discussions}
-meep
+Nuclear waste management and disposal certainly is becoming a more and more pressing issue.
+However, given the technical and political complexity involved in assessing the suitability
+of a permanent repository site,\cite{me06} it seems like dry cask storage should be promoted extensively
+before a permanent repository is ready. Large-scale dry cask storage would ease the pressure on the
+spent fuel pool while easing the pressure on DOE and NRC to rush the scientific analyses.
+But still, we have yet to know the durability of dry casks.
 
-\begin{acknowledgments}
-Thanks Joel and peeps. Arjendu and M. V. Ramana.
-\end{acknowledgments}
+In terms of scientific analyses for a geological repository, much has yet to be done.
+On the performance analysis front, we need
+to think about how to model without enough data and how to deal with orders of magnitude uncertainty.
+Perhaps some other analysis methods are required. Building and comparing the performance of 
+small retrievable repositories that store
+nuclear waste for about 100 years may give us insights on the stability of different geological sites.
 
-\pagebreak
+Newer reprocessing methods (as mentioned in Sec.~\ref{sec:reproc}) are being proposed with
+fast neutron reactors as a solution to the proliferation issue with current reprocessing method.
+A fast neutron reactor uses neutrons of higher energy to induce the fission of plutonium-239
+such that the spent nuclear fuel would contain very little plutonium-239. Reprocessing and
+vitrification techniques can significantly reduce the volume of the waste and make it more
+manageable. With reprocessing, deep boreholes storage may become a more feasible disposal solution 
+for commercial reactor waste.
+
+%\begin{acknowledgments}
+%Thanks Joel and peeps. Arjendu and M. V. Ramana.
+%\end{acknowledgments}
+\newpage
 \bibliographystyle{aip}
 \bibliography{refs}
-\appendix
-\section{Derivation of Fermi's Golden Rule \#2}
-\label{a:fermi}
+%\newpage
+%\appendix
+%\section{Derivation of Fermi's Golden Rule \#2}
+%\label{a:fermi}
 \end{document}

### 2017-02-24 paper.tex
ahhhh

@@
 %\documentclass{article}
-\documentclass[nofootinbib,preprint,aip,pra]{revtex4-1}
+\documentclass[nofootinbib,preprint,aps]{revtex4-1}
 \usepackage{hyperref}
@@
 \newcommand{\red}[1]{\textcolor{red}{\bf #1}}
-
 \begin{document}
-
 \title{Nuclear Waste Management and Disposal}
@@
 Nuclear power continues to play an important role in global energy supply.
-The International Atomic Energy Agency (IAEA) estimates that the power generated by nuclear power
-is continuing to grow and will increase by $40\%$ by 2030.\cite{iaea12}
+The International Atomic Energy Agency (IAEA) estimates that nuclear power
+consumption will increase by $40\%$ by 2030.\cite{iaea12}
 Whereas the ``front end'' of the nuclear fuel cycle, such as reactor physics and power output, draws
-the most attention from the public, the ``back end'' of the fuel cycle, like waste management and
+the most attention from the public, the ``back end'' of the fuel cycle, i.e. waste management and
 disposal, deserves equal, if not more attention.
@@
 the highly radioactive waste in the pools would catch fire and send radioactive smoke
-across a much larger part of Japan, including Tokyo. For some ``fortuitous'' reasons, the fear did not come true.
+across a much larger part of Japan, including Tokyo.\cite{s16} For some ``fortuitous'' reasons, the fear did not come true.
 However, the Nuclear Regulatory Commission (NRC), the independent government agency
@@
 How well is the predicative power of science? How much trust should we put into the uncertainty estimates
-(i.e. how well can science predict its predicative power)?
+(i.e. \red{how well can science predict its predicative power})?
 
@@
 then apply those principles to study
-what physical properties nuclear waste possesses (Section~\ref{sec:waste}), its management and temporary storage
+the physical properties of nuclear waste (Section~\ref{sec:waste}), its management and temporary storage
 (Section~\ref{sec:temp}), and the debates around its eventual disposal (Section~\ref{sec:disposal}).
@@
 \label{sec:phys}
-    Here we detail the nuclear physics underlying most of the things we will talk about
+    Here we detail the nuclear physics underlying most of the discussions
     in this work. We will apply these results in our discussion of nuclear wastes.
-    %\subsection{The nuclear shell model}
-    %The nuclear shell model provides an adequate framework to describe nuclear excitation and
-    %decay. Here we briefly introduce the model and the evidences supporting it.
     \subsection{Radioactivity}
+    Something about radioactivity. The derivations and in this section follows chapter $3$ in
+    Krane's text (Ref.~\onlinecite{k88}) and chapter $4$ of Lilley's text (Ref. \onlinecite{l01}).
     %\subsubsection{Nuclear shell model vs liquid drop model(?)}
@@
         to one billionth of its original energy. Metal is usually more effective against gamma radiation since
-        gamma radiation is a form of electromagnetic wave.\cite{l01}
+        gamma radiation is a form of electromagnetic wave.\cite[chap. 1]{l01}
 
@@
         where $D_R$ is the absorbed dose and $w_R$ the weighting factor of a given kind of radiation $R$
-        (see table \ref{tab:eq} for a list of weighing factors). If there are multiple types of radiation 
+        (see Table~\ref{tab:eq} for a list of weighing factors). If there are multiple types of radiation 
         present, the equivalent does is given by the weighted sum over all contributions. The SI unit for
@@
         \begin{table}
-            \label{tab:eq}
             \centering
@@
             \end{tabular}
+            \label{tab:eq}
             \end{ruledtabular}
@@
         sum of the equivalent dose over individual body parts. Some of the recommended values of tissue
-        weighting factors are given in table \ref{tab:eff}. A weighing factor of $1$ is equivalent to the
+        weighting factors are given in Table~\ref{tab:eff}. A weighing factor of $1$ is equivalent to the
         whole body absorbing the radiation uniformly. The SI unit of effective dose is sievert (Sv),
@@
         \begin{table}
-            \label{tab:eff}
             \centering
@@
                 \end{tabular}
+                \label{tab:eff}
             \end{ruledtabular}
@@
         (UNSCEAR), the additional chance of dying of cancer due to radiation exposure above $100$ mSv is
-        about 3 to 5 percent per sievert.\cite{unscear12}
+        about 3 to 5 percent per sievert.\cite{unscear16}
 
@@
         Sometimes the nucleus can decay by $\gamma$ emission in capture reaction.
-        But most of the time, the excitation energy deforms the nucleus (see Fig. \ref{fig:fission}).
+        But most of the time, the excitation energy deforms the nucleus (see Fig.~\ref{fig:fission}).
         When the
@@
         (for example if new parameters and factors are included), the results from TSPA will not change
-        significantly.\cite{me06,ocrwm02}
+        significantly.e\cite{cv14,ocrwm02}
 
@@
         \begin{figure}[h]
-            \label{fig:volcano}
             \centering
@@
             \caption{A flowchart for the volcanism risk estimation problem. The oval-shaped
-            parts are all probabilistic. See Sec.~\ref{sec:volcano} for details.\cite{me06}}
+            parts are all probabilistic. \ref{sec:volcano} \cite{cv14}}
+            \label{fig:volcanos}
         \end{figure}
 
-        The information necessary to estimate the probability is listed in the flowchart Fig.~\ref{fig:volcano}.
+        The information necessary to estimate the probability is listed in the flowchart Fig.~\ref{fig:volcanos}.
         The boxes represent decision parameters that depends on the judgement of the decision makers and
@@
         This method usually converges reasonably fast -- the error goes down as $1/\sqrt{N}$ where $N$ is the number
-        of samples generated. It is the algorithm of choice in Fig.~\ref{fig:volcano} to
+        of samples generated. It is the algorithm of choice in Fig.~\ref{fig:volcanos} to
         propagate the uncertainty information in one oval to another.
@@
         time between events from 125 thousand years to 250 thousand years. We can see that the choices of
-        models introduce a significant uncertainty to the estimation.\cite{me06}
+        models introduce a significant uncertainty to the estimation.\cite{cv14}
 
@@
         then the NRC estimate of $10^{-7}$ events per year would necessitate further investigation into such
-        a scenario.\cite{me06}
+        a scenario.\cite{cv14}
         \subsection{Deep borehole disposal}
@@
         \begin{figure}[h]
-            \label{fig:borehole}
             \centering
@@
             \caption{Illustration of borehole disposal, with deep repositories' depths for comparison.\cite{c15}}
+            \label{fig:borehole}
         \end{figure}
@@
 However, given the technical and political complexity involved in assessing the suitability
-of a permanent repository site,\cite{me06} it seems like dry cask storage should be promoted extensively
+of a permanent repository site,\cite{m14} it seems like dry cask storage should be promoted extensively
 before a permanent repository is ready. Large-scale dry cask storage would ease the pressure on the

### 2017-02-25 paper.tex
daughter nuclei

@@
 \newpage
-
 \section{Introduction}
@@
 
-\section{Radioactivity and nuclear fission}
+\section{Radioactivity and Reactor Physics}
 \label{sec:phys}
@@
     \subsection{Radioactivity}
-    Something about radioactivity. The derivations and in this section follows chapter $3$ in
-    Krane's text (Ref.~\onlinecite{k88}) and chapter $4$ of Lilley's text (Ref. \onlinecite{l01}).
-    %\subsubsection{Nuclear shell model vs liquid drop model(?)}
-    %\subsubsection{Binding Energy(?)}
+    Radioactive is the process in which most of the known detrimental radiations are generated.
+    The derivations and in this section follows chapter $6$ in
+    Krane's text \cite{k88} and chapter $1$ in Lilley's text \cite{l01}.
         \subsubsection{Quantum theory of radioactive decays}
-        We have learned that the radioactive decay is stochastic in nature and radioactive decay follows
-        an exponential law
-        \begin{equation}
-            \label{eq:lambda}
-            \lambda = -\frac{(dN(t)/dt)}{N},
-        \end{equation}
-        where $N(t)$ is the number of the nuclei at time $t$, and $\lambda$ the decay constant.
-        Equation~\ref{eq:lambda} states that the rate of decay from the higher energy state to
-        lower energy state is independent of time (since $\lambda$ is constant).
-        It corresponds to the fundamental assumption is that the process of radioactive decay is stochastic
-        in nature and the probability of decay for any given nucleus is constant,
-        regardless of the age of the atom.
-
-        Here we outline an quantum mechanics argument for the time-independent decay probability that is perhaps more
-        satisfying.
+        The law of radioactive decay is a direct consequence of it being a quantum-mechanical
+        effect.
         We can approximate the nucleus as a particle situated in a potential well with potential given by
@@
         \begin{equation}
-            \rho(E_f) = \frac{dn_f}{dE_f}.
+            \rho(E_f) = \frac{dn_f}{dE_f},
         \end{equation}
+        where $n_f$ is the number density of the state. \cite[chapt. 6]{k88}
         In simple terms, $\rho(E_f)dE$ gives the proportion of the particles with energy in the
@@
         there is no time dependency in Eq.~\ref{eq:trans},the
-        decay rate should be independent of time, which is the fundamental assumption made in
-        Eq.~\ref{eq:lambda}.
+        decay rate should be independent of time, which can be expressed as
+        \begin{equation}
+            \label{eq:lambda}
+            \lambda = -\frac{(dN(t)/dt)}{N},
+        \end{equation}
+        where $N(t)$ is the number of the nuclei at time $t$, and $\lambda$ the decay constant.
+        Equation~\ref{eq:lambda} is the familiar differential equation that describes radioactive decay.
 
-        One can easily solve the differential equation given by Eq.~\ref{eq:lambda} to get
-        the law of radioactive decay
+        We can easily solve the differential equation given by Eq.~\ref{eq:lambda} to get
+        the the number of nuclei as a function of time $t$
         \begin{equation}
+            \label{eq:decay}
             N(t) = N_0 e^{-\lambda t},
@@
 
+        In addition, the mean lifetime $\tau$ is defined as the averaged lifetime of a collection of nuclei.
+        From Eq.~\ref{eq:lambda}, the number of particles between $t$ and $t+dt$ is $dN=\lambda N dt=\lambda
+        N_0 e^{}-\lamba t}dt$. Hence we have
+        \begin{equation}
+            \tau = \frac{\int t dN}{\int dN}=\frac{\int\limits_0^{\infty}t e^{-\lambda t}dt}
+            {\int\limits_0^{\infty}e^{-\lambda t}} = \frac{1}{\lambda}.
+        \end{equation}
+        The half-life, $t_{1/2}$, is the time after which half of the nuclei in the sample are expected to
+        decay. Simply substitute $N(t)=1/2N_0$ in Eq.~\ref{eq:decay} to get
+        \begin{equation}
+            t_{1/2} = \frac{\ln 2}{\lambda}=\tau \ln 2.
+        \end{equation}
+
         \subsubsection{Quantifying radioactivity}
-        The radioactivity (or simply activity) of a sample $\mathcal{A}(t)$ is given by
+        The radioactivity (also called ``activity'') of a sample $\mathcal{A}(t)$ is given by
         \begin{equation}
@@
         unit is curie (Ci, with $1\text{ Ci}=3.7\times 10^{10}\text{ Bq}$) and it is defined as
-        the activity of radium-266. A commonly found isotope in highly radioactive waste,
-        cesium-137 has an activity of $3.215\times 10^{12}$ Bq
-        per gram. By contrast, the radioactivity of an adult human body (mainly due to the small amount of 
-        potassium-40 and carbon-14), regardless of the weight of the human, is about $8000$ Bq.
+        the activity of $1$ gram of radium-266.
 
-        \subsubsection{Different kinds of radiation and shielding}
-        Here we review different types of radiation as well as how shielding works. In general, shielding
-        of radiation forces the energetic radiated particles (electrons, alpha particles and gamma rays) to
-        interact with the shielding material such that the radiation loses much of its energy before it
-        reaches a living tissue. It is worth noting that water can serve as a decent shielding due to its
-        relatively high density.
-
+        Specific activity, on the other hand, gives the activity per mass of a sample
+        \begin{equation}
+        \mathcal{A}_s = \frac{\mathcal{A}(0)}{{m_{sample}}}.
+        \end{equation}
+        Specific activity can characterize the activity level of a certain isotope or compound.
+        The difference between
+        activity and specific activity resembles the difference between heat capacity and specific heat
+        capacity.
+        The radioactive isotope used for Computer Tomography (CT) scan has a specific activity
+        of $\sim 10^{8}$ Bq/g. See Table~\ref{tab:hl} for the half-lives and specific activities
+        of some isotopes present in highly radioactive nuclear waste.
+        \begin{table}
+            \centering
+            \caption{Specific activity of isotopes commonly present in spent nuclear fuel \cite{toi}}
+            \begin{ruledtabular}
+                \begin{tabular}{l l l}
+                Isotope & Half-life (Years) & Specific Activity (Ci/g)\\
+                \hline
+                Cesium-137 & 37 & 88\\
+                Radium-226 & 1.6\times 10^3 & 1\\
+                Plutonium-239 & 2.4\times 10^4 & 0.063\\
+                Uranium-233 & 1.6\times 10^6 & 9.3\times 10^{-3}\\
+                Thorium-232 & 1.4 \times 10^{10} & 1.1\times 10^{-7}\\
+            \end{tabular}
+            \label{tab:hl}
+            \end{ruledtabular}
+        \end{table}
+        \subsubsection{Types of radiation}
+        There are three main types of radioactive decay processes that generate radiations in nuclear waste: 
+        Alpha decay, which generates Alpha particles (helium-4 nuclei), Beta decay which generates Beta
+        particles (electrons) and Gamma decay which generates gamma rays.
+        The type of radiation (particle beam) generated
+        have different properties and deserve separate trearments when we discuss health effects and shielding.
         \paragraph{Alpha decay}
@@
         \end{equation}
-        Since the alpha particle is charged and massive, it is the easiest radiation to shield.
-        alpha particles can be stopped after traveling through 2 millimeters of water. The human skin
-        also provides adequate shielding against alpha particles. However, our internal organs are
-        vulnerable against alpha particles. Therefore, if the alpha emitter is inhaled or otherwise enters
-        one's body, it may cause more harm than other types of radiations because it is more easily stopped
-        by the tissues and hence deposits more energy.
 
@@
         \end{equation}
-        In a beta decay, the electron emitted carries the most energy. Over 5 centimeter of water usually provides
-        good shielding against beta radiation. Metals are extremely effective for shielding beta radiation.
 
         \paragraph{Gamma decay}
-        Most nuclear reactions, including the alpha and beta decay, leave the resulting nucleus in an excited state.
-        These excited states quickly decay to the ground state via emitting gamma rays with energies in the range
-        of $0.1$ to $10$ MeV, which are characteristic of the energy differences between nuclear states.
-        This process is very much identical to its atomic counterpart. As high energy massless particles, gamma
-        radiation is the hardest to shield, and the thickness of the shield depends on the energy of the photons.
-        Usually it takes 5 meters of water, or about 2 meters of concrete,  to reduce the energy of the photon
-        to one billionth of its original energy. Metal is usually more effective against gamma radiation since
-        gamma radiation is a form of electromagnetic wave.\cite[chap. 1]{l01}
+        Most nuclear reactions, including the alpha and beta decay, leave the resulting nucleus in
+        an excited state.
+        We know that excited atom radiates photons to fall back into ground state. Similarly, excited nuclei
+        decay to ground state via emitting photons, with energies in the range
+        of $0.1$ to $10$ MeV, which are characteristic of the energy differences between nuclear energy states.
+
+        \subsubsection{Daughter nuclei and subsequent decay}
+        The resultant nuclei (called ``daughter products'') from a radioactive decay are oftern radioactive
+        themselves and their decay rate can be larger or smaller than their parents. Consider a decay chain
+        $A\rightarrow B \rightarrow C \rightarrow ...$ with decay constants $\lambda_A, \lambda_B,...$. Suppose
+        we start with a sample of $N_0$ $A$ nuclei. The change of the number of nuclei $B$ increases as $A$
+        decays and decreases s $B$ decay. Hence we have
+        \begin{equation}
+            \frac{dN_B}{dt} = -\lambda_B N_B + \lambda_A N_A.
+        \end{equation}
+        Multiply both sides by $e^{\lambda_B t}$ and substitute in $N_A=N_0 e^{-\lambda_A} t$ to get
+        \begin{align*}
+            \frac{dN_B}{dt}e^{\lambda_B t} + \lambda_B N_B e^{\lambda_B t} &= \lambda_A N_A e^{\lambda_B},\nonumber\\
+            \frac{d}{dt} (N_B e^{\lambda_B t}) &= \lambda_A N_0 e^{(\lambda_B-\lambda_A)t}, \nonumber\\
+            N_B e^{\lambda_B t} &= \lambda_A N_0 \int\limits_0^t e^{(\lambda_B - \lambda_A)u} du \nonumber\\
+                                &= \frac{\lambda_A}{\lambda_B-\lambda_A}N_0 e^{(\lambda_B - \lambda_A)t}
+        \end{align*}
+        Hence we end up with the number of $B$ nuclei as a function of time
+        \begin{equation}
+            N_B(t)=\frac{\lambda_A}{\lambda_B-\lambda_A}N_0(e^{-\lambda_A t} - e^{-\lambda_B t}).
+            \label{eq:daughter}
+        \end{equation}
+        Similar derivations can be done for the rest of the isotopes in the decay chain by considering
+        one decay in a chain at a time. If the daughter isotope has a longer half-life than the parent
+        ($\lambda_A \gg \lambda_B$), for small $t$, Eq.~\ref{eq:daughter} becomes
+        $N_B(t)\sim N_0 e^{-\lambda_A t}$. After a long time,
+        Eq.~\ref{eq:daughter} becomes $N_B(t) \sim N_0 e^{-\lambda_B t}$; the decay
+        chain can be treated as the fast decay of $A$ into $B$, and then the slow decay of $B$
+        into $C$.
+
+        On the other hand, if the parent is longer lived
+        ($\lambda_B \gg \lambda_A)$, after a long time, Eq.~\ref{eq:daughter} can be written as
+        \begin{equation}
+            N_B(t)=\frac{\lambda_A}{\lambda_B}N_0 e^{-\lambda_A t}=\frac{\lambda_A}{\lambda_B}N_A(t).
+        \end{equation}
+        Hence the number of nuclei $A$ and nuclei $B$ are in equilibrium determined by their decay rates, with
+        $A$ dominating the nuclei count. Taking the limit as $\lambda_A \rightarrow 0$, $N_B(t)\rightarrow 0$.
+        The decay is essentially the decay of $A$ directly into $C$. Hence, if we have a chain of reaction
+        where each daughter nuclei has a half-life much shorter than that of its parent, then 
+        isotope $A$ effectively decays into the end of the decay chain and releases a lot more radiations than
+        in a single decay. Moreover, before the decay chain reaches its equilibrium, the \red{activity might do
+        funky stuff}
 
@@
         of energy $E_{\gamma}$
-        (in MeV), one can approximate the dose rate by 
+        (in MeV), one can approximate the dose rate by \cite[chapt. 7]{l01}
         \begin{equation}
@@
             \centering
-            \caption{Weighting factors for different types of radiation\cite{icrp74}}
+            \caption{Weighting factors for different types of radiation \cite{icrp74}}
             \begin{ruledtabular}
@@
         \begin{equation}
-            E = w_T H_T,
+            E = w_T \times H_T,
         \end{equation}
@@
             \centering
-            \caption{Weighting factors for individual organs\cite{icrp74}}
+            \caption{Weighting factors for individual organs \cite{icrp74}}
             \begin{ruledtabular}
@@
         UNSCEAR estimates that risk for severe hereditary effect is about $0.3-0.5$ percent per gray to
-        the first generation following the radiation exposure.\cite{u16, unscear12}
+        the first generation following the radiation exposure.\cite{u16, unscear16}
 
-%        \subsubsection{Limitations of current studies}
-%        \red{TODO}
+        \subsubsection{Limitations of current studies}
 
+    \subsection{Radiation Shielding}
+    \subsubsection{Nuclear cross-section}
+    \subsubsection{Exponential decay law}
+    \subsubsection{Shielding medium}
     \subsection{Fission}
@@
     nuclei that are not fissionable.\cite{lb01} We will explore neutron-induced fission further in
-    section~\ref{sec:fission}. Here we give an example of radiative capture. When a stable form
+    the next section. Here we give an example of radiative capture. When a stable form
     of gold ${}^{197}$Au absorbs a neutron, it forms a highly unstable isotope ${}^{198}$Au in its
@@
     uncommon, as the nucleus formed by neutron capture is usually unstable.
+
+        \begin{figure}[t]
+            \centering
+            \includegraphics[width=0.8\textwidth]{fission.png}
+            \caption{Illustration of the fission process for uranium-235.\cite{l01}}
+            \label{fig:fission}
+        \end{figure}
         \subsubsection{Fission and fission products}
@@
         On average, $2.5$ neutrons are produced per fission.\cite{l01}
-        \begin{figure}[h]
-            \centering
-            \includegraphics[width=0.8\textwidth]{fission.png}
-            \caption{Illustration of the fission process for uranium-235.\cite{l01}}
-            \label{fig:fission}
-        \end{figure}
 
@@
 
-        \begin{figure}[h]
+        \begin{figure}[t]
             \centering
@@
         good isolation for the radionuclides in the waste (See Fig.~\ref{fig:borehole}).
-        \begin{figure}[h]
+        \begin{figure}[t]
             \centering

### 2017-02-26 paper.tex
redo waste classification

@@
 \usepackage{amsmath}
+\usepackage{xspace}
 \usepackage{csquotes}
@@
 \begin{abstract}
-    As nuclear reactors around the world keep operating with no disposal method available for highly
-    radioactive nuclear waste, how we manage nuclear waste is becoming more important. In this work,
-    we examine a few scientific and technical related to nuclear waste management and disposal.
-    The principles of radioactivity and nuclear fission helps us understand the source and properties
-    of different types of nuclear waste. Focusing on the highly radioactive spent fuel, the rest
-    of this work outlines the advantages and drawbacks of different spent fuel management and disposal
-    methods. We study the methods used for performance assessments for Yucca Mountain Nuclear Waste Repository,
-    especially in the context of volcanic disruption. Finally, we discuss some possible options for 
-    nuclear waste management and disposal after
-    the termination of the Yucca Mountain project.\\
-    Word count: 6592\\
+%    As nuclear reactors around the world keep operating with no disposal method available for highly
+%    radioactive nuclear waste, how we manage nuclear waste is becoming more important. In this work,
+%    we examine a few scientific and technical related to nuclear waste management and disposal.
+%    The principles of radioactivity and nuclear fission helps us understand the source and properties
+%    of different types of nuclear waste. Focusing on the highly radioactive spent fuel, the rest
+%    of this work outlines the advantages and drawbacks of different spent fuel management and disposal
+%    methods. We study the methods used for performance assessments for Yucca Mountain Nuclear Waste Repository,
+%    especially in the context of volcanic disruption. Finally, we discuss some possible options for 
+%    nuclear waste management and disposal after
+%    the termination of the Yucca Mountain project.\\
+%    Word count: 6592\\
 \end{abstract}
@@
 
-\section{Radioactivity and Reactor Physics}
+\section{Radioactivity and Its Biological Effects}
 \label{sec:phys}
@@
         From Eq.~\ref{eq:lambda}, the number of particles between $t$ and $t+dt$ is $dN=\lambda N dt=\lambda
-        N_0 e^{}-\lamba t}dt$. Hence we have
+        N_0 e^{-\lambda t}dt$. Hence we have
         \begin{equation}
@@
                 \hline
-                Cesium-137 & 37 & 88\\
-                Radium-226 & 1.6\times 10^3 & 1\\
-                Plutonium-239 & 2.4\times 10^4 & 0.063\\
-                Uranium-233 & 1.6\times 10^6 & 9.3\times 10^{-3}\\
-                Thorium-232 & 1.4 \times 10^{10} & 1.1\times 10^{-7}\\
+                Cesium-137 & $37$ & $88$\\
+                Radium-226 & $1.6\times 10^3$ & $1$\\
+                Plutonium-239 & $2.4\times 10^4$ & $0.063$\\
+                Uranium-233 & $1.6\times 10^6$ & $9.3\times 10^{-3}$\\
+                Thorium-232 & $1.4 \times 10^{10}$ & $1.1\times 10^{-7}$\\
             \end{tabular}
@@
 
-    \subsection{Biological Effects of Radiation}
+        \subsection{Biological effects of radiation\footnote{Much of the discussion in this section follows that in
+        Lilley's text \cite[chapt. 7]{l01}}}
         The biological effects of radiation have been an important topic for decades. A large
@@
         changes that lead to cancer and genetic defects. Here we briefly summarize the known effects.
+        \subsubsection{Interactions of radiations with living tissues}
+        Radiations causes damage to living tissues by depositing energy into the biological molecules.
+        Depending on the type of the radiation
+        and the molecules present in the tissues, the damage mechanism and effects are different.
+        \paragraph{Alpha particles} An alpha particle loses energy predominantly via the interactions between its electric
+        field and the electrons, in the matter it travels through. Interactions with nuclei and head-on collisions
+        are quite rare. Since the speed of an alpha particle is low, it deposits more energy into the matter it
+        interacts with. In tissues, $1$-MeV alpha particles travel through only a few tens of micrometers and thus
+        are easily stopped by skin. However, if alpha particles are ingested and more sensitive organs are irradiated,
+        serious damages can be done.
+        \paragraph{Beta radiation (elections)} Electrons are much lighter and hence are easily scattered.
+        Moreover, they are much more penetrating since their energy lost per distance traveled is lower.
+        Finally, electrons also radiate photons
+        via bremstrahlung radiations as they deccelerate. Therefore, the electrons' energy end up distributing
+        across a much larger volume.
+        \paragraph{Photons} Gamma rays interact with matter via photoelectric effects, Compton scattering and pair productions
+        depending on th energy of the photon and the atoms it interact with.
+        Since biological tissues mainly consist of elements of low atomic number, Compton
+        scattering is the most important interaction. The scattered photons may interact further with the tissues
+        via photoelectric effects and secondary Compton scattering. The intensity of the energy deposition falls
+        off much more slowly and is hence more penetrating. Both electrons and photons can penetrate human body easily,
+        although they do not deposit as much energy, they generally are more dangerous than alpha particles due to their
+        penetration abilities.
+
+        The energy deposition described above is an important cause of irreversible damage to sensitive molecules
+        like the Deoxyribonucleic (DNA) molecules if the energ transferred is large enough.
+        In the cases where the energy transferred is not as large, the radiation ionizes simpler molecules like water
+        to form chemically active free radicals. The free radicals have unpaired electrons. Hence they go on to react and distrupt
+        other molecules that may be crucial to the biological system.
+
         \subsubsection{Quantifying radiation on living tissues}
@@
         exposure time, and the part of the body exposed. Various quantities have been introduced to
-        measure the amount of radiation and the biological impact of radiation. Here we review three
+        the biological impact of radiation. Here we review three
         of these quantities: the absorbed dose, the equivalent dose and the effective dose.
-        The SI units for all of these quantities are equivalent to J/Kg, the amount of energy deposited into
-        per unit mass of living tissue. However, this unit is called gray (Gy) for absorbed dose but sievert
-        for equivalent dose and effective dose.
+        The SI units for dose is siever (Sv), which is equivalent to J/Kg, the amount of energy deposited into
+        per unit mass of living tissue. Some work uses the unit gray (Gy) for abosrbed dose in medical context,
+        but it is the same as sievert.
 
-        The amount of radiation is usually characterized by the amount of energy it carries. Since one of
-        the earliest understood effects of radiation is its ability to ionize gas, one can quantify the
-        strength of the radiation with the ionizing energy of the gas.
+        The effects of radiation is usually characterized by the amount of energy it carries since it constrains
+        the amount of energy it deposits in living tissues.
         Absorbed dose~($D$) quantifies the 
-        radiation energy absorbed per unit mass. The modern unit is Gray(Gy)
-        and $1\text{Gy}=1\text{J kg}^{-1}$.\cite{u16}
-
-        Biological effects not only depend on the total dose applied but also the rate at which the 
-        dose is applied, because mechanisms exist within organisms that allow molecules like deoxyribonucleic acid
-        to be repaired if they are not excessively damaged. Therefore {\it dose rate} is also a useful
-        measure of radiation when considering its biological effect.
+        radiation energy absorbed per unit mass.
 
@@
         of energy $E_{\gamma}$
-        (in MeV), one can approximate the dose rate by \cite[chapt. 7]{l01}
+        (in MeV), one can approximate the dose rate (dose received per unit of time) by \cite[chapt. 7]{l01}
         \begin{equation}
@@
 
-        However, absorbed dose does not take into account different biological implications of different types
-        of radiation.
-        For example, the same dose of alpha particle can do more damage to body than gamma rays. The equivalent
-        dose ($H$) is defined as
-        \begin{equation}
-        H=w_R \times D_R,
-        \end{equation}
-        where $D_R$ is the absorbed dose and $w_R$ the weighting factor of a given kind of radiation $R$
-        (see Table~\ref{tab:eq} for a list of weighing factors). If there are multiple types of radiation 
-        present, the equivalent does is given by the weighted sum over all contributions. The SI unit for
-        equivalent dose is called the sievert (Sv) and $1\text{ Sv}=1\text{ J kg}^{-1}$.
-        \begin{table}
+        \begin{table}[h]
             \centering
@@
         \end{table}
+        However, absorbed dose does not take into account different biological implications of different types
+        of radiation.
+        For example, the same dose of alpha particle deposit more energy than gamma rays. The equivalent
+        dose ($H$) is defined as
+        \begin{equation}
+        H=w_R \times D_R,
+        \end{equation}
+        where $D_R$ is the absorbed dose and $w_R$ the weighting factor of a given kind of radiation $R$
+        (see Table~\ref{tab:eq} for a list of weighing factors).
 
@@
         at question.
-        Evidently, if the radiation affects multiple tissues, the effective dose will be the weighted
-        sum of the equivalent dose over individual body parts. Some of the recommended values of tissue
+        Some of the recommended values of tissue
         weighting factors are given in Table~\ref{tab:eff}. A weighing factor of $1$ is equivalent to the
-        whole body absorbing the radiation uniformly. The SI unit of effective dose is sievert (Sv),
-        identical to that of equivalent dose.
-        \begin{table}
+        whole body absorbing the radiation uniformly.
+        \begin{table}[h]
             \centering
@@
         \end{table}
-        \subsubsection{Effects of radiation on human}
-        Short-term effects of radiation are usually caused by extensive cell death or cell damage.\cite{u16}
-        These effects are deterministic, meaning that below a certain threshold the effect does not occur and
-        the severity of the effect depends on the dose. Some examples are gastrointestinal damage and impairment
-        of fertility.\cite{u16, l01}
+
+        \subsubsection{Effects of radiation on humans}
+        Short-term effects of radiation are usually caused by extensive cell death or cell damage, resulting
+        from large-scale damages of critical molecules.\cite{u16}
+        These effects are deterministic with a threshold,
+        meaning that below a certain threshold the effect will not occur and
+        the severity of the effect depends on the effective dose. Roughly, below the certain threshold,
+        not enough free radicals are created and hence the repairing capabilities of cells can undo the damages.
+        If the radiation level is high, the radiation may generate enough free radicals to disrupt the cells or directly
+        ionize critical molecules. Some examples are gastrointestinal damage and impairment
+        of fertility.\cite{u16}
 
@@
         no apparent threshold and the severity is usually independent of the dose but
-        the chance of occurrence depends on the dose.\cite{u16,l01}
+        the chance of occurrence depends on the dose.\cite{u16}
+        These effects mainly result from the minor disruptions of DNA molecules since the disruptions are replicated
+        as cells reproduce.
         For example, according to a review done by the UN Scientific Committee on the Effects of Atomic Radiation
@@
         Hereditary effects of radiation can occur if the reproduction cells were damaged by radiation.
-        UNSCEAR estimates that risk for severe hereditary effect is about $0.3-0.5$ percent per gray to
+        The mechanism is similar to that of delayed health effects, since DNA molecules are passed on
+        to offpsrings.
+        UNSCEAR estimates that risk for severe hereditary effect is about $0.3-0.5$ percent per siever to
         the first generation following the radiation exposure.\cite{u16, unscear16}
-
-        \subsubsection{Limitations of current studies}
-
-    \subsection{Radiation Shielding}
-    \subsubsection{Nuclear cross-section}
-    \subsubsection{Exponential decay law}
-    \subsubsection{Shielding medium}
-    \subsection{Fission}
-    \label{sec:fission}
-        Nuclear Fission is the process in which a heavy nucleus splits into smaller fragments and releases
-        a huge amount of energy. It is the main physical process taking place in a nuclear reactor. It is
-        also the main mechanism that generates the most radiative wastes that we will discuss later
-        in this work. Therefore, we should first familiarize ourselves with fission.
-    \subsubsection{Neutron absorption}
-    \label{sec:capture}
-    Thermal (slow) neutrons induce fissions in a reactor.
-    Since a neutron does not carry any charge, it experiences a much smaller potential when it interacts
-    with a nucleus. In the nuclear absorption reaction, a neutron crosses the potential barrier to form
-    an excited nucleus.
-    The nuclear absorption reaction is the most crucial type of reactions that takes place in
-    a nuclear reactor. As neutrons travels through materials, they react readily with nuclei.
-    For fast neutrons, many reactions are possible. But for slow or thermal neutrons (as in the
-    case of a nuclear reactor), neutron absorption is the primary mode of interaction.
-    Depending on the path that the compound nucleus takes to return to its ground
-    state, neutron absorption is classified into radiative capture and neutron-induced
-    fission(see section~\ref{sec:fission}).
-
-    In a radiative capture process, the compound nucleus emits photons in the form of gamma rays, whereas in neutron-induced
-    fission, the compound nucleus splits into fragments.
-    In fact, not all nuclei can undergo fission.
-    Therefore, radiative capture is the only possible neutron absorption process for
-    nuclei that are not fissionable.\cite{lb01} We will explore neutron-induced fission further in
-    the next section. Here we give an example of radiative capture. When a stable form
-    of gold ${}^{197}$Au absorbs a neutron, it forms a highly unstable isotope ${}^{198}$Au in its
-    excited state and quickly decays into its ground state by emitting $\gamma$ rays. This can be
-    written as
-    \begin{equation}
-        {}^{197}\text{Au} + n \rightarrow {}^{198}\text{Au} + \gamma.
-    \end{equation}
-    Moreover, the isotope ${}^{198}$Au can spontaneously decay into ${}^{198}$Hg via beta decay. This is not
-    uncommon, as the nucleus formed by neutron capture is usually unstable.
-
-        \begin{figure}[t]
-            \centering
-            \includegraphics[width=0.8\textwidth]{fission.png}
-            \caption{Illustration of the fission process for uranium-235.\cite{l01}}
-            \label{fig:fission}
-        \end{figure}
-        \subsubsection{Fission and fission products}
-        Conjectured by Bohr and proven by Alfred O. Nier in 1939,
-        the fission of uranium-235 induced by a thermal neutron
-        is the best-known example of induced fission. 
-        The neutron capture produces a compound nucleus ${}^{236}U$ that is in an excited state.
-        Sometimes the nucleus can decay by $\gamma$ emission in capture reaction.
-        But most of the time, the excitation energy deforms the nucleus (see Fig.~\ref{fig:fission}).
-        When the
-        deformation reaches a certain point, the Coulomb force overcomes the short-ranged nuclear
-        force (residual strong force) and the nucleus disintegrate into two large fragments and several neutrons. The
-        positively-charged fragments(fission products) attain approximately $170$ MeV of kinetic energy as
-        Coulomb force continues to drive them apart.\cite{l01}
-        A typical reaction is described by
-        \begin{equation}
-        {}^{235}_{92}\text{U} + n \rightarrow {}^{92}_{36}\text{Kr} + {}^{142}_{56}\text{Ba} + 2 n.
-        \end{equation}
-        The energy liberated is about $180$ MeV. The combination of fission products is not
-        unique. \cite{w98, gc01}
-        Very often, photons, electrons and neutrinos are also emitted through beta and gamma decay.
-        On average, $2.5$ neutrons are produced per fission.\cite{l01}
-
-        \subsubsection{Fission energy budget}
-        We can estimate the amount of energy released in fission by considering the binding energy difference
-        and the properties of the product. Essentially, any energy carried by particles that can be slowed down 
-        in the reactor can be harvested as heat to generate power.
-
-        The binding energy per nucleon is about $7.6$ MeV u${}^{-1}$ for uranium and about $8.5$ MeV u${}^{-1}$
-        for a nucleus with atomic number around $117$. Hence, the change of binding energy per nucleon is
-        $0.9$ MeV u${}^{-1}$, which amount to $212$ MeV for uranium-235.
-        About $87\%$ of the total energy is promptly emitted because
-        the fission products, carrying over $90\%$ of the total energy, travel for only a fraction of a
-        millimeter before they are stopped. Besides, fission neutrons emit about $2$ MeV of energy by
-        successive scattering in the reactor.
-        About $13\%$ of the total energy goes into radioactivity. Energy carried by electrons and photons is
-        convertible to heat but the neutrinos, carrying about $50\%$ of the radiation energy, are not stoppable.
-
-        Henceforth, it is generally accepted that about $220$ MeV per fission is recoverable for energy conversion.
-        The energy generated by the fission of uranium-235 is 
-        about 2.5 million times the energy produced by burning coal of the same mass.\cite{e17}
-%        \subsubsection{Chain reaction??}
-
 \section{Nuclear waste: classification and properties}
 \label{sec:waste}
+    In this section, we will first look at a classification scheme of nuclear waste and then
+    examine in details the source, composition and properties of the most detrimental type, high-level waste.
+    \subsection{Classification of nuclear waste}
     Different organizations have slightly different classification schemes for nuclear waste.
@@
     A form of waste may possesses a high activity content due to
-    a high concentration of radionuclides or the presence of radionuclides with high activity.\cite{iaea09}
+    a high concentration of radionuclides and/or the presence of radionuclides with high activity.\cite{iaea09}
     Each form of waste, depending on the half-life, activity content, and temperature, requires different
-    disposal methods.
+    disposal methods. The NRC classification scheme is as follows.\cite{nrc09}
 
-    We will focus on high-level waste (HLW, defined below) in the rest of this work.
-    In this section, we will first examine the composition and properties of HLW and then
-    comment on other categories of waste and disposal methods (following the NRC classification scheme). 
-    \subsection{High-level waste(HLW)}
-    High-level Wastes (HLW) has high activity contents and usually generates a large amount of
+    \paragraph{High-level waste (HLW) and spent fuel}
+    HLW has high activity contents and usually generates a large amount of
     heat as it radiates. The typical source of HLW is spent fuel from the nuclear reactor.
@@
 
-    \subsection{Transuranic waste (TRU)}
-    \label{sec:tru}
-    Transuranic waste is defined as materials that do not belong to high-level waste and are
+    \paragraph{Transuranic waste (TRU)}
+    TRU is defined as materials that do not belong to high-level waste and are
     contaminated by alpha-emitting radionuclides with
@@
 
-    \subsection{Low-level and intermediate-level waste (LLW and ILW)}
-    Low-level waste (LLW) has relatively little radioactivity, produces almost no heat,
+    \paragraph{Low-level waste (LLW)}
+    LLW has relatively little radioactivity, produces almost no heat,
     and contains practically no
@@
     higher radioactivity such that they are classified
-    as intermediate-level waste (ILW). ILL are solidified in concrete and buried in shallow
+    as intermediate-level waste (ILW). ILW are solidified in concrete and buried in shallow
     repositories.\cite{s01} However, US regulations do not make the distinction
@@
 
-    \subsection{Uranium tailings}
+    \paragraph{Uranium tailings}
     Uranium mill tailings from uranium mills are another form of waste with low activity, but
@@
     radioactive thorium and radium and these radionuclides stay in the tailing. Notably, radium
-    can decay into radon gas which is an alpha-emitter. Hence, the alpha radiation can do additional
+    can decay into radon gas which is an alpha-emitter. Hence, the alpha radiation may do additional
     damage to body if the radon enters the body as one breathes.
@@
 
+    Since all categories but high-level waste are relatively easy to handle, we deal with the formation and
+    properties of spent nuclear fuel, the dominant contribution to high level waste, in the rest of this section.
+
+    \subsection{Creation of spent fuel}
+    \subsubsection{Nuclear fuel}
+    For most of the commercial nuclear reactors, the energy comes from the isotope uranium-235 which
+    capable of undergoing fission (see Sec.~\ref{sec:fission}). However, natural uranium consists of $99.3\%$
+    non-fissile uranium-238 and only $0.7\%$ uranium-235. Therefore, for nuclear fuel, natural uranium is enriched
+    via isotope separation to increase the uranium-235 content. Typical fresh nuclear fuel contains $3\%$ uranium-235
+    and $97\%$ uranium-238.
+    \subsubsection{Neutron absorption and radiative capture}
+    \label{sec:capture}
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.6\textwidth]{pu.png}
+            \caption{Schematic illustration of the two different neutron absorption processes:
+            fission of uranium-235 and the radiative absorption of uranium-238.}
+            \label{fig:both}
+        \end{figure}
+    Absorption of thermal neutron (with energy $\sim 0.025$ eV, the most likely speed of a free neutron at 
+    a temperature of $290$ K) is the main reaction that
+    occurs in the reactor. The neutrons come from a startup neutron source and subsequently from the fission
+    process (see Sec.~\ref{sec:fission}).
+    Since a neutron does not carry any charge, it does not experience the Coulomb repulsion potential when
+    it interacts
+    with a nucleus.
+    Henceforth, as neutrons travels through materials, they react readily with nuclei.
+    For fast neutrons, many reactions are possible but they mostly just scatter off heavy nuclei.
+    But for thermal neutrons (as in the case of a nuclear reactor), neutron absorption
+    is the primary mode of interaction between the nuclei and the neutrons.
+    Neutron absorption is the process in which a nucleus absortb a neutron to form an unstable
+    nucleus.
+    Depending on the path that the unstable nucleus takes to return to its ground
+    state, neutron absorption is classified into radiative capture if the nucleus radiates to reach a stable
+    state and neutron-induced fission if the nucleus splits into fragments (see Fig.~\ref{fig:both}).
+    In fact, not all nuclei can undergo fission.
+    Therefore, radiative capture is the only possible neutron absorption process for
+    nuclei that are not fissionable.\cite{lb01} We will explore neutron-induced fission further in
+    the next section. Here we give an example of radiative capture. Thorium-232, which is present in highly radioactive
+    waste, has a half-life longer than the age of the universe (see Table \ref{tab:hl}). However, it can go through
+    radiative capture and then subsequent beta decays to form uranium-233 (neutrinos are omitted in the equations)
+    \begin{align}
+        n + {}^{232}\text{Th} &\rightarrow {}^{233}\text{Th} + \gamma,\\
+        {}^{233}\text{Th} &\rightarrow {}^{233}\text{Pa} + e^-,\\
+        {}^{233}\text{Pa} &\rightarrow {}^{233}\text{U} + e^-.
+    \end{align}
+    The uranium-233 formed has a half-life of $1.4$ million years and hence are a lot more dangerous than
+    thorium-232. Another important radiative capture reaction happening inside nuclear fuel is the radiative
+    capture of uranium-238 (the most abundunt isotope in nuclear fuel) to form plutonium-239, the {\it de facto}
+    isotope to construct atomic bombs
+    \begin{equation}
+        n + {}^{238}\text{U} \rightarrow {}^{239}\text{Pu} + \gamma
+    \end{equation}
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.6\textwidth]{fission.png}
+            \caption{Illustration of the thermal neutron induced fission process of uranium-235.\cite{l01}}
+            \label{fig:fission}
+        \end{figure}
+        \subsubsection{Fission and fission products}
+        \label{sec:fission}
+        Conjectured by Bohr and proven by Alfred O. Nier in 1939,
+        the fission of uranium-235 induced by a thermal neutron
+        is the best-known example of induced fission. 
+        The neutron capture produces a compound nucleus ${}^{236}U$ that is in an excited state.
+        Sometimes the nucleus can decay by $\gamma$ emission in capture reaction.
+        But most of the time, the excitation energy deforms the nucleus into two positively charged
+        parts (see Fig.~\ref{fig:fission}).
+        The nucleus starts to vibrate with higher and higher amplitude as the kinetic energy of the
+        neutron transfers to the vibrational energy of the nucleus.
+        When the
+        amplitude reaches a critical point, the Coulomb force overcomes the short-ranged nuclear
+        force (predominantly strong force) and the nucleus disintegrates into two large fragments and several neutrons.
+        A typical reaction is described by
+        \begin{equation}
+        {}^{235}_{92}\text{U} + n \rightarrow {}^{92}_{36}\text{Kr} + {}^{142}_{56}\text{Ba} + 2 n.
+        \label{eq:u235}
+        \end{equation}
+        Very often, photons, electrons and neutrinos are also emitted through beta and gamma decay.
+        The reactor described by Eq.~\ref{eq:u235} is not unique in different fission products can result
+        from a fission.
+        On average, inside a nuclear reactor, $2.5$ neutrons are produced per fission, causing a ``chain reaction'' of subsequent
+        fissions that generate a huge amount of energy.\cite[chapt. 10]{l01}
+
+        \subsubsection{Fission energy}
+        We can estimate the amount of energy released in fission by considering the binding energy difference
+        between the parent nuclei and the fission products. As a consequence of the masss-energy equivalence in the theory
+        of special relativity, we can calculate the energy released $E$ from the atomic mass differences between the parent $m_{parent}$
+        and the fission products $m_{products}$.
+        \begin{equation}
+            E = (m_{parent}-m_{product})c^2,
+        \end{equation}
+        where $c$ is the speed of light.
+        Essentially, any energy carried by particles that can be slowed down 
+        in the reactor can be harvested as heat to generate power.
+
+        The binding energy per nucleon is about $7.6$ MeV u${}^{-1}$ for uranium and about $8.5$ MeV u${}^{-1}$
+        for a nucleus with atomic number around $117$. Hence, the change of binding energy per nucleon is
+        $0.9$ MeV u${}^{-1}$, which amount to $212$ MeV for a uranium-235 atom.
+        Apart from $6.5\%$ of th energy carried by the neutrinos which does not interact with the surrounding,
+        the rest of the energy can be utilized by virtue of outgoing particles depositing energy in the medium.
+
+        In reality, thermal induced fission of uranium-235 creates a variety of products and hence the energy released
+        in a single reaction vary by tens of MeV. On average, about $220$ MeV per fission is recoverable for energy conversion.
+        The energy generated by the fission of uranium-235 is 
+        about 2.5 million times the energy produced by burning coal of the same mass.\cite{e17}
+
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.6\textwidth]{spent-fuel.png}
+            \caption{Composition of spent nuclear fuel for a light water reactor. The exact composition varies
+            by different kinds of nuclear reactors.}
+            \label{fig:spentfuel}
+        \end{figure}
+        \subsubsection{Composition and properties of spent fuel}
+        neutron, hot,varying half-lives
     \section{Spent fuel management}
@@
     happen in this country for various historical reasons. 
+    \subsection{Radiation shielding}
+    cross section, exponential decay, attenuation
     \subsection{Spent fuel pools}
@@
     50 years.\cite{a11, iaea99}
+
+    \subsection{Dry cask storage}
+    Spent fuel pools have very limited capacities. 
+    Overground dry cask storage provides a low-cost temporary storage before the spent fuel is disposed of.
+    After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks filled with inert gas. 
+The dry casks are often metal cylinders welded to prevent leakage. Each cask is surrounded by
+    additional steel and concrete to provide shielding.
+    The heat from the cask is removed by passive
+    cooling (air circulation). The NRC is hoping that the casks will
+    last safely for as long as 100 years, yet it has been observed that corrosions and cracking occur in 30 years
+    or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
+    overground dry cask storage does not seem to be a long-term solution. The doubts were expressed
+    during the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
+    \begin{displayquote}
+    Leaving the spent fuel on-site for extended periods of time was never intended and is not
+    responsible. (Dry cask manufacturer)
+
+    It is not ethical, basically to plan for long-term storage without pursuing a well defined repository
+    program. (Department of Energy employee speaking as a citizen)
+    \end{displayquote}
+    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in densely
+    packed pools
+    that are close to or over their limiting capacity.\cite{a11,aa12}
+    Therefore, there is clearly a need for a long-term disposal plan, which we will explore in
+    Sec.~\ref{sec:disposal}.
 
@@
 
-    \subsection{Dry cask storage}
-    Spent fuel pools have very limited capacities. 
-    Overground dry cask storage provides a low-cost temporary storage before the spent fuel is disposed of.
-    After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks filled with inert gas. 
-The dry casks are often metal cylinders welded to prevent leakage. Each cask is surrounded by
-    additional steel and concrete to provide shielding.
-    The heat from the cask is removed by passive
-    cooling (air circulation). The NRC is hoping that the casks will
-    last safely for as long as 100 years, yet it has been observed that corrosions and cracking occur in 30 years
-    or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
-    overground dry cask storage does not seem to be a long-term solution. The doubts were expressed
-    during the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
-    \begin{displayquote}
-    Leaving the spent fuel on-site for extended periods of time was never intended and is not
-    responsible. (Dry cask manufacturer)
-
-    It is not ethical, basically to plan for long-term storage without pursuing a well defined repository
-    program. (Department of Energy employee speaking as a citizen)
-    \end{displayquote}
-    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in densely
-    packed pools
-    that are close to or over their limiting capacity.\cite{a11,aa12}
-    Therefore, there is clearly a need for a long-term disposal plan, which we will explore in
-    the next section.
-
 \section{Disposal of HLW}
@@
     The time frame of nuclear waste disposal is about $10^4-10^6$ years, after which the radioactivity
-    of the waste is predicted to decline to a safe level.
+    of the waste is predicted to decline to a safe level (see Sec.\ref{sec:spentfuel}).
     Deep geological repositories that are stable over geological timescales naturally offer a good option for disposal.
@@
 
-        \begin{figure}[t]
-            \centering
-            \includegraphics[width=0.9\textwidth]{volcano.png}
-            \caption{A flowchart for the volcanism risk estimation problem. The oval-shaped
-            parts are all probabilistic. \ref{sec:volcano} \cite{cv14}}
-            \label{fig:volcanos}
-        \end{figure}
-
         The information necessary to estimate the probability is listed in the flowchart Fig.~\ref{fig:volcanos}.
@@
         life span of the repository. Here we look in some detail at the calculation of those parameters.
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.9\textwidth]{volcano.png}
+            \caption{A flowchart for the volcanism risk estimation problem. The oval-shaped
+            parts are all probabilistic. \ref{sec:volcano} \cite{cv14}}
+            \label{fig:volcanos}
+        \end{figure}
+
         \paragraph{Monte-Carlo simulation}
@@
         good isolation for the radionuclides in the waste (See Fig.~\ref{fig:borehole}).
-        \begin{figure}[t]
+        \begin{figure}[h]
             \centering

### 2017-02-27 paper.tex
ready version 2 modulo typos

@@
 \begin{abstract}
-%    As nuclear reactors around the world keep operating with no disposal method available for highly
-%    radioactive nuclear waste, how we manage nuclear waste is becoming more important. In this work,
-%    we examine a few scientific and technical related to nuclear waste management and disposal.
-%    The principles of radioactivity and nuclear fission helps us understand the source and properties
-%    of different types of nuclear waste. Focusing on the highly radioactive spent fuel, the rest
-%    of this work outlines the advantages and drawbacks of different spent fuel management and disposal
-%    methods. We study the methods used for performance assessments for Yucca Mountain Nuclear Waste Repository,
-%    especially in the context of volcanic disruption. Finally, we discuss some possible options for 
-%    nuclear waste management and disposal after
-%    the termination of the Yucca Mountain project.\\
-%    Word count: 6592\\
+    As nuclears around the world continuously generate highly radioactive waste,
+    how we manage nuclear waste is becoming more important. In this work,
+    we examine a few scientific and technical issues related to nuclear waste management and disposal.
+    The principles of radioactivity and nuclear fission help us to understand the sources and properties
+    of different types of nuclear waste. Focusing on the highly radioactive spent fuel, the rest
+    of this work outlines the advantages and drawbacks of different spent fuel management and disposal
+    methods. We study the methods used for performance assessments for Yucca Mountain Nuclear Waste Repository,
+    especially in the context of volcanic disruption. Finally, we discuss some possible options for 
+    nuclear waste management and disposal after
+    the termination of the Yucca Mountain project.\\
+    Word count: 7449\\
 \end{abstract}
@@
 for waste classified as Transuranic, which, by the Nuclear Regulatory Commission standard, does not have particularly
-high radioactivity but only long half-life (See Sec.~\ref{sec:tru}). In particular, WIPP does not accept spent nuclear
+high radioactivity but only long half-life (See Sec.~\ref{sec:waste:classification}). In particular, WIPP does not accept spent nuclear
 fuel.}
@@
 How well is the predicative power of science? How much trust should we put into the uncertainty estimates
-(i.e. \red{how well can science predict its predicative power})?
 
@@
     in this work. We will apply these results in our discussion of nuclear wastes.
-    \subsection{Radioactivity}
-    Radioactive is the process in which most of the known detrimental radiations are generated.
+    \subsection{Radioactivity\footnote{
     The derivations and in this section follows chapter $6$ in
     Krane's text \cite{k88} and chapter $1$ in Lilley's text \cite{l01}.
+    }}
+    Radioactive decay is the process in which most of the known detrimental radiations are generated.
+    Here we review the physical principles and mathematical formalisms behind it.
         \subsubsection{Quantum theory of radioactive decays}
-        The law of radioactive decay is a direct consequence of it being a quantum-mechanical
-        effect.
+        The law of radioactive decay is a direct consequence of quantum mechanics.
         We can approximate the nucleus as a particle situated in a potential well with potential given by
@@
             \centering
-            \caption{Specific activity of isotopes commonly present in spent nuclear fuel \cite{toi}}
+            \caption{Specific activity and hal-life of isotopes commonly present in spent nuclear fuel \cite{toi}}
             \begin{ruledtabular}
@@
                 Cesium-137 & $37$ & $88$\\
-                Radium-226 & $1.6\times 10^3$ & $1$\\
+                Thorium-232 & $1.4 \times 10^{10}$ & $1.1\times 10^{-7}$\\
+                Uranium-233 & $1.6\times 10^6$ & $9.3\times 10^{-3}$\\
                 Plutonium-239 & $2.4\times 10^4$ & $0.063$\\
-                Uranium-233 & $1.6\times 10^6$ & $9.3\times 10^{-3}$\\
-                Thorium-232 & $1.4 \times 10^{10}$ & $1.1\times 10^{-7}$\\
+                Neptunium-237 & $2.1\times 10^6$ & $6.9\times 10^{-4}$\\
             \end{tabular}
@@
         There are three main types of radioactive decay processes that generate radiations in nuclear waste: 
-        Alpha decay, which generates Alpha particles (helium-4 nuclei), Beta decay which generates Beta
-        particles (electrons) and Gamma decay which generates gamma rays.
+        Alpha decay, which generates alpha particles (helium-4 nuclei), Beta decay which generates beta
+        particles (electrons) and gamma decay which generates gamma rays.
         The type of radiation (particle beam) generated
@@
 
-        \subsubsection{Daughter nuclei and subsequent decay}
+        \subsubsection{Daughter nuclei and subsequent decay\footnote{The derivation in this section follows that of
+        \cite[chapt. 1.5]{l01}}}
         The resultant nuclei (called ``daughter products'') from a radioactive decay are oftern radioactive
@@
         isotope $A$ effectively decays into the end of the decay chain and releases a lot more radiations than
-        in a single decay. Moreover, before the decay chain reaches its equilibrium, the \red{activity might do
-        funky stuff}
+        in a single decay. Moreover, before the decay chain reaches its equilibrium, the activity of the sample
+        may increase.
 
@@
         \subsubsection{Interactions of radiations with living tissues}
+        \label{sec:interactions}
         Radiations causes damage to living tissues by depositing energy into the biological molecules.
@@
     \subsection{Classification of nuclear waste}
+    \label{sec:waste:classification}
     Different organizations have slightly different classification schemes for nuclear waste.
@@
     In fact, since there is no commercial reprocessing facilities in the US (see section~\ref{sec:reproc}),
-    almost all of the HLW is in the form of spent nuclear fuel from the reactor. In a nuclear reactor,
-    fission of uranium-235 produces lighter radionuclides like cesium-137 and strontium-90, which
-    account for most of the heat and penetrating radiation (gamma radiations). In addition to
-    the fission process, some uranium-235 atoms go through radiative capture
-    (see section~\ref{sec:capture}) and produce radionuclides that are
-    heavier than uranium (``transuranic'') such as plutonium. The transuranic atoms do not generate
-    as much heat or radiation from their subsequent decay as the fission products, but they tend to have
-    a much longer half-life.
-    For example, strontium-90 and cesium-137 have half-lives of about $30$ years, whereas
-    the transuranic plutonium-239 has a half-life of $24,000$ years.
-
+    almost all of the HLW is in the form of spent nuclear fuel from the reactor.
     \paragraph{Transuranic waste (TRU)}
@@
     facilities after on-site short-term decay storage.
-    Some LLW (for example, metal nuclear fuel cladding) may have a slightly
+    Some LLW (for example, metal nuclear fuel packaging) may have a slightly
     higher radioactivity such that they are classified
@@
     Uranium mill tailings from uranium mills are another form of waste with low activity, but
-    they are not usually classified as LLW. The tailing is what is left after the uranium has been
-    extracted from the ore. The decay of uranium-238 from the mill decays into
-    radioactive thorium and radium and these radionuclides stay in the tailing. Notably, radium
+    they are not usually classified as LLW.
+    Over geological timescales, the stable uranium-238 decays into radioactive thorium and radium 
+    These radionuclides remain after the uraninium (all of its isotopes)
+    has been extracted from the ore. Notably, radium
     can decay into radon gas which is an alpha-emitter. Hence, the alpha radiation may do additional
@@
     \subsection{Creation of spent fuel}
+    \label{sec:waste:spentfuel}
     \subsubsection{Nuclear fuel}
@@
     \label{sec:capture}
-        \begin{figure}[h]
-            \centering
-            \includegraphics[width=0.6\textwidth]{pu.png}
-            \caption{Schematic illustration of the two different neutron absorption processes:
-            fission of uranium-235 and the radiative absorption of uranium-238.}
-            \label{fig:both}
-        \end{figure}
     Absorption of thermal neutron (with energy $\sim 0.025$ eV, the most likely speed of a free neutron at 
@@
     nucleus.
+
     Depending on the path that the unstable nucleus takes to return to its ground
@@
     state and neutron-induced fission if the nucleus splits into fragments (see Fig.~\ref{fig:both}).
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.8\textwidth]{pu.png}
+            \caption{Schematic illustration of the two different neutron absorption processes:
+            fission of uranium-235 and the radiative absorption of uranium-238. Adapted from \cite{iync_2013}}
+            \label{fig:both}
+        \end{figure}
     In fact, not all nuclei can undergo fission.
@@
     \end{equation}
-        \begin{figure}[h]
-            \centering
-            \includegraphics[width=0.6\textwidth]{fission.png}
-            \caption{Illustration of the thermal neutron induced fission process of uranium-235.\cite{l01}}
-            \label{fig:fission}
-        \end{figure}
         \subsubsection{Fission and fission products}
@@
         Very often, photons, electrons and neutrinos are also emitted through beta and gamma decay.
-        The reactor described by Eq.~\ref{eq:u235} is not unique in different fission products can result
+        The reaction described by Eq.~\ref{eq:u235} is not unique because different fission products can result
         from a fission.
-        On average, inside a nuclear reactor, $2.5$ neutrons are produced per fission, causing a ``chain reaction'' of subsequent
+        On average, inside a nuclear reactor, $2.5$ neutrons are produced per fission. These neutrons
+        continue to induce fissions of other uranium-235 nuclei, causing a ``chain reaction'' of subsequent
         fissions that generate a huge amount of energy.\cite[chapt. 10]{l01}
-
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.8\textwidth]{fission.png}
+            \caption{Illustration of the thermal neutron induced fission process of uranium-235.\cite{l01}}
+            \label{fig:fission}
+        \end{figure}
         \subsubsection{Fission energy}
@@
 
+        \subsubsection{Composition and properties of spent fuel}
+        Fission and radiative capture are respectively responsible for the creation of two very
+        different classes of radionuclides in spent fuel
+        (see Fig.~\ref{fig:spentfuel} for an illustration of the composition of spent fuel).
         \begin{figure}[h]
@@
             \caption{Composition of spent nuclear fuel for a light water reactor. The exact composition varies
-            by different kinds of nuclear reactors.}
+            by different kinds of nuclear reactors. ``Minor actinides'' refer to transuranic elements other than plutonium.\cite{gao}}
             \label{fig:spentfuel}
         \end{figure}
-        \subsubsection{Composition and properties of spent fuel}
-        neutron, hot,varying half-lives
+
+        On the one hand, fission of uranium-235 produces lighter radionuclides like cesium-137 and strontium-90, which
+        account for most of the short-term high activity and penetrating radiation (gamma radiations).
+        In addition, for spent fuel freshly removed from the reactor, there is still a small amount ($\sim 1\%$)
+        of uranium-235 left along with a lot of free neutrons released from previous fissions. Thus, a large amount
+        of energy is generated via fission of the leftover uranium-235 and heats up the spent fuel rods.
+
+        On the other hand, some uranium-238 nuclei go through radiative capture
+        and produce radionuclides that are
+        heavier than uranium (``transuranic'') such as plutonium, neptunium and americum.
+        These transuranic nuclei have relatively low activity but their half-lives are from a few thousand years
+        to a million years (see Table~\ref{tab:hl}).
+        Even after a long time, their activity level remains dangerous (see Fig.~\ref{fig:csru} for
+        an illustration of why this is the case).
+        Moreover, these transuranic nuclei can be readily turned into plutonium-239 and proliferation is also
+        an issue when one deals with these elements.
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.6\textwidth]{csur.png}
+            \caption{Activty of $1$ gram of cesium-137 (half-life $30$ years) and uranium-233 (half-life $1.3$ million
+            years) as a function of time in years. The activity of cesium-137 falls off to virtually zero after a few
+        hundred years; whereas the activity of uranium-233 remains about $0.01$ Ci for thousands of years. The activity
+    level of $0.01$ Ci is comparable to that of the radiation source used for Computer Tomography (CT) scans.
+Data from Table~\ref{tab:hl}}
+            \label{fig:csru}
+        \end{figure}
+        Therefore, in the short term, spent fuel emits intense radiations and generates a large amount of heat.
+        In the very long term (up to millions of years), its activity still remains well above the safe level.
+        The combination of these properties makes the management and disposal of spent fuel and high-level waste
+        in general very difficult.
+
     \section{Spent fuel management}
     \label{sec:temp}
-    Spent fuel from nuclear reactors constitutes the major source of HLW. Spent fuel pools at the
-    reactors hold spent fuel cores to cool them down and isolate them for a few years. After that,
-    reprocessing can occur to extract the most radioactive plutonium and uranium isotopes and
+    Spent fuel from nuclear reactors is the major source of HLW.
+    The concept of shielding is crucial in the short-term management of spent fuel due to the high activity
+    it contains. Freshly removed spent fuel rods are stored in spent fuel pools for at least $3$ years before
+    the next step. Reprocessing can extract the long-lived plutonium and uranium isotopes and
     convert the waste into more manageable form (vitrification). However, reprocessing of commercial
-    nuclear spent fuel did not
-    happen in this country for various historical reasons. 
+    nuclear spent fuel did not happen in the US for cost and proliferation concerns. Instead, dry-cask
+    storage, in addition to spent fuel pools, is used as an interim storage methods.
     \subsection{Radiation shielding}
-    cross section, exponential decay, attenuation
+    Since spent fuel contains high activity, shielding of radiations is crucial. Shielding works by
+    forcing the radiation to interact with the shielding medium and hence reducing the amount of energy
+    the radiation carries. In the context of nuclear waste management, the most common shielding medium
+    are water and concrete; and the mode of interactions of radiations with those mediums are similar to
+    the the physical intereaction between radiations and biological molecules (the most abundant of which is
+    water) described in Sec.~\ref{sec:interactions}.
+
+    To quantify the shielding capability of a given material, we introduce cross section $\sigma$. Consider
+    a beam of particles with intensity $I_i$ (with unit (area$\cdot$time)${}^{-1}$) passing through a medium
+    with density of molecules $N_s$. Suppose on average $R_i$ interactions occur per time unit.
+    The cross section of the iteraction between the shielding medium defined as
+    \begin{equation}
+        \sigma = \frac{R_i}{I_i N_s}.
+    \end{equation}
+    Effectively, $\sigma$ has unit of area and
+    is proportional to the interaction probability of the incident radiation with a given nucleus.\cite{k88}
+    Notice that the interaction probability is independent of the distance traveled by the beam in
+    the medium. Written in terms of the beam intensity as a function of distance traveled $z$,
+    the relationship can be expressed as
+    \begin{equation}
+        -\mu = \frac{d I/dz}{I},
+    \end{equation}
+    with the attenuation coefficient $\mu$ defined by
+    \begin{equation}
+        \mu = \sigma N_s
+    \end{equation}
+    where $N_s$ is the number density of molecules in the medium.
+    Solving the differential
+    equation yields another exponential attenuation law
+    \begin{equation}
+        \label{eq:atten}
+        I(z) = I_0 e^{-\mu z},
+    \end{equation}
+    where $I_0$ is the incident beam intensity. Therefore, the beam intensity decays exponentially
+    with the distance traveled. Notice that $\mu$ and $z$ play symmetrical roles in the exponent. Hence
+    in practice, we can trade quantity of shielding medium for the shielding capability of the type of
+    molecule.
+
     \subsection{Spent fuel pools}
-    The heat output of the spent fuel decreases by $99\%$ after the first year and by another factor of $5$
-    by the 5-year mark.\cite{aa12} Therefore, short-term cooling of spent fuel is crucial before
-    further disposal or reprocessing can be done.
-    US nuclear power plants generally have an operation cycle of about 2 years. At the end of each operation cycle,
-    one-quarter to one-third of its fuel assemblies (spent fuel) are removed.
-    Most of the nuclear industry's spent fuel is now stored in spent fuel water pools ``at-reactor (AR)'',
-    which means either within the reactor building or in an adjacent spent fuel building linked to the reactor
-    AR pools minimize the difficulty of transporting spent fuel.\cite{iaea99}
-    US has generated about $65,000$ metric tons of spent fuel, of which $75\%$ is stored in AR pools.\cite{a11}
-    by a transfer tunnel. 
-    Spent fuel pools  are usually about 15-meter deep borated water pools, with storage rack at the bottom to hold the fuel assemblies.
-    The borated water contains boron which are more prone to neutron captures.
-    The water absorbs the heat and shields the radiation. Water is the chosen medium because it is cheap and
-    can provide cooling by circulation. Moreover, water can serve as a shielding from radiation, because it has a
-    fairly high density and hence forces the radiated particles (especially the neutrons) to interact with
-    the water molecules and lose energy. Water temperature is maintained
-    below 50 $^\circ$C via circulation and cooling. The integrity of the storage can persist well beyond 
-    50 years.\cite{a11, iaea99}
+    Short-term cooling of spent fuel is necessary before further disposal or reprocessing can be done.
+    Radiation shielding and cooling are the two important aspects of spent fuel pools.
+
+    Spent fuel pools  are usually about 15-meter deep water pools,
+    with storage rack at the bottom to hold the fuel assemblies. The large depth makes sure that the radiations
+    are well attenuated by the time they reach the surface (see Eq.\ref{eq:atten}).
+
+    The water is enriched with boron atoms that have neutron capture cross section of about $1000$ times
+    the cross section of water molecule. Hence the boron atoms captures the neutrons in the spent fuel
+    and prevent them from causing further fissions and the formation of transuranic elements.
+
+    On the other hand, the water in spent fuel pool is circulated such that the heat transfer from the spent
+    fuel rods to the water via convective heat transfer. The convective heat transfer rate in this case is
+    extremely complicated and we should not try to derive it here. However, we can still gain useful insights
+    from a crude approximation with the Newton's cooling law via convective transfer
+    \begin{equation}
+        \label{eq:cooling}
+        \frac{dQ}{dt} \propto h\cdot (T_{sf} - T_{water}),
+    \end{equation}
+    where $Q$ denotes the heat transferred from the spent fuel to the water,
+    $h>0$ the heat transfer coefficient (which is assumed to be independent
+    of the temperature), $T_{sf}$ the spent fuel temperature, and $T_{water}$ the temperature
+    of the water.
+    $h$ depends of the fluid dynamics properties of the water and can be measured experimentally. The goal of
+    cooling is to prevent heat from accumulating in the spent fuel and thus $dQ/dt \geq 0$ at all time.
+    Apply the first and third laws of thermodynamics to the spent fuel rod,
+    \begin{equation}
+        \label{eq:first}
+        \frac{-\Delta Q}{\Delta T_{sf}} = dS > 0,
+    \end{equation}
+    where $Q\geq 0$ is consistent with the definition above and $S$ denotes entropy, we conclude that $Q$ and $T_{sf}$ always vary in the opposite direction. Rearrange Eq.~\ref{eq:cooling} to get
+    \begin{equation}
+        T_{water} \propto h.T_{sf} - \frac{dQ}{dt}.
+    \end{equation}
+    Hence if we let $dQ/dt=0$ (its min) and $T_{sf}$ be the temperature of the spent
+    fuel rods when it enters the pool (its max), we
+    can obtain an upper bound on $T_{water}$, the temperature of the water. This temperature turns out to
+    be about $50^{\circ}$C to transfer the heat generated by leftover fission in fresh spent fuel. After a
+    certain amount of time, as the fission slows down, $dQ/dt$ decreases and hence $T_sf$ decreases: the spent
+    fuel is cooled.
 
@@
     reprocessed plutonium-239 is the primary fissile element in the production of nuclear weapons.\cite{aa12}
-    On the other hand, France (the world leader in reprocessing technology), India, Japan, Russia, the UK
+    On the other hand, France (the world leader in reprocessing technology), India, Russia, the UK
     and China (since 2010) continues to reprocess spent fuel. To understand the details of reprocessing
@@
     between the heavy elements (uranium and plutonium) from the other lighter fission products.
-    The process involves three steps: (a) adding appropriate solvent (usually nitric acid) to absorb and oxidize 
-    plutonium and uranium while leaving out other elements;
-    (b) adding a reducing (de-oxidizing) agent that reduces (de-oxidizes) only the
-    oxidation state of plutonium to one that is insoluble
-    and thus separable; and finally (c) extracting the plutonium from the mixture. In this way
+    In this way,
     the oxidized plutonium and uranium are separated and can be reduced back to reusable radionuclides.\cite{lb01} 
@@
     uranium extraction) technique extracts the uranium but leaves the plutonium mixed with other
-    radioactive materials. Another popular alternative, pyrochemical reprocessing, inserts giant electrodes
-    into the chemical bath of nuclear waste and electroplates the plutonium and other transuranic elements onto
-    the electrode. However, none of these alternatives have been put into practice due to the high construction
-    and operating costs of reprocessing facilities.\cite{aa12}
-
-    %\red{fast neutron reactors}
+    radioactive materials.\cite{aa12}
 
@@
     The time frame of nuclear waste disposal is about $10^4-10^6$ years, after which the radioactivity
-    of the waste is predicted to decline to a safe level (see Sec.\ref{sec:spentfuel}).
+    of the waste is predicted to decline to a safe level (see Sec.\ref{sec:waste:spentfuel}).
     Deep geological repositories that are stable over geological timescales naturally offer a good option for disposal.
@@
     In this section, we first use the Yucca Mountain Nuclear Waste
-    Repository as a case study to understand the requirements for a permanent HLW disposal site and the debate
+    Repository (see Fig.~\ref{fig:yucca} )
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=\textwidth]{yuccaRepo.jpg}
+            \caption{Scheme of the 300-meter deep Yucca Mountain Waste Repository.\cite{yucpic}}
+            \label{fig:yucca}
+        \end{figure}
+    as a case study to understand the requirements for a permanent HLW disposal site and the debate
     surrounding it.
@@
         located near the Nevada Test Site in Nevada. The project has been highly contested since its
-        inception and the Obama administration ended federal funding for the project in 2011 for political,
+        inception and the Obama administration ended federal funding for the project in 2011, citing political,
         not safety or technical, reasons. We will take a glimpse at the suitability assessment of the site
         as a nuclear waste repository to get a sense of its complexity.
-        \subsubsection{Geological stability of Yucca Mountain}
+
+        \subsubsection{Geological stability of Yucca Mountain\footnote{This section is a short summary of \cite{m06}}}
         The chief concern of the suitability of Yucca Mountain was about water. If moist air comes to
@@
         releasing radionuclides.
-        The community seemed to agree that the cladding will have to carry the primary
+        The community seemed to agree that the packaging will have to carry the primary
         responsibility of preventing the oxidation of the waste, but predicting the performance of some
@@
         experience. Over time, the radiation from the waste will also alter the rock (by altering its composition
-        and creating fractures) and the water in the rock.\cite{m06}
+        and creating fractures) and hence the water percolation rate.
 
@@
         Volcanic activity, meteorite hit and earthquakes may seriously undermine the integrity of the repository.
-        Hot upwelling water from below the repository may corrode the cladding and release radionuclides into
+        Hot upwelling water from below the repository may corrode the packaging and release radionuclides into
         the ground water system. It is impossible to address all of these concerns in this work, but we will look
         at how the risk assessment is carried out in the case of volcanism.
+
         \subsubsection{Total System Performance Assessment (TSPA)}
@@
         (for example if new parameters and factors are included), the results from TSPA will not change
-        significantly.e\cite{cv14,ocrwm02}
+        significantly.\cite{cv14,ocrwm02}
 
@@
         fifteen million and eight million years ago. However, it is the small-volume eruptions, the last of
-        which occurred about $75,000$ years ago, that would pose risk to the repository in its life span.
+        which occurred about $75,000$ years ago, that would the most likely pose risk to the repository
+        in its life span of a few million years.
         The central question is whether the risk of volcanic eruption is acceptable. Mathematically,
@@
         if the probability of a volcanic eruption seriously damaging the repository is less than
-        one in one hundred million per year, it is acceptable.\cite{epa93}
+        one in ten million per year, it is acceptable.\cite{epa93}
         Of course, one has to also consider consequences of a potential radiological releases if an eruption
-        occur. Here we focus on the estimation of occurrence probability.
+        occurs. Here we focus on the estimation of occurrence probability.
 
-        The probability that a future volcanic eruption disrupting a repository is given by
+        The probability that a future volcanic eruption disrupting a repository over
+        the repository's lifetime is given by
         \begin{equation}
@@
         $P(A|B)$ gives the probability of event $A$ happening given that event $B$ happened;
-        $D$ denote the disruption event,
+        $D$ denotes the disruption event,
         $E2$ the event that the eruption affects the repository, $E1$ the event that an eruption occurs.
@@
             \caption{A flowchart for the volcanism risk estimation problem. The oval-shaped
-            parts are all probabilistic. \ref{sec:volcano} \cite{cv14}}
+            boxes are probabilistic models, hexagon shaped the TSPA computational model, and
+        square human decisions.\cite{cv14}}
             \label{fig:volcanos}
@@
         $(x_1,x_2,...,x_n)$.
-        However, if the uncertainty of the input parameter if expressed as a distribution, or if $y$ is
+        However, if the uncertainty of the input parameter is expressed as a distribution, or if $y$ is
         not an analytic function, then the propagation of uncertainty method fails. In our volcano disruption
@@
         more data and hence improved statistics, but longer time interval dated far back may fail to represent
-        future event rates since the geology evolves. The best approach seems to be choosing time interval
+        future event rates since the geology evolves. The best approach seems to be choosing the time interval
         based on geological cycles.
@@
         Given a probability distribution for $E1$, we can estimate $P(E2|E1)$ -- the probability of
-        a given event intersecting the Yucca Mountain site. It is simply the ratio of the area of
+        a given event intersecting the Yucca Mountain site per year assuming that it is independent
+        of time. It is simply the ratio of the area of
         the repository to the entire volcanic zone of a given event. Since the area of the repository
@@
         to predict the size of the volcanic zone is not well determined.
-        Different models predict $P(E2|E1)$ ranging from $0.015$ to $0.0008$, which again has a high
+        Different models predict $P(E2|E1)$ ranging from $0.015$ to $0.0008$ per year, which again has a high
         uncertainty.
@@
         are combined through a Monte-Carlo simulation to produce the disruption probability/rate $P(D)$
-        at the Yucca Mountain repository. DOE estimates a mean disruption rate of $1.5\times 10^8$ with
-        a 90 percent confidence interval of $5.4\times 10^{-9}-4.9\times 10^{-8}$. Hence the DOE's result
+        at the Yucca Mountain repository. DOE estimates a mean disruption rate of $1.5\times 10^8$ events
+        per year with
+        a 90 percent confidence interval of $5.4\times 10^{-9}-4.9\times 10^{-8}$ events per year. Hence the DOE's result
         states that most likely the disruption won't happen more than once every 20 million years.
@@
         Both the NRC and the DOE conclude with a very low volcanic disruption rate that is below the EPA
-        threshold. Nevertheless, the consequences of such a disruption event is still not well understood.
+        $10^{-7}$ events per year threshold of negligible risk.
+        Nevertheless, the consequences of such a disruption event is still not well understood.
         If the leaked dose is low in such an event, then the risk is negligible. If the leaked dose is high,
@@
         Deep borehole disposal utilizes deep and narrow boreholes to dispose of HLW. The boreholes
-        will place the cladded waste as much as 5 kilometers under the surface of the ground and seal it
+        will place the packaged waste as much as 5 kilometers under the surface of the ground and seal it
         with concrete clay.
-        The engineered clad, along with the granite rock with few cracks on which the borehole is drilled,
+        The engineered packaging, along with the granite rock with few cracks on which the borehole is drilled,
         provides
@@
 
-        %\subsection{reversible disposal}
-
 \section{Discussions}
-Nuclear waste management and disposal certainly is becoming a more and more pressing issue.
-However, given the technical and political complexity involved in assessing the suitability
-of a permanent repository site,\cite{m14} it seems like dry cask storage should be promoted extensively
-before a permanent repository is ready. Large-scale dry cask storage would ease the pressure on the
-spent fuel pool while easing the pressure on DOE and NRC to rush the scientific analyses.
-But still, we have yet to know the durability of dry casks.
+Nuclear waste management and disposal is certainly becoming an increasingly pressing issue.
+We explore the creation, management and disposal of highly radioactive nuclear waste and the physical
+princples underlying those processes. The increasingly overpacked spent fuel pools and the challenges
+associated with geological repositories are the weak links in this chain.
 
 In terms of scientific analyses for a geological repository, much has yet to be done.
-On the performance analysis front, we need
+On the performance analysis end, we need
 to think about how to model without enough data and how to deal with orders of magnitude uncertainty.
@@
 A fast neutron reactor uses neutrons of higher energy to induce the fission of plutonium-239
-such that the spent nuclear fuel would contain very little plutonium-239. Reprocessing and
-vitrification techniques can significantly reduce the volume of the waste and make it more
+such that the spent nuclear fuel would contain very little plutonium-239. Reprocessing 
+techniques can significantly reduce the volume of the waste and make it more
 manageable. With reprocessing, deep boreholes storage may become a more feasible disposal solution 

### 2017-03-15 paper.tex
end-of-term update

@@
 \begin{abstract}
-    As nuclears around the world continuously generate highly radioactive waste,
-    how we manage nuclear waste is becoming more important. In this work,
+    As nuclear reactors around the world continuously generate highly radioactive waste,
+    how we manage nuclear waste is becoming more pertinent. In this work,
     we examine a few scientific and technical issues related to nuclear waste management and disposal.
@@
 requires extremely careful disposal methods.
-Nevertheless, there is currently no permanent repository in the world to dispose of these highly radioactive
+Moreover, there is currently no permanent repository in the world to dispose of these highly radioactive
 waste.\footnote{The Waste Isolation Pilot Plant (WIPP) in New Mexico is designed to be a permanent radioactive
     waste repository, but only
-for waste classified as Transuranic, which, by the Nuclear Regulatory Commission standard, does not have particularly
+for waste classified as transuranic, which, by the Nuclear Regulatory Commission standard, does not have particularly
 high radioactivity but only long half-life (See Sec.~\ref{sec:waste:classification}). In particular, WIPP does not accept spent nuclear
@@
 
-Evidently, nuclear waste management and disposal is becoming a increasingly urgent matter globally.
+Evidently, nuclear waste management and disposal are becoming increasingly urgent matters globally.
 In light of the high stake,
@@
 In this work, we will attempt to examine some of the scientific issues underlying nuclear waste management
-and disposal. We will start by building up some
-language and knowledge necessary to understand how nuclear waste is generated (Section~\ref{sec:phys}),
-then apply those principles to study
-the physical properties of nuclear waste (Section~\ref{sec:waste}), its management and temporary storage
-(Section~\ref{sec:temp}), and the debates around its eventual disposal (Section~\ref{sec:disposal}).
+and disposal. We will start by introducing radiations and its biological effects (Section~\ref{sec:phys}); then
+we examine the sources and the physical properties of nuclear waste (Section~\ref{sec:waste}),
+its management and temporary storage (Section~\ref{sec:temp}), and the debates around its eventual disposal (Section~\ref{sec:disposal}).
 
@@
 \label{sec:phys}
-    Here we detail the nuclear physics underlying most of the discussions
-    in this work. We will apply these results in our discussion of nuclear wastes.
+The principles of radioactivity and its biological effects play a central role in our
+discussion of nuclear waste. Hence, we should first review and extend our understanding of 
+radioactivity and radiations.
     \subsection{Radioactivity\footnote{
@@
     Radioactive decay is the process in which most of the known detrimental radiations are generated.
-    Here we review the physical principles and mathematical formalisms behind it.
+    Here we review the physical principles and mathematical formalism behind it.
         \subsubsection{Quantum theory of radioactive decays}
         The law of radioactive decay is a direct consequence of quantum mechanics.
+        Consider a nucleus going through radioactive decay.
         We can approximate the nucleus as a particle situated in a potential well with potential given by
@@
         where $V$ is the time-independent nucleus potential that gives rise to eigenstates $\ket{\psi_1}$
-        and $\ket{\psi_2}$ with $\ket{\psi_2}$ having a lower (expectation value of) energy; $V_p$ is a small
+        and $\ket{\psi_2}$ with $\ket{\psi_2}$ having a lower (expected) energy; $V_p$ is a small
         perturbation that causes the decay.
@@
         We can easily solve the differential equation given by Eq.~\ref{eq:lambda} to get
-        the the number of nuclei as a function of time $t$
+        the number of nuclei as a function of time $t$
         \begin{equation}
@@
         The radioactive isotope used for Computer Tomography (CT) scan has a specific activity
-        of $\sim 10^{8}$ Bq/g. See Table~\ref{tab:hl} for the half-lives and specific activities
+        of $\sim 0.01$ Ci/g. See Table~\ref{tab:hl} for the half-lives and specific activities
         of some isotopes present in highly radioactive nuclear waste.
@@
             \centering
-            \caption{Specific activity and hal-life of isotopes commonly present in spent nuclear fuel \cite{toi}}
+            \caption{Specific activity and half-life of isotopes commonly present in spent nuclear fuel \cite{toi}}
             \begin{ruledtabular}
@@
         There are three main types of radioactive decay processes that generate radiations in nuclear waste: 
-        Alpha decay, which generates alpha particles (helium-4 nuclei), Beta decay which generates beta
+        Alpha decay, which generates alpha particles (helium-4 nuclei), beta minus decay which generates beta
         particles (electrons) and gamma decay which generates gamma rays.
         The type of radiation (particle beam) generated
-        have different properties and deserve separate trearments when we discuss health effects and shielding.
+        have different properties and deserve separate treatments when we discuss health effects and shielding.
         \paragraph{Alpha decay}
@@
         particle and decays into a different element. The emission of an alpha particle is
-        particularly favorable because it does not change the parity of the nucleus and has a particularly
+        particularly favorable because it has a particularly
         high binding energy. The process can be represented by
         \begin{equation}
-            {}^A_Z X_N \rightarrow {}^{A-4}_{Z-2}X'_{N-2} + \alpha.
-        \end{equation}
-
-        \paragraph{Beta decay}
+            {}^A_Z X_N \rightarrow {}^{A-4}_{Z-2}X'_{N-2} + \alpha,
+        \end{equation}
+        where $Z$ denotes the atomic number (number of protons) of a nucleus $A$ the atomic mass, and $N$ the
+        neutron count.
+
+        \paragraph{Beta minus decay}
         There are a variety of beta decay processes. Here we only describe the negative beta decay that is
@@
         \cite[chapt. 1.5]{l01}}}
-        The resultant nuclei (called ``daughter products'') from a radioactive decay are oftern radioactive
+        The resultant nuclei (called ``daughter products'') from a radioactive decay are often radioactive
         themselves and their decay rate can be larger or smaller than their parents. Consider a decay chain
@@
         The decay is essentially the decay of $A$ directly into $C$. Hence, if we have a chain of reaction
-        where each daughter nuclei has a half-life much shorter than that of its parent, then 
+        where each daughter nucleus has a half-life much shorter than that of its parent, then 
         isotope $A$ effectively decays into the end of the decay chain and releases a lot more radiations than
@@
         \paragraph{Alpha particles} An alpha particle loses energy predominantly via the interactions between its electric
-        field and the electrons, in the matter it travels through. Interactions with nuclei and head-on collisions
-        are quite rare. Since the speed of an alpha particle is low, it deposits more energy into the matter it
-        interacts with. In tissues, $1$-MeV alpha particles travel through only a few tens of micrometers and thus
+        field and the electrons in the matter it travels through. Interactions with nuclei and head-on collisions
+        are quite rare. Since the speed of an alpha particle is low, it deposits more energy as it
+        travels. In tissues, $1$-MeV alpha particles travel through only a few tens of micrometers and thus
         are easily stopped by skin. However, if alpha particles are ingested and more sensitive organs are irradiated,
@@
         Finally, electrons also radiate photons
-        via bremstrahlung radiations as they deccelerate. Therefore, the electrons' energy end up distributing
+        via bremsstrahlung radiations as they decelerate. Therefore, the electrons' energy end up distributing
         across a much larger volume.
         \paragraph{Photons} Gamma rays interact with matter via photoelectric effects, Compton scattering and pair productions
-        depending on th energy of the photon and the atoms it interact with.
+        depending on the energy of the photon and the atoms it interact with.
         Since biological tissues mainly consist of elements of low atomic number, Compton
         scattering is the most important interaction. The scattered photons may interact further with the tissues
-        via photoelectric effects and secondary Compton scattering. The intensity of the energy deposition falls
-        off much more slowly and is hence more penetrating. Both electrons and photons can penetrate human body easily,
-        although they do not deposit as much energy, they generally are more dangerous than alpha particles due to their
+        via photoelectric effects and secondary Compton scattering. As a result, the intensity of the energy deposition falls
+        off much more slowly and is hence more penetrating. Both electrons and photons can penetrate human body easily.
+        Although they do not deposit as much energy as alpha particles, they are generally more dangerous due to their
         penetration abilities.
@@
         The energy deposition described above is an important cause of irreversible damage to sensitive molecules
-        like the Deoxyribonucleic (DNA) molecules if the energ transferred is large enough.
+        like the Deoxyribonucleic (DNA) molecules if the energy transferred is large enough.
         In the cases where the energy transferred is not as large, the radiation ionizes simpler molecules like water
-        to form chemically active free radicals. The free radicals have unpaired electrons. Hence they go on to react and distrupt
+        to form chemically active free radicals. The free radicals have unpaired electrons. Hence they go on to react
+        and disrupt
         other molecules that may be crucial to the biological system.
@@
         of these quantities: the absorbed dose, the equivalent dose and the effective dose.
-        The SI units for dose is siever (Sv), which is equivalent to J/Kg, the amount of energy deposited into
-        per unit mass of living tissue. Some work uses the unit gray (Gy) for abosrbed dose in medical context,
+        The SI units for dose is sievert (Sv), which is equivalent to J/Kg, the amount of energy deposited into
+        per unit mass of living tissue. Some work uses the unit gray (Gy) for absorbed dose in medical context,
         but it is the same as sievert.
@@
         The approximate linear dependence on $E_{\gamma}$ is valid from less than $0.1$ MeV to several
-        MeV and therefore should be good for all sources but those generated in an accelerator.\cite{my68}
-        The inverse-squared dependence on distance is hardly surprising.
+        MeV and therefore should be good for all sources we are considering.\cite{my68}
+        The inverse-squared dependence comes from the isotropic radiation assumption.
 
@@
         If the radiation level is high, the radiation may generate enough free radicals to disrupt the cells or directly
-        ionize critical molecules. Some examples are gastrointestinal damage and impairment
+        ionize critical molecules. Some examples of these short-term effects are gastrointestinal damage and impairment
         of fertility.\cite{u16}
@@
         the chance of occurrence depends on the dose.\cite{u16}
-        These effects mainly result from the minor disruptions of DNA molecules since the disruptions are replicated
+        These effects mainly result from the minor disruptions of DNA molecules from the free radicals generated
+        by low doses of radiation. The disruptions are replicated
         as cells reproduce.
@@
         The mechanism is similar to that of delayed health effects, since DNA molecules are passed on
-        to offpsrings.
-        UNSCEAR estimates that risk for severe hereditary effect is about $0.3-0.5$ percent per siever to
+        to offspring.
+        UNSCEAR estimates that risk for severe hereditary effect is about $0.3-0.5$ percent per sievert to
         the first generation following the radiation exposure.\cite{u16, unscear16}
@@
     However, their definitions of high-level waste (HLW), the most detrimental kind, are relatively consistent.
-    The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the activity
-    content and the half-life of the radionuclides.
-    Activity content is a measure of how much radioactivity $\mathcal{A}(t)$, as defined in Eq.~\ref{eq:act},
-    a given amount of waste contains.
-    A form of waste may possesses a high activity content due to
-    a high concentration of radionuclides and/or the presence of radionuclides with high activity.\cite{iaea09}
+    The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the averaged specific activity
+    and the half-life of the radionuclides in the waste.\cite{iaea09}
     Each form of waste, depending on the half-life, activity content, and temperature, requires different
@@
     heat as it radiates. The typical source of HLW is spent fuel from the nuclear reactor.
-    In fact, since there is no commercial reprocessing facilities in the US (see section~\ref{sec:reproc}),
+    In fact, since there is no commercial reprocessing facilities in the U.S. (see section~\ref{sec:reproc}),
     almost all of the HLW is in the form of spent nuclear fuel from the reactor.
@@
     or larger and concentration of activity larger than $3700$ Bq/g.\cite{j83,s01} The activity
-    of TRU is usually low but it remains active for long period of time.
+    of TRU is usually low but it remains active for long period of time. Moreover, TRU can be easily turned
+    into plutonium-239, the {\it de facto} material to fabricate atomic bombs.
 
@@
     as intermediate-level waste (ILW). ILW are solidified in concrete and buried in shallow
-    repositories.\cite{s01} However, US regulations do not make the distinction
+    repositories.\cite{s01} However, U.S. regulations do not make the distinction
     between LLW and ILW but different disposal methods are practiced.\cite{nrc09, s01}
@@
     they are not usually classified as LLW.
-    Over geological timescales, the stable uranium-238 decays into radioactive thorium and radium 
-    These radionuclides remain after the uraninium (all of its isotopes)
+    Over geological timescales, the stable uranium-238 in the ore decays into radioactive thorium and radium 
+    These radionuclides remain after the uranium (all of its isotopes)
     has been extracted from the ore. Notably, radium
@@
 
-    \subsection{Creation of spent fuel}
+    \subsection{Creation and properties of spent fuel}
     \label{sec:waste:spentfuel}
     \subsubsection{Nuclear fuel}
+    Spent fuel comes from used nuclear fuel.
     For most of the commercial nuclear reactors, the energy comes from the isotope uranium-235 which
@@
     via isotope separation to increase the uranium-235 content. Typical fresh nuclear fuel contains $3\%$ uranium-235
-    and $97\%$ uranium-238.
+    and $97\%$ uranium-238 and are in the form of nuclear fuel rods.
+
     \subsubsection{Neutron absorption and radiative capture}
@@
     is the primary mode of interaction between the nuclei and the neutrons.
-    Neutron absorption is the process in which a nucleus absortb a neutron to form an unstable
+    Neutron absorption is the process in which a nucleus absorbs a neutron to form an unstable
     nucleus.
@@
     \end{align}
-    The uranium-233 formed has a half-life of $1.4$ million years and hence are a lot more dangerous than
+    The uranium-233 formed has a half-life of $1.4$ million years and is therefore a lot more dangerous than
     thorium-232. Another important radiative capture reaction happening inside nuclear fuel is the radiative
-    capture of uranium-238 (the most abundunt isotope in nuclear fuel) to form plutonium-239, the {\it de facto}
-    isotope to construct atomic bombs
+    capture of uranium-238 (the most abundant isotope in nuclear fuel) to form plutonium-239.
     \begin{equation}
@@
         On average, inside a nuclear reactor, $2.5$ neutrons are produced per fission. These neutrons
-        continue to induce fissions of other uranium-235 nuclei, causing a ``chain reaction'' of subsequent
-        fissions that generate a huge amount of energy.\cite[chapt. 10]{l01}
+        continue to induce fission of other uranium-235 nuclei, causing a ``chain reaction'' of subsequent
+        fission that generate a huge amount of energy.\cite[chapt. 10]{l01}
         \begin{figure}[h]
@@
         We can estimate the amount of energy released in fission by considering the binding energy difference
-        between the parent nuclei and the fission products. As a consequence of the masss-energy equivalence in the theory
+        between the parent nuclei and the fission products. As a consequence of the mass-energy equivalence in the theory
         of special relativity, we can calculate the energy released $E$ from the atomic mass differences between the parent $m_{parent}$
@@
         $0.9$ MeV u${}^{-1}$, which amount to $212$ MeV for a uranium-235 atom.
-        Apart from $6.5\%$ of th energy carried by the neutrinos which does not interact with the surrounding,
+        Apart from $6.5\%$ of the energy carried by the neutrinos which does not interact with the surrounding,
         the rest of the energy can be utilized by virtue of outgoing particles depositing energy in the medium.
@@
         In addition, for spent fuel freshly removed from the reactor, there is still a small amount ($\sim 1\%$)
-        of uranium-235 left along with a lot of free neutrons released from previous fissions. Thus, a large amount
+        of uranium-235 left along with a lot of free neutrons released from previous fission. Thus, a large amount
         of energy is generated via fission of the leftover uranium-235 and heats up the spent fuel rods.
@@
         and produce radionuclides that are
-        heavier than uranium (``transuranic'') such as plutonium, neptunium and americum.
+        heavier than uranium (``transuranic'') such as plutonium, neptunium and americium.
         These transuranic nuclei have relatively low activity but their half-lives are from a few thousand years
@@
             \includegraphics[width=0.6\textwidth]{csur.png}
-            \caption{Activty of $1$ gram of cesium-137 (half-life $30$ years) and uranium-233 (half-life $1.3$ million
+            \caption{Activity of $1$ gram of cesium-137 (half-life $30$ years) and uranium-233 (half-life $1.3$ million
             years) as a function of time in years. The activity of cesium-137 falls off to virtually zero after a few
@@
     it contains. Freshly removed spent fuel rods are stored in spent fuel pools for at least $3$ years before
-    the next step. Reprocessing can extract the long-lived plutonium and uranium isotopes and
+    the next step. After that, reprocessing can extract the long-lived plutonium and uranium isotopes and
     convert the waste into more manageable form (vitrification). However, reprocessing of commercial
-    nuclear spent fuel did not happen in the US for cost and proliferation concerns. Instead, dry-cask
-    storage, in addition to spent fuel pools, is used as an interim storage methods.
+    nuclear spent fuel did not happen in the U.S. for cost and proliferation concerns. Instead, dry-cask
+    storage, in addition to spent fuel pools, is used as an interim storage method.
     \subsection{Radiation shielding}
@@
     are water and concrete; and the mode of interactions of radiations with those mediums are similar to
-    the the physical intereaction between radiations and biological molecules (the most abundant of which is
+    the physical interaction between radiations and biological molecules (the most abundant of which is
     water) described in Sec.~\ref{sec:interactions}.
@@
     a beam of particles with intensity $I_i$ (with unit (area$\cdot$time)${}^{-1}$) passing through a medium
-    with density of molecules $N_s$. Suppose on average $R_i$ interactions occur per time unit.
-    The cross section of the iteraction between the shielding medium defined as
+    with number density of molecules $N_s$. Suppose on average $R_i$ interactions occur per time unit.
+    The cross section of the interaction between the shielding medium defined as
     \begin{equation}
@@
     \begin{equation}
-        \mu = \sigma N_s
+        \mu = \sigma N_s,
     \end{equation}
@@
     the cross section of water molecule. Hence the boron atoms captures the neutrons in the spent fuel
-    and prevent them from causing further fissions and the formation of transuranic elements.
-
-    On the other hand, the water in spent fuel pool is circulated such that the heat transfer from the spent
+    and prevent them from causing further fission and the formation of transuranic elements.
+
+    On the other hand, the water in spent fuel pool is circulated such that the heat transfers from the spent
     fuel rods to the water via convective heat transfer. The convective heat transfer rate in this case is
@@
     of the water.
-    $h$ depends of the fluid dynamics properties of the water and can be measured experimentally. The goal of
+    The heat transfer coefficient $h$ depends of the fluid dynamics properties of the water and can
+    be measured experimentally. The goal of
     cooling is to prevent heat from accumulating in the spent fuel and thus $dQ/dt \geq 0$ at all time.
@@
     be about $50^{\circ}$C to transfer the heat generated by leftover fission in fresh spent fuel. After a
-    certain amount of time, as the fission slows down, $dQ/dt$ decreases and hence $T_sf$ decreases: the spent
+    certain amount of time, as the fission slows down, $dQ/dt$ decreases and hence $T_{sf}$ decreases: the spent
     fuel is cooled.
-
-    \subsection{Dry cask storage}
-    Spent fuel pools have very limited capacities. 
-    Overground dry cask storage provides a low-cost temporary storage before the spent fuel is disposed of.
-    After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks filled with inert gas. 
-The dry casks are often metal cylinders welded to prevent leakage. Each cask is surrounded by
-    additional steel and concrete to provide shielding.
-    The heat from the cask is removed by passive
-    cooling (air circulation). The NRC is hoping that the casks will
-    last safely for as long as 100 years, yet it has been observed that corrosions and cracking occur in 30 years
-    or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
-    overground dry cask storage does not seem to be a long-term solution. The doubts were expressed
-    during the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
-    \begin{displayquote}
-    Leaving the spent fuel on-site for extended periods of time was never intended and is not
-    responsible. (Dry cask manufacturer)
-
-    It is not ethical, basically to plan for long-term storage without pursuing a well defined repository
-    program. (Department of Energy employee speaking as a citizen)
-    \end{displayquote}
-    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in densely
-    packed pools
-    that are close to or over their limiting capacity.\cite{a11,aa12}
-    Therefore, there is clearly a need for a long-term disposal plan, which we will explore in
-    Sec.~\ref{sec:disposal}.
 
@@
     or fissile weapon material (plutonium-239). Therefore, reprocessing the spent fuel generates plutonium and
-    uranium that may be recycled. The US currently does not have any commercial nuclear fuel
+    uranium that may be recycled. The U.S. currently does not have any commercial nuclear fuel
     reprocessing facility due to the demonstrated high cost of reprocessing
@@
 
+    \subsection{Dry cask storage}
+    Spent fuel pools have very limited capacities. 
+    In the case where the spent fuel is not reprocessed, overground dry cask storage provides a low-cost temporary storage before the spent fuel is disposed of.
+    After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks filled with inert gas. 
+The dry casks are often metal cylinders welded to prevent leakage. Each cask is surrounded by
+    additional steel and concrete to provide shielding.
+    The heat from the cask is removed by passive
+    cooling (air circulation). The NRC is hoping that the casks will
+    last safely for as long as 100 years, yet it has been observed that corrosion and cracking occur in 30 years
+    or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
+    overground dry cask storage does not seem to be a long-term solution. The doubts were expressed
+    during the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
+    \begin{displayquote}
+    Leaving the spent fuel on-site for extended periods of time was never intended and is not
+    responsible. (Dry cask manufacturer)
+
+    It is not ethical, basically to plan for long-term storage without pursuing a well defined repository
+    program. (Department of Energy employee speaking as a citizen)
+    \end{displayquote}
+    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in densely
+    packed pools
+    that are close to or over their limiting capacity.\cite{a11,aa12}
+    Therefore, there is clearly a need for a long-term disposal plan, which we will explore in
+    Sec.~\ref{sec:disposal}.
+
 \section{Disposal of HLW}
@@
     In fact, it is accepted by the international science community that geological repositories in a stable environment over 300 meters below ground
-    is safer than indefinite overground storage of HLW, reprocessed or not.\cite{fmr11}
+    is safer than indefinite overground storage of HLW, reprocessed or not, as above-ground storage
+    are susceptible to catastrophic events like earthquakes and human-caused threats like terrorist attacks.\cite{fmr11}
     In this section, we first use the Yucca Mountain Nuclear Waste
-    Repository (see Fig.~\ref{fig:yucca} )
+    Repository     as a case study to understand the requirements for a permanent HLW disposal site.
+    Then we look at an alternative to a permanent deep geological solution:
+    deep boreholes storage.
         \begin{figure}[h]
@@
             \includegraphics[width=\textwidth]{yuccaRepo.jpg}
-            \caption{Scheme of the 300-meter deep Yucca Mountain Waste Repository.\cite{yucpic}}
+            \caption{Design of the 300-meter deep Yucca Mountain Waste Repository.\cite{yucpic}}
             \label{fig:yucca}
         \end{figure}
-    as a case study to understand the requirements for a permanent HLW disposal site and the debate
-    surrounding it.
-    Then we look at an alternative to a permanent deep geological solution:
-    deep boreholes storage.
     \subsection{Deep geological repository: Yucca Mountain}
@@
         of 1987 to be the nation's permanent deep underground repository for spent nuclear fuel. It is
-        located near the Nevada Test Site in Nevada. The project has been highly contested since its
+        located near the Nevada Nuclear Weapon Test Site in Nevada.
+        It is located 300-meter under the top of the Yucca mountain with impermeable layers of rocks
+        separating the repository
+        from both the surface and the ground water (see Fig.~\ref{fig:yucca}).
+
+        The project has been highly contested since its
         inception and the Obama administration ended federal funding for the project in 2011, citing political,
         not safety or technical, reasons. We will take a glimpse at the suitability assessment of the site
-        as a nuclear waste repository to get a sense of its complexity.
+        as a nuclear waste repository to get a sense of its technical complexity.
 
@@
         The information necessary to estimate the probability is listed in the flowchart Fig.~\ref{fig:volcanos}.
-        The boxes represent decision parameters that depends on the judgement of the decision makers and
+        The boxes represent decision parameters that depends on the judgment of the decision makers and
         the scientists. For example the ``Volcanic Zone Models'' should be picked by the scientists, and
@@
         Each oval is a complex computational model.
-        The large bolded ovals are the $E1$ and $E2$ parameters in Eq.\ref{eq:prob}.
+        The large bold ovals are the $E1$ and $E2$ parameters in Eq.\ref{eq:prob}.
         The $E2$ oval makes
@@
         \paragraph{The probability of a disruption} The probability distribution of $E1$ and $E2$
-        are combined through a Monte-Carlo simulation to produce the disruption probability/rate $P(D)$
+        are combined through a Monte-Carlo simulation to produce the disruption rate $P(D)$
         at the Yucca Mountain repository. DOE estimates a mean disruption rate of $1.5\times 10^8$ events
@@
 
-        The NRC emphasizes the importance of the bounding nature of the problem and uses only a subset
+        The NRC emphasizes the importance of the risk assessment nature of the problem and uses only a subset
         of the DOE model that reflect conservatism. It concludes a disruption rate of
@@
 We explore the creation, management and disposal of highly radioactive nuclear waste and the physical
-princples underlying those processes. The increasingly overpacked spent fuel pools and the challenges
+principles underlying those processes. The increasingly over-packed spent fuel pools and the challenges
 associated with geological repositories are the weak links in this chain.
@@
 
-Newer reprocessing methods (as mentioned in Sec.~\ref{sec:reproc}) are being proposed with
-fast neutron reactors as a solution to the proliferation issue with current reprocessing method.
-A fast neutron reactor uses neutrons of higher energy to induce the fission of plutonium-239
-such that the spent nuclear fuel would contain very little plutonium-239. Reprocessing 
-techniques can significantly reduce the volume of the waste and make it more
+Newer reprocessing methods (as mentioned in Sec.~\ref{sec:reproc}) are being proposed
+as  solutions to the proliferation issue from the current reprocessing method.
+ Reprocessing techniques can significantly reduce the volume of the waste and make it more
 manageable. With reprocessing, deep boreholes storage may become a more feasible disposal solution 

### 2017-03-26 paper.tex
Alex's comments

@@
-%\documentclass{article}
+%documentclass{article}
 \documentclass[nofootinbib,preprint,aps]{revtex4-1}
@@
 \begin{abstract}
-    As nuclear reactors around the world continuously generate highly radioactive waste,
-    how we manage nuclear waste is becoming more pertinent. In this work,
-    we examine a few scientific and technical issues related to nuclear waste management and disposal.
-    The principles of radioactivity and nuclear fission help us to understand the sources and properties
-    of different types of nuclear waste. Focusing on the highly radioactive spent fuel, the rest
-    of this work outlines the advantages and drawbacks of different spent fuel management and disposal
-    methods. We study the methods used for performance assessments for Yucca Mountain Nuclear Waste Repository,
-    especially in the context of volcanic disruption. Finally, we discuss some possible options for 
-    nuclear waste management and disposal after
-    the termination of the Yucca Mountain project.\\
-    Word count: 7449\\
+%    As nuclear reactors around the world continuously generate highly radioactive waste,
+%    how we manage nuclear waste is becoming more pertinent. In this work,
+%    we examine a few scientific and technical issues related to nuclear waste management and disposal.
+%    The principles of radioactivity and nuclear fission help us to understand the sources and properties
+%    of different types of nuclear waste. Focusing on the highly radioactive spent fuel, the rest
+%    of this work outlines the advantages and drawbacks of different spent fuel management and disposal
+%    methods. We study the methods used for performance assessments for Yucca Mountain Nuclear Waste Repository,
+%    especially in the context of volcanic disruption. Finally, we discuss some possible options for 
+%    nuclear waste management and disposal after
+%    the termination of the Yucca Mountain project.\\
+%    Word count: 7449\\
 \end{abstract}
@@
 of the U.S. government overseeing nuclear energy safety,
-estimates that a major fire in a U.S. spent fuel pool is expected to
+estimates that a major fire in a U.S. spent fuel pool would
 displace 3.4 million people in an area larger than New Jersey.\cite{s16}
@@
 In this work, we will attempt to examine some of the scientific issues underlying nuclear waste management
-and disposal. We will start by introducing radiations and its biological effects (Section~\ref{sec:phys}); then
+and disposal. We will start by introducing radiation and its biological effects (Section~\ref{sec:phys}); then
 we examine the sources and the physical properties of nuclear waste (Section~\ref{sec:waste}),
@@
 The principles of radioactivity and its biological effects play a central role in our
-discussion of nuclear waste. Hence, we should first review and extend our understanding of 
-radioactivity and radiations.
+discussions of nuclear waste. Hence, we should first review and extend our understanding of 
+radioactivity and radiation.
     \subsection{Radioactivity\footnote{
-    The derivations and in this section follows chapter $6$ in
+    The derivations in this section follow chapter $6$ in
     Krane's text \cite{k88} and chapter $1$ in Lilley's text \cite{l01}.
     }}
-    Radioactive decay is the process in which most of the known detrimental radiations are generated.
+    Radioactive decay is the process in which a nucleus decays from a higher energy state to a lower
+    energy state and releases energy in the form of radiation. Radioactivity accounts for
+    most of the processes that generate detrimental radiation in nuclear waste.
     Here we review the physical principles and mathematical formalism behind it.
@@
         \subsubsection{Types of radiation}
-        There are three main types of radioactive decay processes that generate radiations in nuclear waste: 
+        There are three main types of radioactive decay processes that generate radiation in nuclear waste: 
         Alpha decay, which generates alpha particles (helium-4 nuclei), beta minus decay which generates beta
         particles (electrons) and gamma decay which generates gamma rays.
-        The type of radiation (particle beam) generated
+        The types of radiation (particle beam) generated
         have different properties and deserve separate treatments when we discuss health effects and shielding.
         \paragraph{Alpha decay}
-        An alpha particle is a helium-4 atom ${}^4_2$He. In an alpha decay, a nucleus emits an alpha
+        An alpha particle is a positively-charged 
+        helium-4 nucleus ${}^4_2$He$^{2+}$. In an alpha decay, a nucleus emits an alpha
         particle and decays into a different element. The emission of an alpha particle is
@@
         an excited state.
-        We know that excited atom radiates photons to fall back into ground state. Similarly, excited nuclei
+        We know that excited atoms radiate photons to fall back into ground state. Similarly, excited nuclei
         decay to ground state via emitting photons, with energies in the range
         of $0.1$ to $10$ MeV, which are characteristic of the energy differences between nuclear energy states.
+        The energy range correponds to that of a gamma ray photon.
 
@@
         \end{equation}
-        Multiply both sides by $e^{\lambda_B t}$ and substitute in $N_A=N_0 e^{-\lambda_A} t$ to get
-        \begin{align*}
+        Multiply both sides by an integrating factor $e^{\lambda_B t}$ and substitute in the number of particle
+        $A$ as a function of time $N_A=N_0 e^{-\lambda_A t}$ to get
+        \begin{align}
             \frac{dN_B}{dt}e^{\lambda_B t} + \lambda_B N_B e^{\lambda_B t} &= \lambda_A N_A e^{\lambda_B},\nonumber\\
@@
                                 &= \frac{\lambda_A}{\lambda_B-\lambda_A}N_0 e^{(\lambda_B - \lambda_A)t}
-        \end{align*}
+        \end{align}
         Hence we end up with the number of $B$ nuclei as a function of time
@@
         where each daughter nucleus has a half-life much shorter than that of its parent, then 
-        isotope $A$ effectively decays into the end of the decay chain and releases a lot more radiations than
+        isotope $A$ effectively decays into the end of the decay chain and releases a lot more radiation than
         in a single decay. Moreover, before the decay chain reaches its equilibrium, the activity of the sample
-        may increase.
+        may even increase.
 
@@
         changes that lead to cancer and genetic defects. Here we briefly summarize the known effects.
-        \subsubsection{Interactions of radiations with living tissues}
+        \subsubsection{Interactions of radiation with living tissues}
         \label{sec:interactions}
-        Radiations causes damage to living tissues by depositing energy into the biological molecules.
+        Radiation causes damage to living tissues by depositing energy into biological molecules.
         Depending on the type of the radiation
-        and the molecules present in the tissues, the damage mechanism and effects are different.
-        \paragraph{Alpha particles} An alpha particle loses energy predominantly via the interactions between its electric
+        and the molecules present in the tissues, the damage mechanisms and effects are different.
+
+        An alpha particle loses energy predominantly via the interactions between its electric
         field and the electrons in the matter it travels through. Interactions with nuclei and head-on collisions
         are quite rare. Since the speed of an alpha particle is low, it deposits more energy as it
-        travels. In tissues, $1$-MeV alpha particles travel through only a few tens of micrometers and thus
+        travels. In living tissues, an $1$-MeV alpha particle travels through only a few tens of micrometers and thus
         are easily stopped by skin. However, if alpha particles are ingested and more sensitive organs are irradiated,
         serious damages can be done.
-        \paragraph{Beta radiation (elections)} Electrons are much lighter and hence are easily scattered.
+
+        Beta radiation consistens of electrons. Electrons are much lighter and hence are easily scattered.
         Moreover, they are much more penetrating since their energy lost per distance traveled is lower.
         Finally, electrons also radiate photons
-        via bremsstrahlung radiations as they decelerate. Therefore, the electrons' energy end up distributing
+        via bremsstrahlung radiation as they decelerate. Therefore, the electrons' energy end up distributing
         across a much larger volume.
-        \paragraph{Photons} Gamma rays interact with matter via photoelectric effects, Compton scattering and pair productions
+
+        Gamma rays interact with matter via photoelectric effects, Compton scattering and pair productions
         depending on the energy of the photon and the atoms it interact with.
@@
         and disrupt
-        other molecules that may be crucial to the biological system.
+        other molecules that may be crucial to the cells.
 
@@
         The biological impact of radiation depends on the type, power and energy of the radiation, the
-        exposure time, and the part of the body exposed. Various quantities have been introduced to
-        the biological impact of radiation. Here we review three
+        exposure time, and the part of the body exposed. Various quantities was developed to
+        describe the biological impacts of radiation. Here we review three
         of these quantities: the absorbed dose, the equivalent dose and the effective dose.
-        The SI units for dose is sievert (Sv), which is equivalent to J/Kg, the amount of energy deposited into
-        per unit mass of living tissue. Some work uses the unit gray (Gy) for absorbed dose in medical context,
-        but it is the same as sievert.
 
@@
         Absorbed dose~($D$) quantifies the 
-        radiation energy absorbed per unit mass.
+        radiation energy absorbed per unit mass. The SI unit for absorbed dose is gray (Gy), which is
+        equivalent to $J/kg$.
 
@@
         \begin{equation}
-            \frac{dD}{dt}(\mu\text{Sv h}^{-1}) \approx \frac{\mathcal{A}\text{(MBq)}\times E_{\gamma}\text{(MeV)}}
+            \frac{dD}{dt}(\mu\text{Gy h}^{-1}) \approx \frac{\mathcal{A}\text{(MBq)}\times E_{\gamma}\text{(MeV)}}
             {6\times [r\text{(m)}]^2}.
@@
 
-        \begin{table}[h]
+        \begin{table}
             \centering
@@
         \end{table}
-        However, absorbed dose does not take into account different biological implications of different types
-        of radiation.
-        For example, the same dose of alpha particle deposit more energy than gamma rays. The equivalent
-        dose ($H$) is defined as
+        However, the effects of the radiation on living tissue depend on the type of radiation.
+        For example, the same dose of alpha particle deposit more energy in living tissue than gamma rays.
+        To quantify the {\it effects} of radiation, we introduce equivalent
+        dose ($H$) as defined by 
         \begin{equation}
@@
         where $D_R$ is the absorbed dose and $w_R$ the weighting factor of a given kind of radiation $R$
-        (see Table~\ref{tab:eq} for a list of weighing factors).
+        (see Table~\ref{tab:eq} for a list of weighing factors). The unit for equivalent dose is sievert (Sv).
+        $1$ Sv is defined as the effects equivalent to $1$ Gy of Gamma radiation on living tissues. It describes
+        the effects rather than the physical dose received by living tissues.
 
@@
         where $w_T$ is the body tissue weighting factor, and $H_T$ the equivalent dose on the tissue
-        at question.
+        at question. Since effective dose quantifies the effects of radiation, the unit for effective dose
+        is also sievert.
         Some of the recommended values of tissue
@@
         whole body absorbing the radiation uniformly.
-        \begin{table}[h]
+        \begin{table}
             \centering
@@
         \subsubsection{Effects of radiation on humans}
-        Short-term effects of radiation are usually caused by extensive cell death or cell damage, resulting
-        from large-scale damages of critical molecules.\cite{u16}
+        Short-term effects of radiation usually involve extensive cell death or cell damage, resulting
+        from large-scale damage to critical molecules.\cite{u16}
         These effects are deterministic with a threshold,
@@
         not enough free radicals are created and hence the repairing capabilities of cells can undo the damages.
-        If the radiation level is high, the radiation may generate enough free radicals to disrupt the cells or directly
+        If the effective dose is high, the radiation may generate enough free radicals to disrupt the cells or directly
         ionize critical molecules. Some examples of these short-term effects are gastrointestinal damage and impairment
@@
         Delayed health effects occur a long time after the exposure and are usually stochastic. There is
-        no apparent threshold and the severity is usually independent of the dose but
+        no apparent threshold and the severity is usually independent of the dose, but
         the chance of occurrence depends on the dose.\cite{u16}
@@
     In this section, we will first look at a classification scheme of nuclear waste and then
-    examine in details the source, composition and properties of the most detrimental type, high-level waste.
+    examine the source, composition, and properties of the most detrimental type--the high-level waste.
     \subsection{Classification of nuclear waste}
@@
     Each form of waste, depending on the half-life, activity content, and temperature, requires different
-    disposal methods. The NRC classification scheme is as follows.\cite{nrc09}
+    disposal methods. The NRC classification scheme is as follows:\cite{nrc09}
 
@@
     sufficiently long half-life ($>20$ years) of elements with atomic number $92$ (atomic number of uranium)
-    or larger and concentration of activity larger than $3700$ Bq/g.\cite{j83,s01} The activity
+    or larger and concentration of activity above $3700$ Bq/g.\cite{j83,s01} The activity
     of TRU is usually low but it remains active for long period of time. Moreover, TRU can be easily turned
@@
     higher radioactivity such that they are classified
-    as intermediate-level waste (ILW). ILW are solidified in concrete and buried in shallow
+    as intermediate-level waste (ILW). ILW are consolidated in concrete and buried in shallow
     repositories.\cite{s01} However, U.S. regulations do not make the distinction
@@
     they are not usually classified as LLW.
-    Over geological timescales, the stable uranium-238 in the ore decays into radioactive thorium and radium 
+    Over geological timescales, the stable isotope uranium-238 
+    in the ore decays into radioactive thorium and radium 
     These radionuclides remain after the uranium (all of its isotopes)
@@
     can decay into radon gas which is an alpha-emitter. Hence, the alpha radiation may do additional
-    damage to body if the radon enters the body as one breathes.
-    The uranium tailings leave the uranium mills in the form of a liquid sludge
-    and can dry. Therefore, it is necessary to prevent the uranium tailings
+    damage to body if the radon gas enters the body as one breathes.
+    Therefore, it is necessary to prevent the uranium tailings
     from contaminating water and air.
 
-    Since all categories but high-level waste are relatively easy to handle, we deal with the formation and
-    properties of spent nuclear fuel, the dominant contribution to high level waste, in the rest of this section.
+    Since all categories but high-level waste are relatively easy to handle, in the rest of
+    this section we will focus on the formation and
+    properties of spent nuclear fuel, the dominant contribution to high level waste,
 
@@
     \label{sec:waste:spentfuel}
+    Spent fuel comes from used nuclear fuel. Here we look at the source and properties of spent fuel.
     \subsubsection{Nuclear fuel}
-    Spent fuel comes from used nuclear fuel.
-    For most of the commercial nuclear reactors, the energy comes from the isotope uranium-235 which
-    capable of undergoing fission (see Sec.~\ref{sec:fission}). However, natural uranium consists of $99.3\%$
+    For most of the commercial nuclear reactors, the energy comes from the isotope uranium-235 capable
+    of undergoing fission.
+    However, natural uranium consists of $99.3\%$
     non-fissile uranium-238 and only $0.7\%$ uranium-235. Therefore, for nuclear fuel, natural uranium is enriched
-    via isotope separation to increase the uranium-235 content. Typical fresh nuclear fuel contains $3\%$ uranium-235
-    and $97\%$ uranium-238 and are in the form of nuclear fuel rods.
+    via isotope separation to increase the uranium-235 content. Typical fresh nuclear fuel, usually packaged in
+    rods, contains $3\%$ uranium-235 and $97\%$ uranium-238.
 
@@
     occurs in the reactor. The neutrons come from a startup neutron source and subsequently from the fission
-    process (see Sec.~\ref{sec:fission}).
+    process (see Sec.~\ref{sec:fission} for more details).
     Since a neutron does not carry any charge, it does not experience the Coulomb repulsion potential when
@@
     state, neutron absorption is classified into radiative capture if the nucleus radiates to reach a stable
-    state and neutron-induced fission if the nucleus splits into fragments (see Fig.~\ref{fig:both}).
+    state, and neutron-induced fission if the nucleus splits into fragments (see Fig.~\ref{fig:both}).
         \begin{figure}[h]
@@
     Therefore, radiative capture is the only possible neutron absorption process for
-    nuclei that are not fissionable.\cite{lb01} We will explore neutron-induced fission further in
-    the next section. Here we give an example of radiative capture. Thorium-232, which is present in highly radioactive
+    nuclei that are not fissionable.\cite{lb01}
+    We will explore neutron-induced fission further in the next section. Here we give an example of radiative capture. Thorium-232, which is present in highly radioactive
     waste, has a half-life longer than the age of the universe (see Table \ref{tab:hl}). However, it can go through
@@
         parts (see Fig.~\ref{fig:fission}).
-        The nucleus starts to vibrate with higher and higher amplitude as the kinetic energy of the
+        The nucleus starts to vibrate with increasing amplitude as the kinetic energy of the
         neutron transfers to the vibrational energy of the nucleus.
@@
         amplitude reaches a critical point, the Coulomb force overcomes the short-ranged nuclear
-        force (predominantly strong force) and the nucleus disintegrates into two large fragments and several neutrons.
+        force (predominantly strong force). As a result, the nucleus disintegrates into two large fragments and several neutrons.
         A typical reaction is described by
@@
         \end{equation}
-        Very often, photons, electrons and neutrinos are also emitted through beta and gamma decay.
+        Very often, photons, electrons, and neutrinos are also emitted through beta and gamma decay.
         The reaction described by Eq.~\ref{eq:u235} is not unique because different fission products can result
@@
         Apart from $6.5\%$ of the energy carried by the neutrinos which does not interact with the surrounding,
-        the rest of the energy can be utilized by virtue of outgoing particles depositing energy in the medium.
+        the rest of the energy can be utilized by virtue of outgoing particles depositing energy in the medium
+        surrounding the fuel inside the reactor.
 
@@
         On the one hand, fission of uranium-235 produces lighter radionuclides like cesium-137 and strontium-90, which
-        account for most of the short-term high activity and penetrating radiation (gamma radiations).
+        account for most of the short-term high activity and penetrating radiation (gamma radiation).
         In addition, for spent fuel freshly removed from the reactor, there is still a small amount ($\sim 1\%$)
-        of uranium-235 left along with a lot of free neutrons released from previous fission. Thus, a large amount
-        of energy is generated via fission of the leftover uranium-235 and heats up the spent fuel rods.
+        of uranium-235 left along with a lot of free neutrons released from previous fissions. Thus, a large amount
+        of energy is generated via fission of the leftover uranium-235, heating up the spent fuel rods.
 
@@
         and produce radionuclides that are
-        heavier than uranium (``transuranic'') such as plutonium, neptunium and americium.
-        These transuranic nuclei have relatively low activity but their half-lives are from a few thousand years
+        heavier than uranium (``transuranic'') such as plutonium, neptunium, and americium.
+        These transuranic nuclei have relatively low activity but their half-lives range from a few thousand years
         to a million years (see Table~\ref{tab:hl}).
@@
         \end{figure}
-        Therefore, in the short term, spent fuel emits intense radiations and generates a large amount of heat.
-        In the very long term (up to millions of years), its activity still remains well above the safe level.
+        Therefore, in the short term, spent fuel emits intense radiation and generates a large amount of heat.
+        Moreover, in the very long term (up to millions of years), its activity still remains well above the safe level.
         The combination of these properties makes the management and disposal of spent fuel and high-level waste
@@
     the next step. After that, reprocessing can extract the long-lived plutonium and uranium isotopes and
-    convert the waste into more manageable form (vitrification). However, reprocessing of commercial
-    nuclear spent fuel did not happen in the U.S. for cost and proliferation concerns. Instead, dry-cask
+    convert the waste into more manageable form. However, reprocessing of commercial
+    nuclear spent fuel was halted in the U.S. for cost and proliferation concerns. Instead, dry-cask
     storage, in addition to spent fuel pools, is used as an interim storage method.
     \subsection{Radiation shielding}
-    Since spent fuel contains high activity, shielding of radiations is crucial. Shielding works by
+    Since spent fuel contains high activity, shielding of radiation is crucial. Shielding works by
     forcing the radiation to interact with the shielding medium and hence reducing the amount of energy
     the radiation carries. In the context of nuclear waste management, the most common shielding medium
-    are water and concrete; and the mode of interactions of radiations with those mediums are similar to
-    the physical interaction between radiations and biological molecules (the most abundant of which is
+    are water and concrete; and the mode of interactions of radiation with those mediums are similar to
+    the physical interactions between radiation and biological molecules (the most abundant of which is
     water) described in Sec.~\ref{sec:interactions}.
@@
     Spent fuel pools  are usually about 15-meter deep water pools,
-    with storage rack at the bottom to hold the fuel assemblies. The large depth makes sure that the radiations
-    are well attenuated by the time they reach the surface (see Eq.\ref{eq:atten}).
+    with storage rack at the bottom to hold the fuel assemblies. The depth ensures that the radiation
+    is well attenuated by the time they reach the surface (see Eq.\ref{eq:atten}).
 
@@
     the cross section of water molecule. Hence the boron atoms captures the neutrons in the spent fuel
-    and prevent them from causing further fission and the formation of transuranic elements.
-
-    On the other hand, the water in spent fuel pool is circulated such that the heat transfers from the spent
-    fuel rods to the water via convective heat transfer. The convective heat transfer rate in this case is
+    and prevent them from causing further fission and formation of transuranic elements.
+
+    Additionally, the water in spent fuel pool is circulated such that heat dissipates from the spent
+    fuel rods to the water via convective heat transfer. The onvective heat transfer rate in this case is
     extremely complicated and we should not try to derive it here. However, we can still gain useful insights
@@
     where $Q$ denotes the heat transferred from the spent fuel to the water,
-    $h>0$ the heat transfer coefficient (which is assumed to be independent
+    $h$ the heat transfer coefficient (which is positive and assumed to be independent
     of the temperature), $T_{sf}$ the spent fuel temperature, and $T_{water}$ the temperature
@@
     \begin{equation}
-        T_{water} \propto h.T_{sf} - \frac{dQ}{dt}.
+        T_{water} \propto h\cdot T_{sf} - \frac{dQ}{dt}.
     \end{equation}
@@
     cooling (air circulation). The NRC is hoping that the casks will
-    last safely for as long as 100 years, yet it has been observed that corrosion and cracking occur in 30 years
+    last safely for as long as 100 years, yet it has been observed that corrosion and cracking can occur in 30 years
     or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
-    overground dry cask storage does not seem to be a long-term solution. The doubts were expressed
-    during the Nuclear Waste Technical Review Board 2009 meeting \cite{nwtb09, aa12}
-    \begin{displayquote}
-    Leaving the spent fuel on-site for extended periods of time was never intended and is not
-    responsible. (Dry cask manufacturer)
-
-    It is not ethical, basically to plan for long-term storage without pursuing a well defined repository
-    program. (Department of Energy employee speaking as a citizen)
-    \end{displayquote}
-    Even more striking is the fact that more than $75\%$ of the spent fuel are still stored in densely
+    overground dry cask storage does not seem to be a long-term solution.
+    Even more striking is the fact that more than $75\%$ of the spent fuel is still stored in densely
     packed pools
     that are close to or over their limiting capacity.\cite{a11,aa12}
-    Therefore, there is clearly a need for a long-term disposal plan, which we will explore in
-    Sec.~\ref{sec:disposal}.
+    Therefore, there is clearly a need for a long-term disposal plan.
 
@@
     Deep geological repositories that are stable over geological timescales naturally offer a good option for disposal.
-    In fact, it is accepted by the international science community that geological repositories in a stable environment over 300 meters below ground
+    In fact, it is accepted by the international science community that geological repositories in a stable environment over 300 meters underground 
     is safer than indefinite overground storage of HLW, reprocessed or not, as above-ground storage
-    are susceptible to catastrophic events like earthquakes and human-caused threats like terrorist attacks.\cite{fmr11}
-    In this section, we first use the Yucca Mountain Nuclear Waste
+    is susceptible to catastrophic events like earthquakes and human-caused
+    threats like terrorist attacks.\cite{fmr11}
+    In this section, we first use Yucca Mountain Nuclear Waste
     Repository     as a case study to understand the requirements for a permanent HLW disposal site.
@@
         maintenance once the waste was deposited. Therefore, scientists and engineers are challenged
-        to predict the geological environment of the Yucca Mountain over a million years and come up
+        to predict the geological environment of Yucca Mountain over the next one million years and come up
         with a design that they will have very little chance to alter once it is implemented.
@@
         of 1987 to be the nation's permanent deep underground repository for spent nuclear fuel. It is
-        located near the Nevada Nuclear Weapon Test Site in Nevada.
-        It is located 300-meter under the top of the Yucca mountain with impermeable layers of rocks
+        located near the Nevada Nuclear Weapon Test Site.
+        It is located 300-meter under the peak of Yucca mountain with impermeable layers of rocks
         separating the repository
@@
         Tectonic motion may change how much the rocks are fractured and thus the percolation rate.
-        Volcanic activity, meteorite hit and earthquakes may seriously undermine the integrity of the repository.
+        Volcanic activity, meteorite strike and earthquakes may seriously undermine the integrity of the repository.
         Hot upwelling water from below the repository may corrode the packaging and release radionuclides into
@@
         \subsubsection{Total System Performance Assessment (TSPA)}
-        The Department of Energy (DOE) created Total System Performance Assessment (TSPA) as a huge probabilistic
+        The Department of Energy (DOE) created Total System Performance Assessment (TSPA) as a complicated
+        probabilistic
         computer model that predicts the potential annual dose to the potential receptor.
         Essentially, the TSPA takes a huge set of parameters characterizing features that DOE deems as important,
-        carry out the geophysical and geochemical simulations,
+        carries out geophysical and geochemical simulations,
         and outputs the expected annual dose leaked from the repository.

### 2017-04-03 paper.tex
final; pending proofreading

@@
 \usepackage{graphicx}
+\usepackage{float}
 \graphicspath{ {images/} }
@@
 \newcommand{\red}[1]{\textcolor{red}{\bf #1}}
+\setlength{\skip\footins}{1.5cm}
 \begin{document}
@@
 \begin{abstract}
-%    As nuclear reactors around the world continuously generate highly radioactive waste,
-%    how we manage nuclear waste is becoming more pertinent. In this work,
-%    we examine a few scientific and technical issues related to nuclear waste management and disposal.
-%    The principles of radioactivity and nuclear fission help us to understand the sources and properties
-%    of different types of nuclear waste. Focusing on the highly radioactive spent fuel, the rest
-%    of this work outlines the advantages and drawbacks of different spent fuel management and disposal
-%    methods. We study the methods used for performance assessments for Yucca Mountain Nuclear Waste Repository,
-%    especially in the context of volcanic disruption. Finally, we discuss some possible options for 
-%    nuclear waste management and disposal after
-%    the termination of the Yucca Mountain project.\\
-%    Word count: 7449\\
+    As nuclear reactors around the world continuously generate highly radioactive waste,
+    the ways in which we manage nuclear waste are becoming more pertinent. In this work,
+    we examine some scientific and technical issues related to nuclear waste management and disposal.
+    The principles of radioactivity, biological effects of radiation, and nuclear fission help us to understand the sources and properties
+    of different types of nuclear waste. Focusing on the highly radioactive spent reactor fuel, the rest
+    of this work outlines the advantages and drawbacks of different spent fuel management
+    methods. We study the methods used for performance assessments for Yucca Mountain Nuclear Waste Repository,
+    especially in the context of volcanic disruption.\\
+    Word count: 7498
 \end{abstract}
@@
 requires extremely careful disposal methods.
-Moreover, there is currently no permanent repository in the world to dispose of these highly radioactive
+Moreover, there are currently no permanent repositories in the U.S. to dispose of these highly radioactive
 waste.\footnote{The Waste Isolation Pilot Plant (WIPP) in New Mexico is designed to be a permanent radioactive
@@
 for waste classified as transuranic, which, by the Nuclear Regulatory Commission standard, does not have particularly
-high radioactivity but only long half-life (See Sec.~\ref{sec:waste:classification}). In particular, WIPP does not accept spent nuclear
-fuel.}
+high radioactivity but only long half-life (See Sec.~\ref{sec:waste:classification}). In particular, WIPP does not accept spent nuclear fuel.\cite[chapt. 4]{aa12}}
 
@@
 Evidently, nuclear waste management and disposal are becoming increasingly urgent matters globally.
-In light of the high stake,
+In light of the high stakes,
 the uncertainty of the stability of a geological repository over a extremely long timescale has evolved into 
 a topic not only of scientific contention but also of policy debate.
-How well is the predicative power of science? How much trust should we put into the uncertainty estimates
+How good is the predictive power of science? How much trust should we put into the uncertainty estimates?
 
@@
 we examine the sources and the physical properties of nuclear waste (Section~\ref{sec:waste}),
-its management and temporary storage (Section~\ref{sec:temp}), and the debates around its eventual disposal (Section~\ref{sec:disposal}).
+its management and temporary storage (Section~\ref{sec:temp}), and its eventual disposal (Section~\ref{sec:disposal}).
 
@@
 
+        \subsubsection{Types of radiation}
+        There are three main types of radioactive decay processes that generate radiation in nuclear waste: 
+        Alpha decay, which ejects alpha particles (helium-4 nuclei), beta minus decay which ejects beta
+        particles (electrons) and gamma decay which ejects gamma rays.
+        Different types of radiation generated
+        have different properties and deserve separate treatments when we discuss health effects and shielding.
+        \paragraph{Alpha decay}
+        An alpha particle is a positively-charged 
+        helium-4 nucleus ${}^4_2$He$^{2+}$. In an alpha decay, a nucleus emits an alpha
+        particle (which has a particularly high binding energy) and decays into a different element.
+        The process can be represented by
+        \begin{equation}
+            {}^A_Z X_N \rightarrow {}^{A-4}_{Z-2}X'_{N-2} + \alpha,
+        \end{equation}
+        where $Z$ denotes the atomic number (number of protons) of a nucleus, $A$ the atomic mass, and $N$ the
+        neutron count.
+
+        \paragraph{Beta minus decay}
+        There are a variety of beta decay processes. Here we only describe the negative beta decay that is
+        relevant in this work. The basic decay process usually comes from the decay of a neutron
+        \begin{equation}
+            n \rightarrow p + e^- + \bar{\nu}_e,
+        \end{equation}
+        where $n$ denotes a neutron, $p$ a proton, $e^-$ an electron and $\bar{\nu}$ an electron antineutrino.
+        This can also happen inside a nucleus (bound neutron), represented by
+        \begin{equation}
+            {}^A_Z X_N \rightarrow {}^A_{Z+1} X'_{N-1} + e^- + \bar{\nu}_e.
+        \end{equation}
+
+        \paragraph{Gamma decay}
+        Most nuclear reactions, including the alpha and beta decay, leave the resulting nucleus in
+        an excited state.
+        We know that excited atoms radiate photons to fall back into ground state. Similarly, excited nuclei
+        decay to ground state via emitting photons, with energies in the range
+        of $0.1$ to $10$ MeV, which are characteristic of the energy differences between nuclear energy states.
+        The energy range corresponds to that of a gamma ray photon.
+
         \subsubsection{Quantifying radioactivity}
@@
         the activity of $1$ gram of radium-266.
-
-        Specific activity, on the other hand, gives the activity per mass of a sample
-        \begin{equation}
-        \mathcal{A}_s = \frac{\mathcal{A}(0)}{{m_{sample}}}.
-        \end{equation}
-        Specific activity can characterize the activity level of a certain isotope or compound.
-        The difference between
-        activity and specific activity resembles the difference between heat capacity and specific heat
-        capacity.
-        The radioactive isotope used for Computer Tomography (CT) scan has a specific activity
-        of $\sim 0.01$ Ci/g. See Table~\ref{tab:hl} for the half-lives and specific activities
-        of some isotopes present in highly radioactive nuclear waste.
         \begin{table}
@@
         \end{table}
-        \subsubsection{Types of radiation}
-        There are three main types of radioactive decay processes that generate radiation in nuclear waste: 
-        Alpha decay, which generates alpha particles (helium-4 nuclei), beta minus decay which generates beta
-        particles (electrons) and gamma decay which generates gamma rays.
-        The types of radiation (particle beam) generated
-        have different properties and deserve separate treatments when we discuss health effects and shielding.
-        \paragraph{Alpha decay}
-        An alpha particle is a positively-charged 
-        helium-4 nucleus ${}^4_2$He$^{2+}$. In an alpha decay, a nucleus emits an alpha
-        particle and decays into a different element. The emission of an alpha particle is
-        particularly favorable because it has a particularly
-        high binding energy. The process can be represented by
-        \begin{equation}
-            {}^A_Z X_N \rightarrow {}^{A-4}_{Z-2}X'_{N-2} + \alpha,
-        \end{equation}
-        where $Z$ denotes the atomic number (number of protons) of a nucleus $A$ the atomic mass, and $N$ the
-        neutron count.
-
-        \paragraph{Beta minus decay}
-        There are a variety of beta decay processes. Here we only describe the negative beta decay that is
-        relevant in this work. The basic decay process usually comes from the decay of a neutron
-        \begin{equation}
-            n \rightarrow p + e^- + \bar{\nu}_e,
-        \end{equation}
-        where $n$ denote  a neutron, $p$ a proton, $e^-$ an electron and $\bar{\nu}$ an electron antineutrino.
-        This can also happen inside a nucleus (bound neutron), represented by
-        \begin{equation}
-            {}^A_Z X_N \rightarrow {}^A_{Z+1} X'_{N+1} + e^- + \bar{\nu}_e.
-        \end{equation}
-
-        \paragraph{Gamma decay}
-        Most nuclear reactions, including the alpha and beta decay, leave the resulting nucleus in
-        an excited state.
-        We know that excited atoms radiate photons to fall back into ground state. Similarly, excited nuclei
-        decay to ground state via emitting photons, with energies in the range
-        of $0.1$ to $10$ MeV, which are characteristic of the energy differences between nuclear energy states.
-        The energy range correponds to that of a gamma ray photon.
-
-        \subsubsection{Daughter nuclei and subsequent decay\footnote{The derivation in this section follows that of
-        \cite[chapt. 1.5]{l01}}}
+
+        Specific activity, on the other hand, gives the activity per mass of a sample
+        \begin{equation}
+        \mathcal{A}_s = \frac{\mathcal{A}(0)}{{m_{sample}}}.
+        \end{equation}
+        Specific activity can characterize the activity level of a certain isotope or compound.
+        The difference between
+        activity and specific activity resembles the difference between heat capacity and specific heat
+        capacity.
+        As an illustration of the magnitude of the unit,
+        the radioactive isotope used for Computer Tomography (CT) scan has a specific activity
+        of $\sim 0.01$ Ci/g. See Table~\ref{tab:hl} for the half-lives and specific activities
+        of some isotopes present in highly radioactive nuclear waste.
+
+        \subsubsection{Daughter nuclei and subsequent decay\footnote{The derivation in this section follows that in
+        Lilley's text \cite[chapt. 1.5]{l01}.}}
         The resultant nuclei (called ``daughter products'') from a radioactive decay are often radioactive
@@
 
+        \begin{figure}
+            \centering
+            \includegraphics[height=0.7\textheight]{chain.png}
+            \caption{The neptunium-237 decay chain. The number below each isotope indicates the half-life.
+            Notice that neptunium-237 can decay into isotopes with significantly shorter half-lives and
+        release more radiation.\cite{chainpic}}
+            \label{fig:chain}
+        \end{figure}
         On the other hand, if the parent is longer lived
@@
         $A$ dominating the nuclei count. Taking the limit as $\lambda_A \rightarrow 0$, $N_B(t)\rightarrow 0$.
-        The decay is essentially the decay of $A$ directly into $C$. Hence, if we have a chain of reaction
+        The decay is essentially the decay of $A$ directly into $C$. Hence, if we have a chain of reactions
         where each daughter nucleus has a half-life much shorter than that of its parent, then 
         isotope $A$ effectively decays into the end of the decay chain and releases a lot more radiation than
-        in a single decay. Moreover, before the decay chain reaches its equilibrium, the activity of the sample
-        may even increase.
+        in a single decay from $A$ into $B$ (see Fig.~\ref{fig:chain} for an example). Therefore, one must consider the entire decay chain of an isotope
+        when assessing the radioactive risks it poses.
 
         \subsection{Biological effects of radiation\footnote{Much of the discussion in this section follows that in
-        Lilley's text \cite[chapt. 7]{l01}}}
+        Lilley's text \cite[chapt. 7]{l01}.}}
         The biological effects of radiation have been an important topic for decades. A large
@@
         field and the electrons in the matter it travels through. Interactions with nuclei and head-on collisions
-        are quite rare. Since the speed of an alpha particle is low, it deposits more energy as it
-        travels. In living tissues, an $1$-MeV alpha particle travels through only a few tens of micrometers and thus
-        are easily stopped by skin. However, if alpha particles are ingested and more sensitive organs are irradiated,
-        serious damages can be done.
-
-        Beta radiation consistens of electrons. Electrons are much lighter and hence are easily scattered.
+        are quite rare. Since the speed of an alpha particle is low, it deposits more energy per distance traveled
+        because there is more time for the particle to interact with the medium.
+        In living tissues, an $1$-MeV alpha particle travels through only a few tens of micrometers and thus
+        are easily stopped by skin. However, if alpha particles are ingested and sensitive internal organs are irradiated,
+        serious damage can be done.
+
+        Beta radiation consists of electrons. Electrons are much lighter and hence are easily scattered.
         Moreover, they are much more penetrating since their energy lost per distance traveled is lower.
         Finally, electrons also radiate photons
-        via bremsstrahlung radiation as they decelerate. Therefore, the electrons' energy end up distributing
+        via bremsstrahlung radiation as they decelerate. Therefore, the electrons' energy ends up distributing
         across a much larger volume.
@@
         via photoelectric effects and secondary Compton scattering. As a result, the intensity of the energy deposition falls
-        off much more slowly and is hence more penetrating. Both electrons and photons can penetrate human body easily.
+        off much more slowly and is hence more penetrating.
+
+        Compared to alpha particles, electrons and photons can penetrate human body easily.
         Although they do not deposit as much energy as alpha particles, they are generally more dangerous due to their
@@
         The biological impact of radiation depends on the type, power and energy of the radiation, the
-        exposure time, and the part of the body exposed. Various quantities was developed to
-        describe the biological impacts of radiation. Here we review three
+        exposure time, and the part of the body exposed. Various quantities were developed to
+        quantify the biological impacts of radiation. Here we review three
         of these quantities: the absorbed dose, the equivalent dose and the effective dose.
 
-        The effects of radiation is usually characterized by the amount of energy it carries since it constrains
+        The effect of radiation is usually characterized by the amount of energy it carries since it constrains
         the amount of energy it deposits in living tissues.
@@
         of energy $E_{\gamma}$
-        (in MeV), one can approximate the dose rate (dose received per unit of time) by \cite[chapt. 7]{l01}
+        (in MeV) from $r$ meters away,
+        one can approximate the dose rate (dose received per unit of time) by \cite[chapt. 7]{l01}
         \begin{equation}
@@
         MeV and therefore should be good for all sources we are considering.\cite{my68}
-        The inverse-squared dependence comes from the isotropic radiation assumption.
+        The inverse-squared distance dependence comes from the isotropic radiation assumption.
 
@@
         However, the effects of the radiation on living tissue depend on the type of radiation.
-        For example, the same dose of alpha particle deposit more energy in living tissue than gamma rays.
+        For example, the same dose of alpha particles deposit more energy in living tissue than gamma rays.
         To quantify the {\it effects} of radiation, we introduce equivalent
@@
         where $D_R$ is the absorbed dose and $w_R$ the weighting factor of a given kind of radiation $R$
-        (see Table~\ref{tab:eq} for a list of weighing factors). The unit for equivalent dose is sievert (Sv).
+        (see Table~\ref{tab:eq} for a list of weighting factors). The unit for equivalent dose is sievert (Sv).
         $1$ Sv is defined as the effects equivalent to $1$ Gy of Gamma radiation on living tissues. It describes
@@
 
-        Hereditary effects of radiation can occur if the reproduction cells were damaged by radiation.
+        Hereditary effects of radiation can occur if the reproduction cells are damaged by radiation.
         The mechanism is similar to that of delayed health effects, since DNA molecules are passed on
@@
 \label{sec:waste}
-    In this section, we will first look at a classification scheme of nuclear waste and then
-    examine the source, composition, and properties of the most detrimental type--the high-level waste.
+    In this section, we will look at the classification of nuclear waste and
+    apply some of the concepts outlined in the last section to 
+    analyzing the source, composition, and properties of the most detrimental type--the high-level waste.
     \subsection{Classification of nuclear waste}
@@
 
-    \paragraph{High-level waste (HLW) and spent fuel}
-    HLW has high activity contents and usually generates a large amount of
+    High-level waste (HLW) has high activity contents and usually generates a large amount of
     heat as it radiates. The typical source of HLW is spent fuel from the nuclear reactor.
-    In fact, since there is no commercial reprocessing facilities in the U.S. (see section~\ref{sec:reproc}),
+    In fact, since there are no commercial reprocessing facilities in the U.S. (see Sec.~\ref{sec:reproc} for
+    more details),
     almost all of the HLW is in the form of spent nuclear fuel from the reactor.
-    \paragraph{Transuranic waste (TRU)}
-    TRU is defined as materials that do not belong to high-level waste and are
+
+    Transuranic waste (TRU) is defined as materials that do not belong to high-level waste and are
     contaminated by alpha-emitting radionuclides with
@@
     of TRU is usually low but it remains active for long period of time. Moreover, TRU can be easily turned
-    into plutonium-239, the {\it de facto} material to fabricate atomic bombs.
-
-    \paragraph{Low-level waste (LLW)}
-    LLW has relatively little radioactivity, produces almost no heat,
+    into plutonium-239 (the {\it de facto} material to fabricate atomic bombs), via neutron capture and alpha decay
+
+    Low-level waste (LLW) has relatively little radioactivity, produces almost no heat,
     and contains practically no
-    transuranic elements. Examples are papers and tools used in cleanup operations in the reactor.
+    transuranic elements. Examples are papers and tools used in routine cleanup operations in the reactor.
     Most LLW does not require shielding and may be buried in landfill 
@@
 
-    \paragraph{Uranium tailings}
     Uranium mill tailings from uranium mills are another form of waste with low activity, but
@@
     Over geological timescales, the stable isotope uranium-238 
-    in the ore decays into radioactive thorium and radium 
+    in the ore radioactively decays into radioactive thorium and radium 
     These radionuclides remain after the uranium (all of its isotopes)
@@
     from contaminating water and air.
-
-    Since all categories but high-level waste are relatively easy to handle, in the rest of
-    this section we will focus on the formation and
+    \subsection{Creation and properties of spent fuel}
+    Since all categories but high-level waste are relatively easy to handle,
+    we will focus on the formation and
     properties of spent nuclear fuel, the dominant contribution to high level waste,
-
-    \subsection{Creation and properties of spent fuel}
     \label{sec:waste:spentfuel}
-    Spent fuel comes from used nuclear fuel. Here we look at the source and properties of spent fuel.
     \subsubsection{Nuclear fuel}
-    For most of the commercial nuclear reactors, the energy comes from the isotope uranium-235 capable
+        \begin{figure}
+            \centering
+            \includegraphics[width=\textwidth]{rod.jpg}
+            \caption{Photo of a nuclear fuel rod.\cite{rodpic}}
+            \label{fig:rod}
+        \end{figure}
+    For most of the commercial nuclear reactors, the energy generated comes from the isotope uranium-235 capable
     of undergoing fission.
@@
     via isotope separation to increase the uranium-235 content. Typical fresh nuclear fuel, usually packaged in
-    rods, contains $3\%$ uranium-235 and $97\%$ uranium-238.
+    rods (Fig.~\ref{fig:rod}), contains $3\%$ uranium-235 and $97\%$ uranium-238.
 
@@
     \label{sec:capture}
-    Absorption of thermal neutron (with energy $\sim 0.025$ eV, the most likely speed of a free neutron at 
-    a temperature of $290$ K) is the main reaction that
+    Absorption of thermal neutrons (with energy $\sim 0.025$ eV, the most likely speed of a free neutron at 
+    a temperature of $290$ K) by uranium-235 is the main reaction that
     occurs in the reactor. The neutrons come from a startup neutron source and subsequently from the fission
@@
     state, neutron absorption is classified into radiative capture if the nucleus radiates to reach a stable
-    state, and neutron-induced fission if the nucleus splits into fragments (see Fig.~\ref{fig:both}).
-        \begin{figure}[h]
-            \centering
-            \includegraphics[width=0.8\textwidth]{pu.png}
+    state, and neutron-induced fission if the nucleus splits into fragments (see Fig.~\ref{fig:both} for an illustration
+    of the two decay pathways).
+        \begin{figure}
+            \centering
+            \includegraphics[width=0.6\textwidth]{pu.png}
             \caption{Schematic illustration of the two different neutron absorption processes:
@@
     waste, has a half-life longer than the age of the universe (see Table \ref{tab:hl}). However, it can go through
-    radiative capture and then subsequent beta decays to form uranium-233 (neutrinos are omitted in the equations)
+    radiative capture and then subsequent beta decays to form uranium-233: (neutrinos are omitted in the equations)
     \begin{align}
@@
     thorium-232. Another important radiative capture reaction happening inside nuclear fuel is the radiative
-    capture of uranium-238 (the most abundant isotope in nuclear fuel) to form plutonium-239.
+    capture of uranium-238 (the most abundant isotope in nuclear fuel) to form plutonium-239, given
+    by
     \begin{equation}
@@
     \end{equation}
-        \subsubsection{Fission and fission products}
+    \subsubsection{Fission and fission products\footnote{
+    This section follows the discussions in Lilley's text \cite[chapt. 10]{l01}.}}
         \label{sec:fission}
@@
         the fission of uranium-235 induced by a thermal neutron
-        is the best-known example of induced fission. 
+        is the best-known example of induced fission.\cite[chapt. 1]{gc01}
         The neutron capture produces a compound nucleus ${}^{236}U$ that is in an excited state.
@@
         continue to induce fission of other uranium-235 nuclei, causing a ``chain reaction'' of subsequent
-        fission that generate a huge amount of energy.\cite[chapt. 10]{l01}
-        \begin{figure}[h]
+        fission that generate a huge amount of energy in a very short time.
+        \begin{figure}
             \centering
@@
         \begin{equation}
-            E = (m_{parent}-m_{product})c^2,
+            E = (m_{parent}-m_{products})c^2,
         \end{equation}
@@
         for a nucleus with atomic number around $117$. Hence, the change of binding energy per nucleon is
-        $0.9$ MeV u${}^{-1}$, which amount to $212$ MeV for a uranium-235 atom.
-        Apart from $6.5\%$ of the energy carried by the neutrinos which does not interact with the surrounding,
+        $0.9$ MeV u${}^{-1}$, which amounts to $212$ MeV for a uranium-235 atom.
+        Apart from $6.5\%$ of the energy carried by the neutrinos which do not interact with the surrounding,
         the rest of the energy can be utilized by virtue of outgoing particles depositing energy in the medium
@@
 
-        In reality, thermal induced fission of uranium-235 creates a variety of products and hence the energy released
+        In reality, induced fission of uranium-235 creates a variety of products and hence the energy released
         in a single reaction vary by tens of MeV. On average, about $220$ MeV per fission is recoverable for energy conversion.
@@
         (see Fig.~\ref{fig:spentfuel} for an illustration of the composition of spent fuel).
-        \begin{figure}[h]
+        \begin{figure}
             \centering
             \includegraphics[width=0.6\textwidth]{spent-fuel.png}
-            \caption{Composition of spent nuclear fuel for a light water reactor. The exact composition varies
+            \caption{Composition of spent nuclear fuel for a reactor. The exact composition varies
             by different kinds of nuclear reactors. ``Minor actinides'' refer to transuranic elements other than plutonium.\cite{gao}}
@@
         account for most of the short-term high activity and penetrating radiation (gamma radiation).
-        In addition, for spent fuel freshly removed from the reactor, there is still a small amount ($\sim 1\%$)
+        In addition, for spent fuel freshly removed from the reactor, there is still a small amount ($\sim 1\%$ of
+        the spent fuel)
         of uranium-235 left along with a lot of free neutrons released from previous fissions. Thus, a large amount
@@
         an issue when one deals with these elements.
-        \begin{figure}[h]
+        \begin{figure}
             \centering
@@
     level of $0.01$ Ci is comparable to that of the radiation source used for Computer Tomography (CT) scans.
-Data from Table~\ref{tab:hl}}
+    Graph produced by author with data from Table~\ref{tab:hl}}
             \label{fig:csru}
@@
     Spent fuel from nuclear reactors is the major source of HLW.
+    In this section, we will study in details some important processes involved in the short-term management
+    of spent fuel.
     The concept of shielding is crucial in the short-term management of spent fuel due to the high activity
@@
     a beam of particles with intensity $I_i$ (with unit (area$\cdot$time)${}^{-1}$) passing through a medium
-    with number density of molecules $N_s$. Suppose on average $R_i$ interactions occur per time unit.
+    with $N_s$ molecules interacting with the incident particles.
+    Suppose on average $R_i$ interactions occur per time unit.
     The cross section of the interaction between the shielding medium defined as
@@
     with the distance traveled. Notice that $\mu$ and $z$ play symmetrical roles in the exponent. Hence
-    in practice, we can trade quantity of shielding medium for the shielding capability of the type of
+    in practice, we can trade the depth of shielding medium for the shielding capability of the type of
     molecule.
-
+        \begin{figure}
+            \centering
+            \includegraphics[width=0.8\textwidth]{pool.jpg}
+            \caption{Photo of a spent fuel pool with racks for spent fuel rods at the bottom.\cite{s16}}
+            \label{fig:pool}
+        \end{figure}
     \subsection{Spent fuel pools}
     Short-term cooling of spent fuel is necessary before further disposal or reprocessing can be done.
-    Radiation shielding and cooling are the two important aspects of spent fuel pools.
+    Typically, the spent fuel undergoes cooling in spent fuel cools (Fig.~\ref{fig:pool}) which also shield the radiation.
 
@@
     The water is enriched with boron atoms that have neutron capture cross section of about $1000$ times
-    the cross section of water molecule. Hence the boron atoms captures the neutrons in the spent fuel
+    the cross section of water molecule. Hence the boron atoms capture the neutrons produced by fission
+    in the spent fuel
     and prevent them from causing further fission and formation of transuranic elements.
@@
     Additionally, the water in spent fuel pool is circulated such that heat dissipates from the spent
-    fuel rods to the water via convective heat transfer. The onvective heat transfer rate in this case is
+    fuel rods to the circulated water via forced convective heat transfer.
+    The convective heat transfer rate in this case is
     extremely complicated and we should not try to derive it here. However, we can still gain useful insights
@@
     of the temperature), $T_{sf}$ the spent fuel temperature, and $T_{water}$ the temperature
-    of the water.
+    of the water.\cite[chapt. 1]{l13}
     The heat transfer coefficient $h$ depends of the fluid dynamics properties of the water and can
     be measured experimentally. The goal of
-    cooling is to prevent heat from accumulating in the spent fuel and thus $dQ/dt \geq 0$ at all time.
+    cooling is to prevent heat from accumulating in the spent fuel and thus $dQ/dt \geq 0$ for all time.
     Apply the first and third laws of thermodynamics to the spent fuel rod,
@@
     \end{equation}
-    where $Q\geq 0$ is consistent with the definition above and $S$ denotes entropy, we conclude that $Q$ and $T_{sf}$ always vary in the opposite direction. Rearrange Eq.~\ref{eq:cooling} to get
+    where $Q\geq 0$ is consistent with the definition above and $S$ denotes entropy, we conclude that $Q$ and $T_{sf}$ always varies in the opposite direction. Rearrange Eq.~\ref{eq:cooling} to get
     \begin{equation}
@@
     fuel is cooled.
+        \begin{figure}
+            \centering
+            \includegraphics[width=0.8\textwidth]{dry_cask.jpg}
+            \caption{A photo of dry casks.\cite{dcpic}}
+            \label{fig:cask}
+        \end{figure}
+    \subsection{Dry cask storage}
+    Spent fuel pools have very limited capacities. 
+    In the case where the spent fuel is not reprocessed, overground dry cask storage (Fig.~\ref{fig:cask}) provides a low-cost temporary storage before the spent fuel is disposed of.
+    After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks weighing
+    100 tons filled with inert gas. 
+    The dry casks are often metal cylinders welded to prevent leakage. Each cask is surrounded by
+    additional steel and concrete to provide shielding.
+    The heat from the cask is removed by passive
+    cooling (air circulation) and therefore no power supply is needed.
+
+    It is useful to draw a comparison between dry cask storage and spent fuel pools here.
+    Spent fuel pools require active cooling with a power source. As a result, natural disasters (like the Fukushima
+    tsunami) and terrorist attacks could potentially shut down the power supply and cause a fire that would
+    send radioactive smoke into the air.
+    On the other hand, dry casks have been observed to withstand earthquakes and floods.
+    The NRC estimates that dry casks will
+    last safely for as long as 100 years, yet it has been observed that corrosion and cracking can occur in 30 years
+    or less.\cite{aa12} But overall, dry cask storage is a much safer option than spent fuel pools.\cite{r14} Nevertheless, in the US, the nuclear power industry has been reluctant to shift toward dry cask storage since it
+    involves (albeit modest) extra cost than simply storing spent fuel in pools.\cite{r14}
+    Even more striking is the fact that more than $75\%$ of the spent fuel is still stored in densely
+    packed pools
+    that are close to or over their limiting capacity.\cite{a11,aa12}
 
@@
     for reprocessing.
-    Spent nuclear fuel contains a large amount of plutonium and uranium that are suitable as nuclear fuel
+    Spent nuclear fuel contains a large amount of plutonium and uranium-235 that are suitable as nuclear fuel
     or fissile weapon material (plutonium-239). Therefore, reprocessing the spent fuel generates plutonium and
@@
     reprocessing facility due to the demonstrated high cost of reprocessing
-    during the West Valley Demonstration Project and water contamination issues at the Hanford Site.
+    during the West Valley Demonstration Project (a reprocessing experiment project)
+    and water contamination issues at the Hanford Site.
     Concerns about proliferation issues also arise, as the 
-    reprocessed plutonium-239 is the primary fissile element in the production of nuclear weapons.\cite{aa12}
-    On the other hand, France (the world leader in reprocessing technology), India, Russia, the UK
-    and China (since 2010) continues to reprocess spent fuel. To understand the details of reprocessing
+    reprocessed plutonium-239 is the primary fissile element in the production of nuclear fission weapons.\cite{aa12}
+    On the other hand, other countries like France, the UK and China
+    continue to reprocess spent fuel.\cite{v14,h14,h15}
+    To understand the details of reprocessing
     requires advanced chemistry and is beyond the scope of this work. However, the issue remains interesting
@@
     Almost all of the reprocessing plants use the PUREX (Plutonium Uranium Redox EXtraction) process
-    to reprocess spent fuel. This process utilizes the differences in oxidation potential 
+    to reprocess spent fuel. This process utilizes the differences in chemical potential in oxidation reactions
     between the heavy elements (uranium and plutonium) from the other lighter fission products.
@@
     waste can be turned into liquid form, mixed with frit (the substance from which glass is made) and then
-    vitrified (made into glass). The vitrification process immobilizes the radioactive particles and makes
-    the waste much more safe and manageable, but vitrification can only be applied to reprocessed spent fuel.
-
-    It is worthwhile to note that the PUREX processed was designed specifically to separate plutonium for
+    vitrified (made into glass). The process of turning nuclear waste into glass, called vitrification,
+    immobilizes the radioactive particles and makes
+    the waste much more safe and manageable, but only reprocessed spent fuel can be vitrified.\cite{v14}
+
+    It is worthwhile to note that the PUREX process was designed specifically to separate plutonium for
     bombs. Therefore, literally any other reprocessing approach would be less prone to proliferation.
     Researchers are actively investigating alternatives to the PUREX technology. For example, the UREX+ (for
-    uranium extraction) technique extracts the uranium but leaves the plutonium mixed with other
-    radioactive materials.\cite{aa12}
-
-    \subsection{Dry cask storage}
-    Spent fuel pools have very limited capacities. 
-    In the case where the spent fuel is not reprocessed, overground dry cask storage provides a low-cost temporary storage before the spent fuel is disposed of.
-    After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks filled with inert gas. 
-The dry casks are often metal cylinders welded to prevent leakage. Each cask is surrounded by
-    additional steel and concrete to provide shielding.
-    The heat from the cask is removed by passive
-    cooling (air circulation). The NRC is hoping that the casks will
-    last safely for as long as 100 years, yet it has been observed that corrosion and cracking can occur in 30 years
-    or less.\cite{aa12} As such, without the prospect of a permanent geological repository,
-    overground dry cask storage does not seem to be a long-term solution.
-    Even more striking is the fact that more than $75\%$ of the spent fuel is still stored in densely
-    packed pools
-    that are close to or over their limiting capacity.\cite{a11,aa12}
-    Therefore, there is clearly a need for a long-term disposal plan.
-
-\section{Disposal of HLW}
+    uranium extraction) technique extracts the uranium but leaves the plutonium mixed in low concentration
+    with other
+    radioactive materials that are eventually disposed of.\cite[chapt. 7]{aa12}
+
+\section{Disposal of High-level Waste}
 \label{sec:disposal}
-    The time frame of nuclear waste disposal is about $10^4-10^6$ years, after which the radioactivity
+    The time frame of nuclear waste storage is about $10^4-10^6$ years, after which the radioactivity
     of the waste is predicted to decline to a safe level (see Sec.\ref{sec:waste:spentfuel}).
@@
     Repository as a case study to understand the requirements for a permanent HLW disposal site.
-    Then we look at an alternative to a permanent deep geological solution:
-    deep boreholes storage.
+
     \begin{figure}[h]
             \centering
-            \includegraphics[width=\textwidth]{yuccaRepo.jpg}
+            \includegraphics[width=0.9\textwidth]{yuccaRepo.jpg}
             \caption{Design of the 300-meter deep Yucca Mountain Waste Repository.\cite{yucpic}}
@@
         \end{figure}
+
     \subsection{Deep geological repository: Yucca Mountain}
@@
         from both the surface and the ground water (see Fig.~\ref{fig:yucca}).
-
-        The project has been highly contested since its
-        inception and the Obama administration ended federal funding for the project in 2011, citing political,
-        not safety or technical, reasons. We will take a glimpse at the suitability assessment of the site
+        In this section we will take a glimpse at the suitability assessment of the site
         as a nuclear waste repository to get a sense of its technical complexity.
 
-        \subsubsection{Geological stability of Yucca Mountain\footnote{This section is a short summary of \cite{m06}}}
+        \subsection{Geological stability of Yucca Mountain\footnote{This section is a short summary of \cite{m06}}}
         The chief concern of the suitability of Yucca Mountain was about water. If moist air comes to
@@
         releasing radionuclides.
-        The community seemed to agree that the packaging will have to carry the primary
-        responsibility of preventing the oxidation of the waste, but predicting the performance of some
-        engineered product over the timescale of a million years is not something with which we have had much 
+        The community seemed to agree that the waste packaging will have to carry the primary
+        responsibility of preventing the oxidation of the waste, but we have not had a lot of experience
+        predicting the performance of some
+        engineered product over the timescale of a million years
         experience. Over time, the radiation from the waste will also alter the rock (by altering its composition
@@
 
-        \subsubsection{Total System Performance Assessment (TSPA)}
+        \subsection{Total System Performance Assessment (TSPA)}
         The Department of Energy (DOE) created Total System Performance Assessment (TSPA) as a complicated
@@
 
-        \subsubsection{Risk and uncertainty estimation: volcanism}
+        \subsection{Risk and uncertainty estimation: volcanism}
         \label{sec:volcano}
@@
         one in ten million per year, it is acceptable.\cite{epa93}
-        Of course, one has to also consider consequences of a potential radiological releases if an eruption
+        Of course, one has to also consider consequences of a potential radiological release if an eruption
         occurs. Here we focus on the estimation of occurrence probability.
 
-        The probability that a future volcanic eruption disrupting a repository over
+        The probability of a future volcanic eruption disrupting a repository over
         the repository's lifetime is given by
@@
         The information necessary to estimate the probability is listed in the flowchart Fig.~\ref{fig:volcanos}.
-        The boxes represent decision parameters that depends on the judgment of the decision makers and
+        The boxes represent decision parameters that depend on the judgment of the decision makers and
         the scientists. For example the ``Volcanic Zone Models'' should be picked by the scientists, and
@@
         life span of the repository. Here we look in some detail at the calculation of those parameters.
-        \begin{figure}[h]
+        \begin{figure}
             \centering
@@
         are combined through a Monte-Carlo simulation to produce the disruption rate $P(D)$
-        at the Yucca Mountain repository. DOE estimates a mean disruption rate of $1.5\times 10^8$ events
+        at the Yucca Mountain repository. DOE estimates a mean disruption rate of $1.5\times 10^{-8}$ events
         per year with
@@
         a scenario.\cite{cv14}
-        \subsection{Deep borehole disposal}
-        Deep borehole disposal utilizes deep and narrow boreholes to dispose of HLW. The boreholes
-        will place the packaged waste as much as 5 kilometers under the surface of the ground and seal it
-        with concrete clay.
-        The engineered packaging, along with the granite rock with few cracks on which the borehole is drilled,
-        provides
-        good isolation for the radionuclides in the waste (See Fig.~\ref{fig:borehole}).
-        \begin{figure}[h]
-            \centering
-            \includegraphics[width=0.8\textwidth]{borehole.jpg}
-            \caption{Illustration of borehole disposal, with deep repositories' depths for comparison.\cite{c15}}
-            \label{fig:borehole}
-        \end{figure}
-
-        Deep boreholes have several advantages over deep geological repository. For example, suitable
-        rocks can be more readily found for boreholes (since it does not take up a lot of space and can be
-        of smaller scale),
-        whereas large repository sites are hard to identified. \cite{b09,c15}
-        However, the borehole solution has its own limitations.
-        Drilling a hole as deep as 5 kilometers poses a daunting engineering challenge.
-        Moreover, reactor rods from commercial nuclear plants in the U.S. usually size up to 2 meters in diameter,
-        which would be too large for the borehole and require extensive repackaging.
-        Nevertheless, deep borehole disposal remains a valid option.
-        Sandia lab has embarked on a five-year plan to experiment with the deep borehole design.\cite{c15}
+
+        Just like volcanism, many risk factors in the  Yucca Mountain project turned out much more complicated
+        than people initially expected. The project has been highly contested since its
+        inception due to scientific and political reasons. The Obama administration ended federal funding for the project in 2011, citing political,
+        not safety or technical, reasons.\cite{m14} Nevertheless, the Trump administration is considering
+        reviving the project.\cite{z17}
 
@@
 Nuclear waste management and disposal is certainly becoming an increasingly pressing issue.
-We explore the creation, management and disposal of highly radioactive nuclear waste and the physical
-principles underlying those processes. The increasingly over-packed spent fuel pools and the challenges
+We explored the creation, management and disposal of highly radioactive nuclear waste and the physical
+principles underlying these processes. The increasingly over-packed spent fuel pools and the challenges
 associated with geological repositories are the weak links in this chain.
@@
 
-Newer reprocessing methods (as mentioned in Sec.~\ref{sec:reproc}) are being proposed
+Newer reprocessing methods are being proposed and tested
 as  solutions to the proliferation issue from the current reprocessing method.
@@
 
-%\begin{acknowledgments}
-%Thanks Joel and peeps. Arjendu and M. V. Ramana.
-%\end{acknowledgments}
+\begin{acknowledgments}
+I would like to thank my primary advisor Joel Weisberg for being an incredible resource and
+source of support, secondary advisor Cindy Blaha and peer advisor Alex McMurty for providing extremely
+insightful and helpful comments, and M. V. Ramana for discussions to get me started on this topic.
+\end{acknowledgments}
 \newpage
@@
 \bibliography{refs}
-%\newpage
-%\appendix
-%\section{Derivation of Fermi's Golden Rule \#2}
-%\label{a:fermi}
 \end{document}

### 2017-04-03 paper.tex
done

@@
 most of which is spent nuclear fuel from nuclear reactors.\cite{iaea08, r12}
-The world will have over $457,000$ metric tons of highly radioactive waste stock by the end of 2020.\cite{r12}
+The world will have over $457,000$ metric tons of highly radioactive waste in stock by the end of 2020.\cite{r12}
 Highly radioactive waste remains hazardous for up to $10^6$ years after it is produced and thus
 requires extremely careful disposal methods.
-Moreover, there are currently no permanent repositories in the U.S. to dispose of these highly radioactive
+Moreover, there are currently no permanent repositories in the U.S. to dispose of the highly radioactive
 waste.\footnote{The Waste Isolation Pilot Plant (WIPP) in New Mexico is designed to be a permanent radioactive
@@
 of the U.S. government overseeing nuclear energy safety,
-estimates that a major fire in a U.S. spent fuel pool would
+estimated that a major fire in a U.S. spent fuel pool would
 displace 3.4 million people in an area larger than New Jersey.\cite{s16}
@@
 In light of the high stakes,
-the uncertainty of the stability of a geological repository over a extremely long timescale has evolved into 
+the uncertainty of the stability of a permanent repository over a extremely long timescale has evolved into 
 a topic not only of scientific contention but also of policy debate.
@@
         \end{equation}
-        where $n_f$ is the number density of the state. \cite[chapt. 6]{k88}
+        where $n_f$ is the number density of the state.\cite[chapt. 6]{k88}
         In simple terms, $\rho(E_f)dE$ gives the proportion of the particles with energy in the
@@
         the decay probability is small (meaning $V_p\ll V$), the result from perturbation theory should hold. Since
-        there is no time dependency in Eq.~\ref{eq:trans},the
+        there is no time dependency in Eq.~\ref{eq:trans}, the
         decay rate should be independent of time, which can be expressed as
@@
 
-        We can easily solve the differential equation given by Eq.~\ref{eq:lambda} to get
+        We can solve the differential equation given by Eq.~\ref{eq:lambda} to get
         the number of nuclei as a function of time $t$
@@
         In addition, the mean lifetime $\tau$ is defined as the averaged lifetime of a collection of nuclei.
-        From Eq.~\ref{eq:lambda}, the number of particles between $t$ and $t+dt$ is $dN=\lambda N dt=\lambda
+        From Eq.~\ref{eq:lambda}, the number of particles during time interval $t$ and $t+dt$ is $dN=\lambda N dt=\lambda
         N_0 e^{-\lambda t}dt$. Hence we have
@@
         The half-life, $t_{1/2}$, is the time after which half of the nuclei in the sample are expected to
-        decay. Simply substitute $N(t)=1/2N_0$ in Eq.~\ref{eq:decay} to get
+        decay. Substitute $N(t)=1/2N_0$ in Eq.~\ref{eq:decay} to get
         \begin{equation}
@@
         There are three main types of radioactive decay processes that generate radiation in nuclear waste: 
-        Alpha decay, which ejects alpha particles (helium-4 nuclei), beta minus decay which ejects beta
+        alpha decay, which ejects alpha particles (helium-4 nuclei), beta minus decay which ejects beta
         particles (electrons) and gamma decay which ejects gamma rays.
-        Different types of radiation generated
+        Different types of radiation 
         have different properties and deserve separate treatments when we discuss health effects and shielding.
@@
         helium-4 nucleus ${}^4_2$He$^{2+}$. In an alpha decay, a nucleus emits an alpha
-        particle (which has a particularly high binding energy) and decays into a different element.
+        particle $\alpha$ (which has a particularly high binding energy) and decays into a different element.
         The process can be represented by
@@
         As an illustration of the magnitude of the unit,
-        the radioactive isotope used for Computer Tomography (CT) scan has a specific activity
+        the radioactive isotope used for Computer Tomography (CT) scans has a specific activity
         of $\sim 0.01$ Ci/g. See Table~\ref{tab:hl} for the half-lives and specific activities
@@
 
-        \subsubsection{Daughter nuclei and subsequent decay\footnote{The derivation in this section follows that in
+        \subsubsection{Daughter nuclei and subsequent decays\footnote{The derivation in this section follows that in
         Lilley's text \cite[chapt. 1.5]{l01}.}}
         The resultant nuclei (called ``daughter products'') from a radioactive decay are often radioactive
-        themselves and their decay rate can be larger or smaller than their parents. Consider a decay chain
+        themselves and their decay rate can be larger or smaller than their parents'. Consider a decay chain
         $A\rightarrow B \rightarrow C \rightarrow ...$ with decay constants $\lambda_A, \lambda_B,...$. Suppose
-        we start with a sample of $N_0$ $A$ nuclei. The change of the number of nuclei $B$ increases as $A$
+        we start with a sample of $N_0$ nuclei $A$. The change of the number of nuclei $B$ increases as $A$
         decays and decreases s $B$ decay. Hence we have
@@
         \end{equation}
-        Multiply both sides by an integrating factor $e^{\lambda_B t}$ and substitute in the number of particle
+        Multiply both sides by an integrating factor $e^{\lambda_B t}$ and substitute in the number of nuclei
         $A$ as a function of time $N_A=N_0 e^{-\lambda_A t}$ to get
@@
         \end{align}
-        Hence we end up with the number of $B$ nuclei as a function of time
+        Hence we end up with the number of nuclei $B$ as a function of time
         \begin{equation}
@@
         Similar derivations can be done for the rest of the isotopes in the decay chain by considering
-        one decay in a chain at a time. If the daughter isotope has a longer half-life than the parent
+        one decay in a chain at a time. If the daughter isotope has a longer half-life than does the parent
         ($\lambda_A \gg \lambda_B$), for small $t$, Eq.~\ref{eq:daughter} becomes
@@
         $A$ dominating the nuclei count. Taking the limit as $\lambda_A \rightarrow 0$, $N_B(t)\rightarrow 0$.
-        The decay is essentially the decay of $A$ directly into $C$. Hence, if we have a chain of reactions
+        The decay is essentially the decay of $A$ directly into $C$. Hence, if we have a chain of decays 
         where each daughter nucleus has a half-life much shorter than that of its parent, then 
         isotope $A$ effectively decays into the end of the decay chain and releases a lot more radiation than
-        in a single decay from $A$ into $B$ (see Fig.~\ref{fig:chain} for an example). Therefore, one must consider the entire decay chain of an isotope
+        in a single decay of $A$ into $B$ (see Fig.~\ref{fig:chain} for an example). Therefore, one must consider the entire decay chain of an isotope
         when assessing the radioactive risks it poses.
@@
         because there is more time for the particle to interact with the medium.
-        In living tissues, an $1$-MeV alpha particle travels through only a few tens of micrometers and thus
-        are easily stopped by skin. However, if alpha particles are ingested and sensitive internal organs are irradiated,
+        In living tissues, an $1$-MeV alpha particle travels through only a few tens of micrometers and
+        is thus easily stopped by skin. However, if alpha particles are ingested and sensitive internal organs are irradiated,
         serious damage can be done.
 
-        Beta radiation consists of electrons. Electrons are much lighter and hence are easily scattered.
+        Beta radiation consists of electrons. Electrons are much lighter and are hence easily scattered.
         Moreover, they are much more penetrating since their energy lost per distance traveled is lower.
@@
 
-        Gamma rays interact with matter via photoelectric effects, Compton scattering and pair productions
-        depending on the energy of the photon and the atoms it interact with.
+        Gamma rays interact with matter via photoelectric effects, Compton scattering, and pair productions
+        depending on the energy of the photon and the atom it interact with.
         Since biological tissues mainly consist of elements of low atomic number, Compton
@@
         via photoelectric effects and secondary Compton scattering. As a result, the intensity of the energy deposition falls
-        off much more slowly and is hence more penetrating.
+        off much more slowly, resulting in more penetrating radiation.
 
@@
         \subsubsection{Quantifying radiation on living tissues}
-        The biological impact of radiation depends on the type, power and energy of the radiation, the
+        The biological impact of radiation depends on the type, power as well as energy of the radiation, the
         exposure time, and the part of the body exposed. Various quantities were developed to
         quantify the biological impacts of radiation. Here we review three
-        of these quantities: the absorbed dose, the equivalent dose and the effective dose.
+        of these quantities: the absorbed dose, the equivalent dose, and the effective dose.
 
@@
         radiation energy absorbed per unit mass. The SI unit for absorbed dose is gray (Gy), which is
-        equivalent to $J/kg$.
+        equivalent to J/kg.
 
@@
         \end{equation}
-        The approximate linear dependence on $E_{\gamma}$ is valid from less than $0.1$ MeV to several
+        The approximate linear dependence on $E_{\gamma}$ is valid for energy from less than $0.1$ MeV to several
         MeV and therefore should be good for all sources we are considering.\cite{my68}
@@
         However, the effects of the radiation on living tissue depend on the type of radiation.
-        For example, the same dose of alpha particles deposit more energy in living tissue than gamma rays.
+        For example, the same dose of alpha particles deposits more energy in living tissue than that of gamma rays.
         To quantify the {\it effects} of radiation, we introduce equivalent
@@
         (see Table~\ref{tab:eq} for a list of weighting factors). The unit for equivalent dose is sievert (Sv).
-        $1$ Sv is defined as the effects equivalent to $1$ Gy of Gamma radiation on living tissues. It describes
-        the effects rather than the physical dose received by living tissues.
+        $1$ Sv is defined as the effects equivalent to $1$ Gy of gamma radiation on living tissues. It describes
+        the {\it effects} rather than the {\it physical dose} received by the living tissues.
 
@@
         Therefore, effective dose~($E$) is defined as the equivalent dose weighted by the body part
-        radiated
+        irradiated
         \begin{equation}
@@
         These effects are deterministic with a threshold,
-        meaning that below a certain threshold the effect will not occur and
+        meaning that below a certain threshold the effect will not occur and that
         the severity of the effect depends on the effective dose. Roughly, below the certain threshold,
-        not enough free radicals are created and hence the repairing capabilities of cells can undo the damages.
+        not enough free radicals are created and hence the repairing capabilities of cells can undo the damage.
         If the effective dose is high, the radiation may generate enough free radicals to disrupt the cells or directly
@@
         to offspring.
-        UNSCEAR estimates that risk for severe hereditary effect is about $0.3-0.5$ percent per sievert to
+        UNSCEAR estimates that the risk for severe hereditary effects is about $0.3-0.5$ percent per sievert for 
         the first generation following the radiation exposure.\cite{u16, unscear16}
@@
     apply some of the concepts outlined in the last section to 
-    analyzing the source, composition, and properties of the most detrimental type--the high-level waste.
+    analyzing the source, composition, and properties of the most detrimental type--high-level waste.
     \subsection{Classification of nuclear waste}
@@
     Different organizations have slightly different classification schemes for nuclear waste.
-    However, their definitions of high-level waste (HLW), the most detrimental kind, are relatively consistent.
+    However, their definitions of high-level waste (HLW), the most detrimental kind, are consistent.
     The International Atomic Energy Agency (IAEA) classifies nuclear waste by considering the averaged specific activity
@@
     contaminated by alpha-emitting radionuclides with
-    sufficiently long half-life ($>20$ years) of elements with atomic number $92$ (atomic number of uranium)
+    sufficiently long half-lives ($>20$ years) of elements with atomic number $92$ (the atomic number of uranium)
     or larger and concentration of activity above $3700$ Bq/g.\cite{j83,s01} The activity
-    of TRU is usually low but it remains active for long period of time. Moreover, TRU can be easily turned
+    of TRU is usually low but it remains active for a long period time. Moreover, TRU can be easily turned
     into plutonium-239 (the {\it de facto} material to fabricate atomic bombs), via neutron capture and alpha decay
@@
     Some LLW (for example, metal nuclear fuel packaging) may have a slightly
-    higher radioactivity such that they are classified
+    higher radioactivity such that it is classified
     as intermediate-level waste (ILW). ILW are consolidated in concrete and buried in shallow
@@
     can decay into radon gas which is an alpha-emitter. Hence, the alpha radiation may do additional
-    damage to body if the radon gas enters the body as one breathes.
+    damage to the body if the radon gas enters the body as one breathes.
     Therefore, it is necessary to prevent the uranium tailings
@@
     \subsection{Creation and properties of spent fuel}
+    \label{sec:waste:spentfuel}
     Since all categories but high-level waste are relatively easy to handle,
     we will focus on the formation and
-    properties of spent nuclear fuel, the dominant contribution to high level waste,
-    \label{sec:waste:spentfuel}
+    properties of spent nuclear fuel, the dominant contribution to high level waste.
     \subsubsection{Nuclear fuel}
@@
     with a nucleus.
-    Henceforth, as neutrons travels through materials, they react readily with nuclei.
+    Henceforth, as neutrons travel through materials, they react readily with nuclei.
     For fast neutrons, many reactions are possible but they mostly just scatter off heavy nuclei.
@@
             \caption{Schematic illustration of the two different neutron absorption processes:
-            fission of uranium-235 and the radiative absorption of uranium-238. Adapted from \cite{iync_2013}}
+            the fission of uranium-235 and the radiative absorption of uranium-238. Adapted from \cite{iync_2013}}
             \label{fig:both}
@@
         The neutron capture produces a compound nucleus ${}^{236}U$ that is in an excited state.
-        Sometimes the nucleus can decay by $\gamma$ emission in capture reaction.
+        Sometimes the nucleus can decay by emitting a $\gamma$ photon.
         But most of the time, the excitation energy deforms the nucleus into two positively charged
@@
         continue to induce fission of other uranium-235 nuclei, causing a ``chain reaction'' of subsequent
-        fission that generate a huge amount of energy in a very short time.
+        fission that generates a huge amount of energy in a very short time.
         \begin{figure}
@@
         of special relativity, we can calculate the energy released $E$ from the atomic mass differences between the parent $m_{parent}$
-        and the fission products $m_{products}$.
+        and the fission products $m_{products}$
         \begin{equation}
@@
         In reality, induced fission of uranium-235 creates a variety of products and hence the energy released
-        in a single reaction vary by tens of MeV. On average, about $220$ MeV per fission is recoverable for energy conversion.
+        in a single reaction varies by tens of MeV. On average, about $220$ MeV per fission is recoverable for energy conversion.
         The energy generated by the fission of uranium-235 is 
@@
         the spent fuel)
-        of uranium-235 left along with a lot of free neutrons released from previous fissions. Thus, a large amount
+        of uranium-235 left along with a lot of free neutrons released from previous fission. Thus, a large amount
         of energy is generated via fission of the leftover uranium-235, heating up the spent fuel rods.
@@
         The combination of these properties makes the management and disposal of spent fuel and high-level waste
-        in general very difficult.
+        very difficult.
 
@@
     Spent fuel from nuclear reactors is the major source of HLW.
-    In this section, we will study in details some important processes involved in the short-term management
+    In this section, we will study in detail some important processes involved in the short-term management
     of spent fuel.
@@
     Since spent fuel contains high activity, shielding of radiation is crucial. Shielding works by
-    forcing the radiation to interact with the shielding medium and hence reducing the amount of energy
-    the radiation carries. In the context of nuclear waste management, the most common shielding medium
-    are water and concrete; and the mode of interactions of radiation with those mediums are similar to
+    forcing the radiation to interact with the shielding medium, thereby reducing the amount of energy
+    the radiation carries. In the context of nuclear waste management, the most common shielding media
+    are water and concrete; and the mode of interactions of radiation with those media are similar to
     the physical interactions between radiation and biological molecules (the most abundant of which is
@@
     Suppose on average $R_i$ interactions occur per time unit.
-    The cross section of the interaction between the shielding medium defined as
+    The cross section of the interaction between the shielding medium and the incident beam is defined as
     \begin{equation}
@@
     Short-term cooling of spent fuel is necessary before further disposal or reprocessing can be done.
-    Typically, the spent fuel undergoes cooling in spent fuel cools (Fig.~\ref{fig:pool}) which also shield the radiation.
+    Typically, the spent fuel cools in spent fuel pools (Fig.~\ref{fig:pool}) which also shield the radiation.
 
     Spent fuel pools  are usually about 15-meter deep water pools,
-    with storage rack at the bottom to hold the fuel assemblies. The depth ensures that the radiation
+    with storage racks at the bottom to hold the fuel assemblies. The depth ensures that the radiation
     is well attenuated by the time they reach the surface (see Eq.\ref{eq:atten}).
 
-    The water is enriched with boron atoms that have neutron capture cross section of about $1000$ times
-    the cross section of water molecule. Hence the boron atoms capture the neutrons produced by fission
+    The water is enriched with boron atoms that have a neutron capture cross section of about $1000$ times
+    the cross section of a water molecule. Hence the boron atoms capture the neutrons produced by fission
     in the spent fuel
@@
     of the water.\cite[chapt. 1]{l13}
-    The heat transfer coefficient $h$ depends of the fluid dynamics properties of the water and can
+    The heat transfer coefficient $h$ depends on the fluid dynamics properties of water and can
     be measured experimentally. The goal of
@@
     \end{equation}
-    where $Q\geq 0$ is consistent with the definition above and $S$ denotes entropy, we conclude that $Q$ and $T_{sf}$ always varies in the opposite direction. Rearrange Eq.~\ref{eq:cooling} to get
+    where $Q\geq 0$ is consistent with the definition above and $S$ denotes entropy, we conclude that $Q$ and $T_{sf}$ always vary in the opposite direction. Rearrange Eq.~\ref{eq:cooling} to get
     \begin{equation}
@@
     Hence if we let $dQ/dt=0$ (its min) and $T_{sf}$ be the temperature of the spent
-    fuel rods when it enters the pool (its max), we
+    fuel rods when they enter the pool (its max), we
     can obtain an upper bound on $T_{water}$, the temperature of the water. This temperature turns out to
@@
     After cooling in the spent fuel pools for at least a year, the spent fuel can be stored in dry casks weighing
-    100 tons filled with inert gas. 
+    100 tons and filled with inert gas. 
     The dry casks are often metal cylinders welded to prevent leakage. Each cask is surrounded by
@@
         predicting the performance of some
-        engineered product over the timescale of a million years
-        experience. Over time, the radiation from the waste will also alter the rock (by altering its composition
+        engineered product over the timescale of a million years.
+        Over time, the radiation from the waste will also alter the rock (by altering its composition
         and creating fractures) and hence the water percolation rate.
@@
         fifteen million and eight million years ago. However, it is the small-volume eruptions, the last of
-        which occurred about $75,000$ years ago, that would the most likely pose risk to the repository
+        which occurred about $75,000$ years ago, that would most likely pose risk to the repository
         in its life span of a few million years.
@@
         However, if the uncertainty of the input parameter is expressed as a distribution, or if $y$ is
-        not an analytic function, then the propagation of uncertainty method fails. In our volcano disruption
+        not an analytic function, then this method fails. In our volcano disruption
         problem, $y$ can be seen as the complex computational model like TSPA. However, there is a
@@
         from observed data. The choice of time range is not trivial: longer time range certainly gives us
-        more data and hence improved statistics, but longer time interval dated far back may fail to represent
+        more data and therefore improved statistics, but a longer time range reaching far back may fail to represent
         future event rates since the geology evolves. The best approach seems to be choosing the time interval
@@
         $1.5\times 10^{-6}$ to $8\times 10^{-6}$ events per year. These estimates translate to an estimated
-        time between events from 125 thousand years to 250 thousand years. We can see that the choices of
-        models introduce a significant uncertainty to the estimation.\cite{cv14}
+        time between events from 125 thousand years to 250 thousand years. Therefore, the choice of
+        model can vary the estimation result significantly.\cite{cv14}
 
@@
  Reprocessing techniques can significantly reduce the volume of the waste and make it more
-manageable. With reprocessing, deep boreholes storage may become a more feasible disposal solution 
-for commercial reactor waste.
+manageable.
 

### 2018-01-13 presentation/demo.tex
Cleaned up and added presentation.

\documentclass[10pt]{beamer}
\usepackage{pgfpages}
\setbeameroption{show notes}
\setbeameroption{show notes on second screen=right}
\usetheme[progressbar=frametitle]{metropolis}
% workaround to set theme back with pdfpages notes
\makeatletter 
\def\beamer@framenotesbegin{% at beginning of slide
     \usebeamercolor[fg]{normal text}
      \gdef\beamer@noteitems{}% 
      \gdef\beamer@notes{}% 
}
\makeatother

\usepackage{natbib}
\usepackage{appendixnumberbeamer}
\usepackage{pbox}
\usepackage{booktabs}
\usepackage[scale=2]{ccicons}
\usepackage{caption}
\usepackage{pgfplots}
\usepgfplotslibrary{dateplot}
\usepackage[]{amsmath}
\usepackage{xspace}
\newcommand{\themename}{\textbf{\textsc{metropolis}}\xspace}
\bibliographystyle{plain}
\title{Nuclear Waste: Too Hot to Touch}
\subtitle{Physics Comps Presentation}
\date{\today}
\date{}
\author{Yuping Huang}
\institute{Department of Physics and Astronomy, Carleton College}
\titlegraphic{\hfill\includegraphics[height=1.5cm]{carleton.pdf}}
\AtBeginSection[]
{
   \begin{frame}{Outline}
  \setbeamertemplate{section in toc}[sections numbered]
       \tableofcontents[currentsection]
   \end{frame}
}
\begin{document}
\begin{frame}
\begin{figure}
\includegraphics[height=\textheight]{last}
\caption*{Image credit XKCD.com}
\end{figure}
\end{frame}
\maketitle
\note[itemize]{
\item thanks and welcome
\item set the scene--some number
}
\begin{frame}{Introduction: why does it matter?}
\begin{itemize}
	\item World nuclear energy consumption will increase by 40\% by 2030
    \item Over \alert{10,000} metric tons of highly radioactive waste generated annually
    \item Highly radioactive waste may remain hazardous for up to millions of years
    \item We will have over \alert{457,000} tons of highly radioactive waste by 2020 without any foreseeable disposal
    scheme
\end{itemize}
\end{frame}
\note[itemize]{
\item 40\% increase energy
\item $10,000$ tons annually
\item harzadous
\item \alert{457,000} tons by 2020
}
\begin{frame}{Spent fuel pool ``near miss'' at Fukushima}
\only<1>{
		\begin{figure}
        \centering
		\includegraphics[width=0.8\textwidth]{vapor}
		\caption*{Water vapor from hydrogen burning in a reactor building at Fukushima}
		\end{figure}
        }
\only<2>{
		\begin{figure}
        \centering
		\includegraphics[width=0.8\textwidth]{exploded}
		\caption*{The explosion in the reactor building}
		\end{figure}
        }
\only<3>{
		\begin{figure}
        \centering
		\includegraphics[width=0.8\textwidth]{fukushima_nocool}
		\caption*{The Japan Times article on the cooling halt of the spent fuel pool}
		\end{figure}
        }
\end{frame}

\begin{frame}{The nightmare scenario}
\note[item]{1 megabecquerel per square meter -- trigger relocation order}
\note[item]{jan, apr, jul, oct, 2015}
\note[item]{convince you of the gravity of this topic}
\begin{figure}
\centering
\only<1>{\includegraphics[width=0.6\textwidth]{jan}\caption*{Jan, 2015}}
\only<2>{\includegraphics[width=0.6\textwidth]{oct}\caption*{Oct, 2015}}
\end{figure}
An estimate of the spread of Cesium-137 radioactive cloud if the spent fuel pools catch fire at the Peach Bottom nuclear power plant in Pennsylvania at different time of the year. \cite{s16}
\end{frame}

\begin{frame}{Outline}
  \setbeamertemplate{section in toc}[sections numbered]
  \tableofcontents[hideallsubsections]
\end{frame}

\section{Radiation and Doses}
\begin{frame}{Unstable nuclei}
  \begin{columns}[T,onlytextwidth]
    \column{0.4\textwidth}
      \begin{itemize}
      	\item Configuration decides stability
      	\item Binding energy = Coulomb + Nuclear
      	\item Minimum energy at $Z\approx N$
      \end{itemize}

    \column{0.6\textwidth}
		\begin{figure}
		\centering
		\includegraphics[height=0.8\textheight]{stability.png}
		\caption*{half-lives of isotopes with different neutron and proton numbers. (Image credit Wikipedia)}
        \end{figure}
  \end{columns}
\end{frame}
\note[itemize]{
\item binding energy as a function of number of protons and neutrons, nuclear force + Coulomb force
\item stable configurations minimize the function
\item as there are more protons. need neutrons to weaken the long-range Coulomb repulsion
}

\begin{frame}{Radioactive decay}
\note[item]{set N=1/2$N_0$ to get half-life}
\note[item]{subsequent decay of daughter nuclei could happen which makes things harder; if half life of daughter much shorter then we can treat it as one exponential; otherwise there will be some equilibrium of daughter nuclei and their subsequent decay products}
\begin{itemize}[<+->]
\item Exponential decay: constant probability of decay $\lambda$ from unstable high
energy nucleus state to lower energy state
\[\lambda = -\frac{(dN(t)/dt)}{N(t)} \Rightarrow N(t)=N_0 e^{-\lambda t}\]
\item Half-life $t_{1/2}=\frac{\ln 2}{\lambda}$: the time required for half of the sample to decay on average
\item{(Radio)activity: number of decays per unit time}
\[\mathcal{A}(t)\equiv \lambda N(t)\]
\item Units: becquerel (Bq) = 1 decay/s or curie (Ci) = $3.7\times 10^{10}$ Bq
\item Specific activity: number of decays per unit time per mass
\end{itemize}
\end{frame}

\begin{frame}{Modes of decay}
\begin{itemize}[<+->]
\item $\alpha$ decay: emits helium-4 atom ${}^4_2$He
\[{}^A_Z X_N \rightarrow {}^{A-4}_{Z-2}X'_{N-2} + \alpha.\]
\item $\beta^-$ decay: emits electrons ($\beta$ radiation) and neutrinos  
\[{}^A_Z X_N \rightarrow {}^A_{Z+1} X'_{N+1} + e^- + \bar{\nu}_e\]
\item $\gamma$ decay: normally occurs after other forms of decays for nuclei in excited state,
emits $\gamma$ ray in the range of $0.1$ to $10$ MeV
\end{itemize}
\end{frame}

\begin{frame}{Half-life and specific activity of radioactive isotopes}{fragile}
  \begin{table}
  \centering
    \begin{tabular}{llll}
      \toprule
      Isotope & Half-life & Specific Activity (Ci/g) \\
      \midrule
      Cesium-137 & 37 years & $88$\\
      Radium-226 & 1600 years  & $1$\\
      Plutonium-239 & 24,000 years & $0.063$ \\
      Uranium-233 & 1.6 million years & $9.3\times 10^{-3}$\\
      Thorium-232 & 14 billion years & $1.1\times 10^{-7}$ \\
      \bottomrule
    \end{tabular}
  \end{table}
  NB: The isotope used for computerized tomography (CT) scan has activity $\sim 10^{-2}$ Ci.
\end{frame}
\note[itemize]{
\item cesium-137 most concerning
\item plutonium makes atomic bombs
\item u-232 long half life and not safe level of radiation
\item thorium-232 can go thru neutron capture to form uranium-233
}

\begin{frame}{Health Effects and Dose}
\begin{itemize}[<+->]
\item Radiation deposit energy and ionize biological molecules, e.g. the DNA
\item \emph{Equivalent dose} $H_R$: energy deposited by the radiation per mass weighted by the type of radiation
\[H_R = w_R \times \frac{E_R}{m_b},\]
$E_R$: energy of radiation, $m_b$ mass of recipient, $w_R$: weighting factor of radiation type.
\item Unit: sievert (Sv) = J/kg, equivalent to the effect of $1$J of x-rays on $1$kg of biological mass
\item For $\gamma$ radiations, the equivalent dose rate is approximately
\[\frac{dH_R}{dt}(\mu\text{Sv h}^{-1}) \approx \frac{\mathcal{A}\text{(MBq)}\times E_{\gamma}\text{(MeV)}}{6\times [r\text{(m)}]^2}\]
\end{itemize}
\end{frame}

\begin{frame}{Weighting factor $w_R$}
The equivalent dose $H_R = w_R \times \frac{E_R}{m_b}.$
\begin{table}
\begin{tabular}{l c c}
\toprule
Type of radiation & Energy range & Weighting factor, $w_R$\\
\hline
Photons, electrons & All energies & 1\\
Neutrons & $<10$ keV & 5 \\
& $10-100$ keV & 10 \\
& $100$ keV-$2$ MeV & 20 \\
& $2-20$ MeV & 10 \\
Protons & $<20$ MeV & 5 \\
$\alpha$ particles& & $20$\\
\bottomrule
\end{tabular}
\caption*{Weighting factors $w_R$ for different types of radiations. \cite{icrp74}}
\end{table}
\end{frame}
\begin{frame}{Cesium-137}
\note[item]{Image shown is the city of Priyat in Ukraine which is inhabitable but is now open to tourist visit}
\note[item]{Dose level still 80 times higher than the recommended maximum for external irradiation of the human body}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{Priyat}
\caption*{After the Chernobyl accident in which 27 kg of cesium-137 were released, a 4,000 km${}^2$ area still currently contains too much cesium-137 to be inhabited (with dose rate $1\mu$Sv/h)}
\end{figure}

\end{frame}

\begin{frame}{Long half-life vs short half-life}
\begin{figure}
\centering
\only<1->{
\includegraphics[width=0.6\textwidth]{csur.png}}
\note[item]{0.01 Ci is enough for medical radiation diagnosis like (CT)}
\note[item]{it's not gonna kill you immediately, but it is not safe}
\only<1>{\caption*{Activity of 1 gram of cesium-137 vs uranium-233 over time, where blue indicates cesium-137 and green uranium-233}}
\only<2>{\caption*{\alert{Conclusion: Long half-life isotopes can be tricky to handle}}}
\end{figure}
\end{frame}

\section{Sources of Nuclear Waste}
\begin{frame}{Categories of nuclear waste (defined by the NRC) \cite{s01}}
\begin{itemize}[<+->]
\item \alert<13>{High level waste}
	\begin{itemize}
    \item typically comes from spent fuel of nuclear reactor
	\item generates large amount of heat and radiation
    \item contains isotopes of varying half life
    \end{itemize}
\item Transuranic waste
	\begin{itemize}
    \item materials contaminated with artificially made radioactive isotopes that have atomic number larger
    than uranium
    \item mostly comes from the military
    \item low activity but long half life
    \end{itemize}
\item Low-level waste
	\begin{itemize}
    \item Low activity waste. Examples are papers and tools used in cleanup operations in the reactor.
    \end{itemize}
\item Uranium tailings
	\begin{itemize}
	\item radioactive thorium and radium left in the uranium mine after the uranium is extracted
    \end{itemize}
\end{itemize}
\end{frame}

\note[itemize]{
\item tailing: radium can decay into radon gas which is an alpha-emitter; be careful not to pollute water and air
}

\begin{frame}{What happens to nuclear fuel: neutron capture}
\note{Thermal neutrons means most probable temperature at $17\deg$C}
\begin{itemize}[<+->]
\item Fresh nuclear fuel: 97\% uranium-238 and 3\% uranium-235
\item The fuel rod is irradiated with \alert{thermal ($0.025$ eV)} neutrons to induce neutron absorption
\item Two main outcome: fission and radiative capture
\end{itemize}
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{rod}
\end{figure}
\end{frame}

\begin{frame}{Nuclear fission}
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{fission}
\caption*{Schematic illustration of the thermal neutron induced fission of uranium-235. \cite{l01}}
\end{figure}

\end{frame}
\note[itemize]{
\item Thermal neutron gets absorbed; the nucleus becomes unstable and split into two parts
\item the two parts repel; KE of neutron converted to vibrational energy
\item as the amplitude gets higher nuclear force potential drops off sharply and coulomb potential
dominates
\item boom. gigantic binding energy difference produces energy.
}

\begin{frame}{Radiative capture}
\begin{itemize}[<+->]
\item For a nucleus that is not fissile, it absorbs a neutron, form a heavier nuclei and radiates to return to ground state
\item uranium-238 in nuclear fuel can generate plutonium-239
\[{}^{238}\text{U} \xrightarrow{\text{radiative capture}} {}^{239}\text{U} \xrightarrow{\text{half-life 23 min}} {}^{239}\text{Np}\xrightarrow{\text{half-life } 2.4 \text{ d}} {}^{239}\text{Pu}\]
\item Thorium-232 also goes through radiative capture to form uranium-233
\end{itemize}
\begin{figure}
\centering
\includegraphics[height=0.5\textheight]{pu}
\end{figure}
\end{frame}

\begin{frame}{Composition of spent fuel}
\begin{itemize}
\item Nuclear fission produces fission products like cesium-137, strontium-90 with half-lives shorter than
40 years
\item Neutron capture produces uranium-233, plutonium and other transuranic elements (minor actinides) with half-lives up to
millions of years
\end{itemize}
\begin{figure}
\centering
\only<1>{\includegraphics[height=0.5\textheight]{pu}}
\only<2>{\includegraphics[height=0.5\textheight]{spent-fuel}}
\only<2>{\caption*{Composition of spent fuel from light-water reactor; exact composition varies by reactor designs}}
\end{figure}
\end{frame}

\section{Nuclear Waste Management}
\begin{frame}{Shielding radiations}
	\note[item]{SLOW DOWN}
    \note[item]{Interaction probability is independent of distance traveled}
    \note[item]{Decay probability is independent of the time passed}
	\begin{itemize}[<+->]
    	\item For spent fuel, radiations and neutrons need to be shielded
    	\item Shielding forces the radiations to interact with the shielding media
        \item Nuclear cross section $\sigma$ quantifies the effective interaction probability
        \item The interaction probability is independent of the distance traveled
        \item Intensity $I(z)$ of incident beam falls off exponentially as thickness $z$ increases
        \[I(z)=I(0) e^{-\mu z}, \mu \propto \sigma.\]
    \end{itemize}
\end{frame}

\begin{frame}{Nuclear cross section\ footnote{section adapted from \cite{k88}}}
	\note[item]{erase the board and do dimensional analysis}
	\note[item]{tells you how many reactions can happen with one nucleus}
    \note[item]{make the point about 10 times as thick water}
	\begin{itemize}[<+->]
	\item Nuclear cross section characterizes reaction probability and has the unit of area per nucleus
    \[\sigma = \frac{R}{I_i N},\]
    where $I_i$ is incident beam intensity, $R$ is the number of reaction occurred per time, $N$ the number of nuclei in the target
	\item Typical nuclei radii are $\sim 10^{-14}$ and have effective area $\pi r^2\sim 10^{-28}=1$ barn
    \item Boron nucleus neutron capture has $\sigma = 3840$ barns
    \item $\gamma$ ray interaction with water has $\sigma \sim 0.01$ barns, with concrete $\sigma\sim 0.1$ barns
    \end{itemize}
\end{frame}

\begin{frame}{Spent fuel pools}
\note[item]{heat drops by more than 90\% for the first year}
\note[item]{Two years is enough but there aren't enough subsequent storage space available}
  \begin{columns}[T,onlytextwidth]
    \column{0.5\textwidth}
      \begin{itemize}[<+->]
      	\item Removed spent fuel typically spend 5 years or more in spent fuel pools
        \item Filled with boron-rich water because boron atoms have a larger neutron capture cross-section
        \item The water is circulated and cooled at $50^{\circ}$C
        \item At least 15-meter deep to shield gamma rays
      \end{itemize}

    \column{0.5\textwidth}
		\begin{figure}
		\includegraphics[width=0.9\textwidth]{fukushimafuel-rods}
		\caption*{Spent fuel pool at Fukushima}
		\end{figure}
  \end{columns}
\end{frame}

\begin{frame}{Dry cask storage}

  \begin{columns}[T,onlytextwidth]
    \column{0.5\textwidth}
      \begin{itemize}[<+->]
      	\item After 5 years in the pool, the spent fuel can be transfered to dry cask storage
        \item Dry casks are sturdy metal and concrete structures designed to withstand harsh conditions
        \item Air is circulated to cool the cask
        \item The NRC believes that dry casks storage will be safe for at least 100 years
      \end{itemize}

    \column{0.5\textwidth}
		\begin{figure}
		\includegraphics[width=0.9\textwidth]{dry_storage.jpg}
		\caption*{A dry cask storage stie}
		\end{figure}
  \end{columns}
\end{frame}
\section{Nuclear Waste Disposal}
	\begin{frame}{Deep Geological Repository}
    \begin{itemize}[<+->]
    \item The goal of nuclear waste disposal is to minimize radiation exposure for over $10^6$ years and to prevent proliferation
    \item The science community generally accepts that deep (over $300$m) underground disposal for high-level waste is much safer than above-ground storage
    \item Other alternatives include transmutation and space storage, which are too costly
    \end{itemize}
    \end{frame}

    \begin{frame}{Yucca Mountain Nuclear Waste Repository}
    \note[item]{established Nuclear Waste Policy Act of 1987}
    \note[item]{later became the only site studied for feasibility}
    \note[item]{the Department of Energy was directed not to develop complementary approaches}
    \note[item]{risks: oxidation from moise air; water percolation to water table; hot upwelling water from below}
    \only<1>{
    \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{yuccaRepo}
    \caption*{Scheme of the Yucca Mountain Nuclear Waste Repository; the depth of the repository is about $300$ meters}
    \end{figure}
    }
    \only<2->{
    \note[item]{prevent oxidation is the responsibility of the packaging of the waste, but it is hard to predict its integrity for over 1 million years}
    Risk factors:\\
    \begin{itemize}
    	\item Oxidation of the waste from moist air
        \item Water percolation from the surface to the water table
        \item \alert<2>{Tectonic motion and climate change}
        \item Hot upwelling water from below
        \item \alert<3>{Catastrophic events}
    \end{itemize}
    }
    \end{frame}
	\begin{frame}{Volcanism: risk assessment}
    \note[item]{disruptive volcanic eruption releasing radioactive material into the air}
    \note[item]{observation and inferences; human choices; computational models}
    \centering
        \begin{figure}
		\includegraphics[height=0.6\textheight]{volcano}
        \caption*{A flowchart for the volcanism risk estimation problem. The oval-shaped parts are observations and statistical inferences, the rectangular ones human choices, and the hexagon the DOE computational model. \cite{me06}}
		\end{figure}
    \end{frame}
	\begin{frame}{How to deal with uncertainty?}
    	\only<1>{
        \begin{figure}
		\includegraphics[height=0.8\textheight]{meme}
		\end{figure}
        }

        \only<2>{
    	\begin{figure}
		\includegraphics[height=0.8\textheight]{bill}
		\caption*{Propagation of uncertainty. Courtesy of Bill Titus}
		\end{figure}
        }
        \only<3>{
        \alert{BUT}
        \begin{itemize}
        \item We don't necessarily have an analytical expression for the model
        \item The linear approximation might fail if the uncertainties of the input parameters
        are large
        \item The error bar may not contain all the information about the uncertainty
        \end{itemize}
        }
    \end{frame}

    \begin{frame}{Monte Carlo simulations}
    	Suppose we have a function $y(x_1,x_2,...,x_n)$ that is not necessarily analytic (for example, $y$ can be the expected leaked dose calculated from a simulation with a bunch of input parameters).\\
        \begin{itemize}
            \item Generate a large number of $(x_1,x_2,...x_n)$ samples according to their respective distributions.
            \item For each $(x_1,x_2,...,x_n)$ sample, calculate a $y$ value (with the computational model).
            \item Aggregate all the $y$'s generated to estimate its distribution or uncertainty.
        \end{itemize}
    \end{frame}

%    \begin{frame}{Monte Carlo simulations: an example}
%		\begin{figure}
%		\includegraphics[width=0.9\textwidth]{Monte-Carlo-laundry-example}
%		\caption*{An example of Monte Carlo simulation to infer environmental impact of laundry based
%        on a user's laundry habit}
%		\end{figure}
%    \end{frame}
    \begin{frame}{Monte Carlo simulation: an example}{fragile}
    \only<1>{
        \alert{You are in a lab.}
    	\begin{itemize}
        \item Task: find out how many nuclei will remain after $t=200$s for the sample you have
        \item You \alert{don't} know the exponential decay law $N(t)=N_0 e^{-\lambda t}$
        \item You have: a program \textsc{NOTNDSOLVE} takes in $N_0$, $\lambda$, $t$ and predicts $N(t)$
        \item Measurements made: $N_0= (1\pm 0.2)\times 10^{27}$, $\lambda = 0.003\pm 0.001$/s
        \item Assume each parameter is uniformly distributed inside its error bar 
        \end{itemize}
      }

      \only<2>{
      \alert{Here's what you do}\\
      Repeat the following $10,000$ times
      \begin{itemize}
      \item Generate a random $N_0$ and a random $\lambda$ from their respective distributions
      \item Calculate $N(t=20s)$ using $NOTNDSOLVE$
      \end{itemize}
      Now plot a histogram of all the $N(t=20s)$ numbers you have. It quantifies the uncertainty of your prediction
      }

      \only<3->{
      Probability distribution for the predicted $N(t=200s)$
      \begin{figure}
      \centering
      \only<3>{
      \includegraphics[height=0.7\textheight]{nList}
      \caption*{$10^4$ samples;Mean: $5.52\times 10^{26}$; standard deviation $5.53\times 10^{26}$}
      }
      \only<4>{
      \includegraphics[height=0.7\textheight]{more}
      \caption*{$10^6$ samples; Mean: $5.52\times 10^{26}$; standard deviation $9.04\times 10^{25}$}
      }
      \end{figure}

      }
    \end{frame}

    \begin{frame}{Volcanic disruption problem revisited}
        \begin{figure}
		\includegraphics[height=0.6\textheight]{volcano}
        \caption*{A flowchart for the volcanism risk estimation problem. The oval-shaped parts are observations and statistical inferences, the rectangular ones human choices, and the hexagon the DOE computational model. \cite{me06}}
		\end{figure}
    \end{frame}

    \begin{frame}{The volcanism probability: NRC vs DOE}
    \note[item]{on the edge situation}
    \note[item]{if the expected leaked dose is really high, then the risk is higher than when the leaked dose is low}
    \begin{itemize}[<+->]
    	\item EPA guideline: if the disruptive event rate is lower than \alert{$10^{-8}$} events per year then the risk is negligible
    	\item The DOE estimates an disruptive volcanic eruption rate of $1.5\times 10^{-9}$ events per year
        with 90 percent confidence bound \alert{$5.4\times 10^{-9}$-$4.9\times 10^{-8}$} events per year

        \item The NRC uses a subset of the DOE's models that reflect conservatism and concludes a disruption rate of \alert{$10^{-7}$ - $10^{-8}$} events per year

        \item In an edge case like this, we need to understand better the effects of the eruption on the leaked dose to assess the risk
    \end{itemize}
    \end{frame}

\begin{frame}{Status of Yucca Mountain}
\note[item]{Obama administration defunded the project in 2011}
Steven Chu, former secretary of energy said in 2009 \cite{aa12}:\\
\vspace{0.5cm}
``Yucca Mountain as a repository is off the table. What we’re going to be doing is saying, let’s step back. We realize that we know a lot more today than we did 25 or 30 years ago...(dry cask storage) gives us time to figure out what we should do for a long-term strategy.''\\
\vspace{0.5cm}
In 2011, the project was defunded by the Obama administration
\end{frame}

\begin{frame}{Newsflash}
\only<1>{
\begin{figure}
\centering
\caption*{Las Vegas Review Journal, Jan 19, 2017}
\includegraphics[width=0.8\textwidth]{rickperry.png}
\end{figure}
}
\only<2>{
\begin{figure}
\centering
\caption*{Las Vegas Review Journal, Jan 31, 2017}
\includegraphics[width=0.8\textwidth]{fuck.png}
\end{figure}
}
\end{frame}

\section{Discussions}

\begin{frame}{Discussions: what else can be done?}
\begin{itemize}[<+->]
\item Deep underground storage is complicated, both politically and scientifically
\item Dry cask storage can buy the US a few decades to come up with different solutions (e.g. fast neutron reactors)
\item France is building a reversible long-term ($\sim 100$ years) underground storage
\item Sweden and Finland started constructing deep geological repositories
\end{itemize}

\end{frame}

\begin{frame}[allowframebreaks]{References}
  \bibliography{refs}

\end{frame}

\begin{frame}{Acknowledgment}
\begin{itemize}
\item Thanks to you for coming
\item Thanks to Yuhao Wan, Sam McHugh, and Gaston for attending the practice talk
\item Thanks to advisors Joel Weisberg and Alex McMurty
\end{itemize}
\end{frame}

\begin{frame}[standout]
\usebeamercolor[bg]{normal text}
Questions?
\end{frame}

\end{document}
